---
title: "Research Rapport Results"
format:
  html:
    css: style.css
---

This document will produce all the figures and tables that will be used in the research rapport. It also will serve as a compilation of all code that has been written thus far. 

# 1. Purpose + proceedings
The purpose of this thesis is twofold. First, it aims to show the shortcomings of current methods and second it aims to propose a new method that overcomes these shortcomings:

1. We first need to show that current SOTA methods such as the RI-CLPM and the DPM are not equipped to handle baseline confounding effects that change over time. 
2. Second, we need to show that adjustment per occasion in a linear fashion results in biased estimates when the baseline confounding effects are non-linear AND change over time.

We need to run two simulation studies to show these points. The first simulation study will only focus on the inability of the SOTA methods to handle varying effects of baseline confounders. It will show that we need to explicitly model the confounder effects in order to get unbiased estimates of the cross-lagged effects. The second simulation study will then expand on the first by including non-linear confounder effects and showing that linear adjustment per occasion results in biased estimates of the cross-lagged effects. In this second study, we also need to include the integration of machine learning methods with residualisation to flexibly model the confounders effect on each occasion. An added advantage is that this approach does not suppose that we know the true functional form of the confounder effects, but instead we can learn them from the data. 

## 1.1 Conditions
There are various simulation conditions that need to be considered. Note that for this study, we will not vary all conditions, but focus on defining the problem and providing a proof of concept for our proposed solution. We will briefly consider the following conditions:

* 1.1.1 Type of time-varying effect 
* 1.1.2 Strength of confounding effect (and covariation among the confounders)
* 1.1.3 Sample size
* 1.1.4 Number of occasions
* 1.1.5 The values of the cross-lagged effects, auto-regressive effects, and residual correlations
* 1.1.6 Type of the non-linear effect (only for study 2)
* 1.1.7 Ammount of non-linearity (only for study 2)
* 1.1.8 Methods used 

### 1.1.1 Type of time-varying effect
There are various ways in which an effect can change over time. Some of these are more realistic than others. For this simulation study, we will consider the following types of time-varying effects:

* Constant over time (i.e., no time-varying effect). This will serve as the baseline condition.
* Linear increase/decrease over time. At every occasion, a constant c is added/subtracted to/from the effect. For example, the effect of age on social media (or genAI) use might decrease over the years as it is integrated into daily use more. 
* Circular/sinusoidal effect over time. At every occasion, the effect of the confounder increases and decreases in a sinusoidal pattern. For example, the effect of the weather on mood might follow a seasonal pattern.
* Stepwise effect over time. A change in effects at a specific occasion due to some one-off event. E.g. the effect of age on social media usage changes drastically after COVID-19 lockdowns.
* Random walk. The effect is drawn from a normal distribution centered around the previous occasion. As such, effects will slowly drift into a direction over time, for example due to societal changes (concept drift). 

In order to determine how much variation there is between the occasions, it would be neat to have some measure of variation. For example, we could calculate the standard deviation of the effects across occasions. This would give us an idea of how much the effects vary over time. Then, we could use this measure to determine make sure all conditions have similar levels of variation. We can compute this measure as follows:

```{r effect variation function}

# function that takes in the matrix B, which holds the effects of the confounders on X and Y
# where rows = X, Y and columns = occasions
confounder_variation <- function(B) {
  
  # Check input
  if (!is.matrix(B) || nrow(B) != 2) {
    stop("B must be a 2 x k matrix (rows = X, Y; columns = occasions).")
  }
  
  # Compute SD across occasions for each row
  sd_X <- stats::sd(B[1, ])
  sd_Y <- stats::sd(B[2, ])
  
  # Optionally: pooled SD across all betas (flatten matrix)
  sd_overall <- stats::sd(as.vector(B))
  
  # Return results
  list(
    sd_X      = sd_X,
    sd_Y      = sd_Y,
    sd_overall = sd_overall
  )
}
 
```

### 1.1.2 Strength of confounding effect
Confounders affect the variables of interest (X and Y). These variables receive an ammount of variance from the confounders. Since the variance of each variable is 1, this variance can be calculated from the regression coefficients directly. This is useful, because we can make all scenarios have the same level of confounding by ensuring that the R squared of the confounders on X and Y is the same across conditions. For the research rapport, we will consider only 1 level of confounding, but later this may be changed. We can calculate the variance received from the confounders as follows:

$$ 
R^2 = B \Psi B^T = B I B^T = B B^T,
$$

Where $B$ is the matrix of effects of the confounders on X and Y, and $\Psi$ is the covariance matrix of the confounders (which is the identity matrix if the confounders are uncorrelated and have variance 1). We can denote the variance in our variables of interest that is explained by the confounders as $R^2$. Note that we could choose to make the confounders correlated, but for the research rapport we will assume them uncorrelated. Also note that (as will be elaborated upon later) the non-linear parts of each confounder will be orthogonalised w.r.t. their linear 'parent' meaning the covariance between the linear and non-linear component is 0. We can use the following script to determine the strength of the confounding effects from the matrix $B$:

```{r confounder strength function}

# function to calculate the strength of confounding effects
# from the matrix B, which holds the effects of the confounders on X and Y
# where rows = X, Y and columns = confounders
confounder_strength <- function(B) {
  
  # check if the input is a matrix with 2 rows (for X and Y)
  if (!is.matrix(B) || nrow(B) != 2) {
    stop("B must be a 2 x k matrix (rows = X, Y; columns = confounders).")
  }
  
  # with Psi = I, S_U = B %*% t(B)
  S_U <- B %*% t(B)
  
  # extract variances and covariance
  var_X <- S_U[1, 1]
  var_Y <- S_U[2, 2]
  cov_XY <- S_U[1, 2]
  
  # save results in a list
  list(
    S_U   = S_U,
    var_X = var_X,
    var_Y = var_Y,
    cov_XY = cov_XY
  )
}

```

### 1.1.3 Sample size
We will choose various sample sizes that range from small (i.e. 50-100), to large (1000-5.000), to population level (1000000+). This is important for two reasons. 

1. The first is that at small sample sizes, estimation of some models may show much greater variability. This is problematic, since estimates may be correct on average, but we are in a scenario where they are not very often. This is especially important when using machine learning methods, since these often require larger sample sizes to perform well.
2. The second is that we would like to see how well the methods perform at population level. This is important, since we would like to see that the methods converge to the true values when sample size increases.
3. An extra bonus point is that sample sizes heavily influence computational time. This is important both for the CLPMS and the machine learning methods.

For this research report, we will consider only a sample size of 1000. 

### 1.1.4 Number of occasions
The number of occasions influences our problem in three ways:

1. More points mean less information per point. This means that we need a larger sample size to get stable estimates.
2. More points mean more opportunity for time-varying effects to change. This means that we might see more bias the further in time we go from the baseline. 
3. This in turn means that our added variance will increase, which potentially causes the stationarity equation to fail. 

For this research report, we will consider only 5 occasions.

### 1.1.5 Cross-lagged & auto-regressive effects + residual correlations
It is important to pick realistic values for the substantively interesting parameters. Usually, the autoregressive effect is larger than the cross-lagged effect, and lies somewhere around 0.2-0.3. The cross lagged effects themselves are usually smaller, around 0.1-0.2. The residual correlations are usually quite large, and lie around 0.2-0.3. They refer to confounders that affect both variables at one time-point, but not across time, such as the weather. We will not vary these parameters over conditions in the research rapport. 

### 1.1.6 Type of the non-linear effect (only for study 2)
We will consider various types of non-linear effects from the confounders on the variables of interest. All of the types used for the research rapport will be generated in the following way:

1. Generate the linear components of the confounders as standard normal variables.
2. Generate the non-linear components of the confounders by applying a non-linear transformation to the linear conponent
3. Orthogonalise the non-linear component w.r.t. the linear component to ensure that they are uncorrelated. It will become apparent later why this is important.

We have chosen to add the following four kinds of non-linear transformations here:

* Exponential: $f(x) = exp(x)$
* Quadratic: $f(x) = x^2$ + Cubic: $f(x) = x^3$
* Sine: $f(x) = sin(x)$
* Tanh: $f(x) = tanh(x)$

These forms were chosen because they allow for independence of their linear component without messing up the interpretation of that linear component. There are some forms of non-linearity very common in the social sciences that are to be added for the final thesis, which are not included here for the research rapport. We will add these forms later, but these require special attention as they cannot be orthogonalised w.r.t. their linear component without messing up the interpretation of that linear component. Examples are:

* piecewise linear functions (threshold effects)
* interactions (two-way, three-way, etc.)

Below are some plots visualising each type of non-linearity used in this study. 

```{r visualising non-linearity, echo=FALSE, message=FALSE, warning=FALSE}

# libraries
library(ggplot2)
library(patchwork)

# reproducibility
set.seed(123)

# generate x
n  <- 200
x  <- seq(-3, 3, length.out = n)

# data for each plot
df_exp <- data.frame(
  x = x,
  y = exp(x) + rnorm(n, 0, 0.3),
  fn = "Exponential"
)

df_qc <- data.frame(
  x = x,
  y = x^2 + x^3 + rnorm(n, 0, 2),
  fn = "Quadratic + Cubic"
)

df_sin <- data.frame(
  x = x,
  y = sin(x) + rnorm(n, 0, 0.1),
  fn = "Sine"
)

df_tanh <- data.frame(
  x = x,
  y = tanh(x) + rnorm(n, 0, 0.05),
  fn = "Tanh"
)

# assign a distinct color per plot
plot_colors <- c(
  "Exponential"       = "#1f77b4",
  "Quadratic + Cubic" = "#d62728",
  "Sine"              = "#2ca02c",
  "Tanh"              = "#9467bd"
)

# plot maker
make_plot <- function(data, title) {
  ggplot(data, aes(x, y)) +
    geom_point(color = plot_colors[unique(data$fn)], alpha = 0.5, size = 1.3) +
    geom_smooth(method = "lm", se = FALSE, color = "black", linewidth = 0.9) +
    theme_minimal(base_size = 12) +
    labs(title = title, x = "", y = "")
}

p1 <- make_plot(df_exp,  "Exponential")
p2 <- make_plot(df_qc,   "Quadratic + Cubic")
p3 <- make_plot(df_sin,  "Sine")
p4 <- make_plot(df_tanh, "Tanh")

# combine in a 2×2 grid
(p1 | p2) /
(p3 | p4)

```

### 1.1.7 Ammount of non-linearity (only for study 2)
Even though we have various forms of non-linearity, we would like the ammount of non-linearity between functions to be comparable. Lucklily, we can straightforwardly calculate the ammount of variance that is explained by the non-linear component of each confounder as $B B^T$, where $B$ is the matrix of effects of the non-linear components of the confounders on X and Y. We can then compute the R squared values for the linear component and the non-linear component separately. This is done so easily, because the linear components and the non-linear components are orthogonalised w.r.t. each other. Now we can compute the measure of non-linearity as follows:

$$
\eta = \frac{R^2_{non-linear}}{R^2_{linear} + R^2_{non-linear}},
$$

where $\eta$ is a number between 0 and 1 that denotes the share of the variance explained by the non-linear components of the confounders in the variables of interest. We can append a subscript $t$ to denote the explained variance per occasion. Our general measure of non-linearity across all occasions can then be computed as:

$$
\eta_{t} =\frac{R^2_{non-linear, t}}{R^2_{linear, t} + R^2_{non-linear, t}},
$$ 

We can also define an R function to compute the non-linearity index $\eta$ from the matrix $B$:

```{r non-linearity index function}

# this function computes the non-linearity index eta
nonlinearity_index <- function(B, k_lin) {
  
  # check B structure
  if (!is.matrix(B) || nrow(B) != 2) {
    stop("B must be a 2 × k matrix (rows = X, Y).")
  }
  
  # split into linear and nonlinear parts
  B_lin <- B[, 1:k_lin, drop = FALSE]
  B_nl  <- B[, (k_lin + 1):ncol(B), drop = FALSE]
  
  # variance using same logic as confounder_strength()
  get_var <- function(v) {
    as.numeric(v %*% v)
  }
  
  # X variances (row 1)
  var_X_lin <- get_var(B_lin[1, ])
  var_X_nl  <- get_var(B_nl[1, ])
  
  # Y variances (row 2)
  var_Y_lin <- get_var(B_lin[2, ])
  var_Y_nl  <- get_var(B_nl[2, ])
  
  # nonlinearity proportions
  eta_X <- var_X_nl / (var_X_lin + var_X_nl)
  eta_Y <- var_Y_nl / (var_Y_lin + var_Y_nl)
  
  # return mean nonlinearity
  mean(c(eta_X, eta_Y))
}

```

### 1.1.8 Methods used
We will consider four classic models:

* The standard CLPM, where we do not model baseline confounders explicitly.
* The RI-CLPM, where we do not model baseline confounders explictly. 
* The DPM, where we do not model baseline confounders explicitly.
* The standard CLPM with linear adjustment per occasion.

In addition, we will consider models that residualise the variables of interest w.r.t. the confounders per occasion using various machine learning methods. These methods include:

* Linear regression (equivalent to linear adjustment per occasion).
* Lasso Regression.
* Random Forests (Ranger).
* Bayesian Additive Rregression trees (BART).
* Extreme Gradient Boosting (XGBoost).
* Optional: TabPFN.

After residualising, we will fit a standard CLPM to the residualised variables.

### 1.1.9 Conditions table
All in all, we can construct the following conditions table for study 1: 

```{r conditions table 1, echo=FALSE, message=FALSE, warning=FALSE}

library(kableExtra)

df <- data.frame(
  Scenario = paste("Scenario", 1:5),
  `Varying effect` = c("None", "Linear", "Stepwise", "Sinusoidal", "Random walk"),
  `CLPM` = "✓",
  `RI-CLPM` = "✓",
  `DPM` = "✓",
  `CLPM + Linear Adj.` = "✓"
)

kbl(
  df,
  booktabs = TRUE,
  caption = "Conditions table study 1",
  caption_placement = "bottom"
) |>
  kable_styling(full_width = FALSE)

```

Note that here the models are not true conditions, as they do not influence the data generation process. They are rather different methods to analyse the same data. We now move on to the conditions table for study 2, which omits the methods used here for brevity. The methods are applied to each condition in the same way as in study 1, now including the residualisation methods as well. 

```{r conditions table 2, echo=FALSE, message=FALSE, warning=FALSE}

library(kableExtra)

df <- data.frame(
  Scenario = paste("Scenario", 1:5),
  `Varying effect` = c("None", "Linear", "Stepwise", "Sinusoidal", "Random walk"),
  `NL = 0.1` = "✓",
  `NL = 0.25` = "✓",
  `NL = 0.5` = "✓"
)

kbl(
  df,
  booktabs = TRUE,
  caption = "Conditions table study 2: NL abbreviates non-linearity index",
  caption_placement = "bottom"
) |>
  kable_styling(full_width = FALSE)

```

All variables that are not mentioned in the conditions table are held constant across conditions. Namely:

* Cross-lagged effects ($a_{xy}$): 0.1
* Auto-regressive effects ($b_{xy}$): 0.2
* Residual correlations ($\rho$): 0.15
* Sample size ($N$): 1000
* Number of occasions ($T$): 5
* Strength of confounding effects at baseline ($R^2$): 0.25 for both X and Y
* Standard deviation of confounding effects over time: 0.1
* Number of linear confounders: 3
* Number of non-linear components per confounder: 1

## 1.2 Performance metrics
For the evaluation of our methods we will look at the following performance metrics:

* Relative bias: the deviation from the true value relative to the true value.
* Empirical standard error: the standard deviation of the estimates across replications.
* Convergence: the proportion of replications where the model converged successfully.

## 1.3 Sampling beta matrices 
We have defined the number of confounders, their level of confounding and their level of non-linearity. However, we still run to an issue here. We can simply pick coefficients for the confounders that satisfy these conditions, but results would not be very general. Instead, we could also come up with a function that samples random coefficients that satisfy these conditions. Note that because all of our confounders are generated in the same way, we can feed the same beta-matrix into all simulation functions. 

Sampling such that we get a matrix that satisfies these conditions is hard. For example, if we try to sample coefficients, and reject those when they do not meet the conditions, we are sampling for a very long time. With 3 linear and 3 non-linear confounders per variable, we have 12 coefficients to sample. With only 2 conditions, we are basically sampling from a 10-dimensional space, which is very inefficient. A visualisation of this problem can be seen below in a 2 and 3-dimensional space:

```{r visualising thin surfaces, echo=FALSE, message=FALSE, warning=FALSE}

### 2D circle inside square ###
set.seed(1)

# number of random points in the square
n <- 5000

# sample uniformly from [-1,1]^2
theta1 <- runif(n, -1, 1)
theta2 <- runif(n, -1, 1)

# radius of the circle (constraint)
r <- 0.8

# distance from origin
dist2 <- sqrt(theta1^2 + theta2^2)

# points "near" the circle: thin band of width eps
eps <- 0.02
on_circle <- abs(dist2 - r) <= eps

# plot
plot(theta1, theta2,
     col = ifelse(on_circle, "red", "grey80"),
     pch = 16, cex = 0.6,
     xlab = expression(theta[1]),
     ylab = expression(theta[2]),
     main = "Thin surface in 2D: circle inside square")
abline(h = 0, v = 0, col = "grey50", lty = 3)
box()


### 3D sphere inside cube ###
library(scatterplot3d)

set.seed(2)

n <- 8000

# sample uniformly from [-1,1]^3
theta1 <- runif(n, -1, 1)
theta2 <- runif(n, -1, 1)
theta3 <- runif(n, -1, 1)

# sphere radius
r <- 0.9

# distance from origin
dist3 <- sqrt(theta1^2 + theta2^2 + theta3^2)

# narrow band around the sphere
eps <- 0.03
on_sphere <- abs(dist3 - r) <= eps

# 3D scatter: cube in grey, thin shell in red
scatterplot3d(theta1, theta2, theta3,
              color = ifelse(on_sphere, "red", "grey80"),
              pch   = 16,
              xlab  = expression(theta[1]),
              ylab  = expression(theta[2]),
              zlab  = expression(theta[3]),
              main  = "Thin surface in 3D: sphere shell inside cube")

```

One way of solving this issue would be to allow for some tolerance (i.e. thicken the band), but since our problem is rather simple, we can do better. For this we need to first realize that we can calculate the variance contributed by the confounders as the sum of squares of the coefficients:

$$
R^2_X = \sum_{i=1}^{k} \beta_{X,t,i}^2
$$

$$
R^2_Y = \sum_{i=1}^{k} \beta_{Y,t,i}^2,
$$

where $k$ is the number of confounders (both linear and non-linear), and $\beta_{X}$ and $\beta_{Y}$ are the vectors of coefficients for X and Y respectively. The i is just to denote we sum all of the coefficients, up to coefficient k. First we must note that:

$$
R^2_t = R^2_{X,t} + R^2_{Y,t},
$$

Where $R^2,t$ denotes the variance explained by the confounders (as a whole) at occasion $t$. $R^2_{X,t}$ is the variance explained by the confounders on the variable that we have labelled $X$. In order to make this problem identifiable, we can impose an extra constraint. As such we can devide the variance equally between X and Y:

$$ 
\frac{R^2_t}{2} = R^2_{X,t} = R^2_{Y,t}.
$$

However, this need not be the case. You can imagine that we make the share of variance a parameter as well, which we can denote as $\delta$. The general form of the constraint would then be:

$$
R^2_{X,t} = \delta R^2_t
$$

$$
R^2_{Y,t} = (1 - \delta) R^2_t,
$$

It would for example be possible to sample from a distribution to pick a number between 0 and 1 for this parameter $\delta$. This would make the simulation scenarios more general, but for this research rapport, we will stick to the simpler equal variance case. This is mainly done because we are working in a 'doubly robust' scenario. Correct adjustment for the confounders on either X or Y is sufficient to get unbiased estimates of the cross-lagged effects. To show the scenarios where SOTA methods do not work, we should thus first focus on the case where both variables suffer from this scenario equally.

We have a similar choice situation regarding the ammount of non-linearity $\eta$. We can either choose to divide the non-linearity equally between X and Y, or we can choose to divide it unequally. For the research rapport, we will again choose the simpler equal non-linearity case:

$$
\eta_t = \eta_{X,t} = \eta_{Y,t},
$$

where $\eta_{X,t}$ is the non-linearity index for variable X at occasion t, and $\eta_{Y,t}$ is the non-linearity index for variable Y at occasion t. We could again choose to make this a parameter that is sampled from a distribution, which we could call $\gamma$. The general form of this constraint would then be:

$$
\eta_{X,t} = \gamma \eta_t
$$

$$
\eta_{Y,t} = (1 - \gamma)\eta_t,
$$

For this research rapport, we will again stick to the simpler equal non-linearity case. Partly because this makes the sampling procedure easier, and partly because we are in a doubly robust scenario again. Now recall that in section 1.1.7 we defined the non-linearity index as:

$$
\eta_t = \frac{R^2_{non-linear, t}}{R^2_{linear, t} + R^2_{non-linear, t}} = \frac{R^2_{non-linear, t}}{R^2_t}.
$$

We can split this equation up for X and Y:

$$
\eta_X = \frac{R^2_{\text{non-linear}, X}}{R^2_t}
$$

$$
\eta_Y = \frac{R^2_{\text{non-linear}, Y}}{R^2_t},
$$

with our assumption of equal non-linearity, we can rewrite this as:

$$
R^2_{\text{non-linear}, X}
  = \eta_t R^2_X
  = \eta_t \frac{R^2_t}{2}
$$

$$
R^2_{\text{linear}, X}
  = (1 - \eta_t) R^2_X
  = (1 - \eta_t) \frac{R^2_t}{2},
$$

where $R^2_{non-linear, X}$ is the variance explained by the non-linear components of the confounders on X, and $R^2_{linear, X}$ is the variance explained by the linear components of the confounders on X. $\eta_t$ is the non-linearity index at occasion t, and $R^2_{X, t}$ denotes the total variance explained by the confounders on X at occasion t. Finally, $R^2_t$ is the total variance explained by the confounders on both X and Y at occasion t. The point being is that the variance explained by the linear and non-linear parts is detemined by the total variance explained and the non-linearity index, when assuming equal non-linearity and variance between X and Y. We can do the same for Y:

$$
R^2_{\text{non-linear}, Y}
  = \eta_t R^2_Y
  = \eta_t \frac{R^2_t}{2}
$$

$$
R^2_{\text{linear}, Y}
  = (1 - \eta_t) R^2_Y
  = (1 - \eta_t) \frac{R^2_t}{2},
$$

The more general form of these formulas, where we do not assume equal non-linearity and variance between X and Y, would be:

$$
R^2_{\text{non-linear}, X} = \frac{\delta \gamma R^2_t}{2}\, ,
$$

$$
R^2_{\text{linear}, X} = \frac{\delta (1 - \gamma) R^2_t}{2}\, ,
$$

$$
R^2_{\text{non-linear}, Y} = \frac{(1 - \delta)\gamma R^2_t}{2}\, ,
$$

$$
R^2_{\text{linear}, Y} = \frac{(1 - \delta)(1 - \gamma) R^2_t}{2}\, ,
$$

Now is it not a coincidence that we defined the variance explained by the confounders as the sum of squares of the coefficients, obtaining the following equations for these components:

$$
R^2_{\text{non-linear}, X} = \sum_{i=1}^{k_{nl}} \beta_{X,t,i}^2\, ,
$$

$$
R^2_{\text{linear}, X} = \sum_{i=1}^{k_{l}} \beta_{X,t,i}^2\, ,
$$

$$
R^2_{\text{non-linear}, Y} = \sum_{i=1}^{k_{nl}} \beta_{Y,t,i}^2\, ,
$$

$$
R^2_{\text{linear}, Y} = \sum_{i=1}^{k_{l}} \beta_{Y,t,i}^2\, ,
$$

where $k_{nl}$ is the number of non-linear confounder components, and $k_{l}$ is the number of linear confounder components. The general case where we do not assume equal non-linearity and variance are ommited here for brevity. If there would be 3 confounders, with 1 non-linear component each, it follows that each of the explained variances ($R^2$) would be the sum of 3 squared coefficients. Now since the value of the squared coefficients also follows directly from the values picked for $\eta$ and $R^2$, given our constraints, we can split the sampling procedure. We must only sample 3 coefficients, such that their squared sum is equal to the desired explained variance. We can do this by using one final trick. First we sample the coefficients from any distribution, but let's use the example of a standard normal:

$$
\beta_{i} \sim \mathcal{N}(0, 1) \quad \text{for } i = 1, 2, 3
$$

The sum of these squared coefficients may still have any value, but we need to make their sum equal to the desired explained variance. We can do this by scaling the coefficients:

$$
\sum{i=1}^{3} \left( \beta_{i} \cdot \sqrt{\frac{R^2_{\text{component}}}{\sum_{j=1}^{3} \beta_{j}^2}} \right)^2 = R^2_{\text{component}},
$$

by which we obtain the desired sum of squares. To recap, we:

1. Pick a desired explained variance of the confounders on X and Y ($R^2_t$)
2. Pick a desired non-linearity index ($\eta_t$)
3. Calculate the explained variance for the linear and non-linear components on X and Y using the formulas above.
4. Sample 3 coefficients from a standard normal distribution for each component (linear X, non-linear X, linear Y, non-linear Y).
5. Scale the coefficients such that their sum of squares is equal to the desired explained variance.

We can implement this procedure (assuming equal non-linearity and variance) in the following R function. Do note that here we sample for the first occasion. Later occasions may differs slightly as we implement variation over time. Also note that here we chose to reject any coefficients larger than 0.3 or smaller than -0.3. 

```{r sampling beta matrices function}

sample_gamma <- function(
    k_lin,                                             # number of linear confounders
    k_nonlin,                                          # number of nonlinear confounders
    R2_1,                                              # total variance explained by confounders at occasion 1
    eta_1,                                             # nonlinearity index at occasion 1
    min_abs = 0.01,                                    # min and max absolute effect sizes for rejection sampling
    max_abs = 0.30,
    max_tries = 1e6                                    # max number of tries
) {

  k <- k_lin + k_nonlin

  ## allocate variance budgets
  LX <- (1 - eta_1) * R2_1 / 2
  LY <- (1 - eta_1) * R2_1 / 2
  NX <- eta_1 * R2_1 / 2 
  NY <- eta_1 * R2_1 / 2 

  for (try in 1:max_tries) {

    # ---- sample random directions ----
    u_x_lin  <- rnorm(k_lin)
    u_y_lin  <- rnorm(k_lin)
    u_x_nlin <- rnorm(k_nonlin)
    u_y_nlin <- rnorm(k_nonlin)

    # ---- normalize ----
    u_x_lin  <- u_x_lin  / sqrt(sum(u_x_lin^2))
    u_y_lin  <- u_y_lin  / sqrt(sum(u_y_lin^2))
    u_x_nlin <- u_x_nlin / sqrt(sum(u_x_nlin^2))
    u_y_nlin <- u_y_nlin / sqrt(sum(u_y_nlin^2))

    # ---- scale to proper total variance ----
    gamma_x_lin  <- sqrt(LX) * u_x_lin
    gamma_y_lin  <- sqrt(LY) * u_y_lin
    gamma_x_nlin <- sqrt(NX) * u_x_nlin
    gamma_y_nlin <- sqrt(NY) * u_y_nlin

    # ---- assemble full vectors ----
    gamma_x <- c(gamma_x_lin,  gamma_x_nlin)
    gamma_y <- c(gamma_y_lin,  gamma_y_nlin)

    # ---- rejection check (absolute bounds!) ----
    if (
      all(abs(gamma_x) >= min_abs & abs(gamma_x) <= max_abs &
          abs(gamma_y) >= min_abs & abs(gamma_y) <= max_abs)
    ) {
      return(list(
        gamma_x = gamma_x,
        gamma_y = gamma_y
      ))
    }
  }

  stop("Failed to find a valid sample within max_tries.")
}

```

## 1.4 Output plots 
Now it is time to define what we want as the output of our two simulation studies. Some plots need to serve as visual aids to explain in an intuitive manner: What is happening? Other plots are true output plots that will show how our methods perform. From simulation study 1, we want to show the inability of RI-CLPM and DPM to handle time-varying effects of baseline confounders in our 5 varying scenarios. Lets produce the following plots:

### 1.4.1 Simulation study 1

* Plot 1: On the X-axis we have the occasion, and on the Y-axis the effect of the confounder on X. A different plot of the same kind will serve to show the confounder effect on Y, yielding 2 plots in total for each scenario. Since we will sample beta matrixes, each beta will have a different starting point, but all will make 'the same jump'. Each line in the plot thus represents a different run of the simulation. We will visualise the effect of only one confounder as it serves purely for understanding. Such a plot would look like:

```{r visualising time-varying effects, echo=FALSE, message=FALSE, warning=FALSE}

# libs
library(tidyverse)
library(patchwork)

# reproducibility
set.seed(42)

# set the parameters
n_runs     <- 500                                       # number of simulated lines per scenario
T_occasions <- 5                                        # number of time points
scenarios <- c("Constant", "Linear", "Sinusoidal", 
              "Stepwise", "Random Walk")

# function to generate the path per scenario
generate_path <- function(beta0, scenario, T_occasions) {
  
  # for occasion t up to T
  t <- 1:T_occasions
  
  # adjustments per scenario
  path <- switch(
    scenario,
    
    # repeat the starting beta
    "Constant" = rep(beta0, T_occasions),
    
    # add constant
    "Linear" = beta0 + seq(0, 0.4, length.out = T_occasions),  
    
    # vary sinusoidally
    "Sinusoidal" = beta0 + sin(seq(0, 2*pi, length.out = T_occasions)) * 0.2,
    
    # make a stepwise jump at t = 3
    "Stepwise" = {
      step_jump <- 0.4
      out <- rep(beta0, T_occasions)
      out[3:T_occasions] <- out[3:T_occasions] + step_jump  
      out
    },
    
    # random variation , but in one direction 
    "Random Walk" = {
      rw <- cumsum(c(beta0, rnorm(T_occasions - 1, mean = 0, sd = 0.05)))
      rw
    }
  )
  
  # save in tibble
  tibble(
    occasion = t,
    beta = path
  )
}

# generate data for all scenarios and runs 
demo_data <- map_df(

  # imput scenarios
  scenarios,

  # define function per scenario
  function(scen) {

    # apply over number of runs
    map_df(
      1:n_runs,
      function(run_id) {

        # random starting point
        beta0 <- rnorm(1, 0.1, 0.05)  

        # use the generation function
        generate_path(beta0, scen, T_occasions) %>%
          mutate(
            run = run_id,
            scenario = scen
          )
      }
    )
  }
)

# save data for X and Y effects
demo_data_xy <- demo_data %>%
  mutate(
    beta_X = beta,
    beta_Y = beta   
  )

# create a blue → magenta palette for the line colours
blue_magenta <- c(
  "#3B4CC0",  
  "#5E60CE",
  "#7A70D1",
  "#9775D3",
  "#B67BD2",
  "#D084CF",
  "#E28FC4",
  "#F29BC1",
  "#FF9ACD"
)

# repeat the palette so the number of colours matches the number of runs
n_runs <- length(unique(demo_data_xy$run))
blue_magenta_extended <- rep(blue_magenta, length.out = n_runs)

# plot the effects on X
plot_X <- ggplot(demo_data_xy,
                 aes(x = occasion, y = beta_X,
                     group = run, color = factor(run))) +
  geom_line(linewidth = 1) +
  facet_wrap(~ scenario) +
  scale_color_manual(values = blue_magenta_extended) +
  labs(
    x = "Occasion",
    y = "Effect on X"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 10)
  )

plot_X

```

* Plot 2: On the X-axis we again have the occasion, and on the Y-axis the explained variance ($R^2$) of the confounder on X. Each line in the plot thus represents a different run of the simulation. We will visualise the effect of only one confounder as it serves purely for understanding. Such a plot would look like:

```{r visualising time-varying R2, echo=FALSE, message=FALSE, warning=FALSE}

# compute explained variance R² for X 
demo_data_r2 <- demo_data_xy %>%
  mutate(
    R2_X = beta_X^2,
    R2_Y = beta_Y^2
  )

plot_R2_X <- ggplot(demo_data_r2,
                    aes(x = occasion, y = R2_X,
                        group = run, color = factor(run))) +
  geom_line(linewidth = 1) +
  facet_wrap(~ scenario) +
  scale_color_manual(values = blue_magenta_extended) +
  labs(
    x = "Occasion",
    y = expression(R^2 ~ "of confounder on X")
  ) +
  theme_minimal(base_size = 10) +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 10)
  )

plot_R2_X

```

* Plot 3: The first of the true results plots. On the X-axis we again have the occasion, and on the Y-axis we plot the relative bias. Now do note that relative bias is the (|estimate - true value|) / true value. Taking the absolute value here is important, since otherwise, positive and negative bias cancel out. We need to make two plots: bias for the autoregressive parameter and bias for the cross-lagged parammeter. For each scenario we will thus obtain: 2 (each parameter) * 5 (scenarios) * 5 (methods) = 50 plots in total. We will make a grid where we have the 5 scenarios, and then we get the plot for each method. We will have 2 of those plots, one for each parameter. Note that we would get twice the grid size if we were to include these for Y too. Here we will show just one example plot: 

```{r visualising relative bias example, echo=FALSE, message=FALSE, warning=FALSE}

# reproducibility
set.seed(123)

# methods
methods <- c("RI_CLPM", "DPM", "CLPM", "CLPM_LBCA", "CLPM_Adj")

# add true values
demo_bias <- demo_data_xy %>%
  mutate(
    true_cross = 0.10,
    true_auto  = 0.25
  )

# fake estimates
demo_bias <- demo_bias %>%
  mutate(
    est_RI_CLPM   = true_cross + rnorm(n(), 0, 0.03),
    est_DPM       = true_cross + rnorm(n(), 0, 0.03),
    est_CLPM      = true_cross + rnorm(n(), 0, 0.03),
    est_CLPM_LBCA = true_cross + rnorm(n(), 0, 0.03),
    est_CLPM_Adj  = true_cross + rnorm(n(), 0, 0.03),

    estA_RI_CLPM   = true_auto + rnorm(n(), 0, 0.04),
    estA_DPM       = true_auto + rnorm(n(), 0, 0.04),
    estA_CLPM      = true_auto + rnorm(n(), 0, 0.04),
    estA_CLPM_LBCA = true_auto + rnorm(n(), 0, 0.04),
    estA_CLPM_Adj  = true_auto + rnorm(n(), 0, 0.04)
  )

# relative bias
demo_bias <- demo_bias %>%
  mutate(
    rb_RI_CLPM   = abs(est_RI_CLPM   - true_cross) / true_cross,
    rb_DPM       = abs(est_DPM       - true_cross) / true_cross,
    rb_CLPM      = abs(est_CLPM      - true_cross) / true_cross,
    rb_CLPM_LBCA = abs(est_CLPM_LBCA - true_cross) / true_cross,
    rb_CLPM_Adj  = abs(est_CLPM_Adj  - true_cross) / true_cross
  )

# mean bias per occasion × scenario × method
demo_plotdata <- demo_bias %>%
  select(scenario, occasion, run, starts_with("rb_")) %>%
  pivot_longer(
    cols = starts_with("rb_"),
    names_to = "method",
    values_to = "rel_bias"
  ) %>%
  mutate(
    method = gsub("rb_", "", method),
    method = factor(method, levels = methods)
  ) %>%
  group_by(scenario, method, occasion) %>%
  summarise(mean_bias = mean(rel_bias), .groups = "drop")

# colours
method_colors <- c(
  "RI_CLPM"   = "#3B4CC0",
  "DPM"       = "#5E60CE",
  "CLPM"      = "#7A70D1",
  "CLPM_LBCA" = "#B67BD2",
  "CLPM_Adj"  = "#EE3A8C"
)

# plot
plot_bias_cross <- ggplot(demo_plotdata,
                          aes(x = occasion, y = mean_bias,
                              color = method, group = method)) +
  geom_point(size = 2) +
  geom_line(linewidth = 0.8) +
  facet_grid(method ~ scenario) +
  scale_color_manual(values = method_colors) +
  labs(
    x = "Occasion",
    y = "Mean Relative Bias (Cross-Lagged)"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 10)
  )

plot_bias_cross

```

* Plot 4: produce the same plot, but for the emprical standard error instead of the relative bias.

```{r visualising empirical SE example, echo=FALSE, message=FALSE, warning=FALSE}

# empirical standard error per occasion × scenario × method
demo_se <- demo_bias %>%
  select(scenario, occasion, run, starts_with("rb_")) %>%
  pivot_longer(
    cols = starts_with("rb_"),
    names_to = "method",
    values_to = "rel_bias"
  ) %>%
  mutate(
    method = gsub("rb_", "", method),
    method = factor(method, levels = methods)
  ) %>%
  group_by(scenario, method, occasion) %>%
  summarise(se_bias = sd(rel_bias), .groups = "drop")

# colours
method_colors <- c(
  "RI_CLPM"   = "#3B4CC0",
  "DPM"       = "#5E60CE",
  "CLPM"      = "#7A70D1",
  "CLPM_LBCA" = "#B67BD2",
  "CLPM_Adj"  = "#EE3A8C"
)

# plot empirical SE
plot_se_cross <- ggplot(demo_se,
                        aes(x = occasion, y = se_bias,
                            color = method, group = method)) +
  geom_point(size = 2) +
  geom_line(linewidth = 0.8) +
  facet_grid(method ~ scenario) +
  scale_color_manual(values = method_colors) +
  labs(
    x = "Occasion",
    y = "Empirical SE (Cross-Lagged)"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 10)
  )

plot_se_cross

```

* plot 5: Even though the previous plots are very detailed, which is great, it would be nice to have one summary plot. It should simply show a boxplot of the mean absolute estimate with each S.e. . As such we should get 5 boxplots for each method, one per scenario. Do note that this doubles once we want to include a plot for the autoregressive parameter too. 

```{r visualising estimates summary, echo=FALSE, message=FALSE, warning=FALSE}

# true cross-lagged value
true_beta <- 0.10

# build correct per-scenario summaries
summary_est <- demo_bias %>%
  select(scenario, occasion, run, starts_with("est_")) %>%
  pivot_longer(
    cols = starts_with("est_"),
    names_to = "method",
    values_to = "estimate"
  ) %>%
  mutate(
    method = gsub("est_", "", method),
    method = factor(method, levels = methods)
  ) %>%
  group_by(scenario, method) %>%
  summarise(
    mean_est = mean(estimate),   # raw estimate mean per scenario × method
    sd_est   = sd(estimate),     # empirical SD
    .groups = "drop"
  ) %>%
  mutate(true_val = true_beta)   # <- explicitly add truth

method_colors <- c(
  "RI_CLPM"   = "#3B4CC0",
  "DPM"       = "#5E60CE",
  "CLPM"      = "#7A70D1",
  "CLPM_LBCA" = "#B67BD2",
  "CLPM_Adj"  = "#EE3A8C"
)

plot_estimates_scen <- ggplot(summary_est,
                              aes(x = method, y = mean_est, color = method)) +

  geom_errorbar(aes(ymin = mean_est - sd_est,
                    ymax = mean_est + sd_est),
                width = 0.15, linewidth = 0.8, alpha = 0.85) +

  geom_point(size = 2) +

  geom_hline(aes(yintercept = true_val),
             linetype = "dashed", linewidth = 0.9) +

  facet_grid(scenario ~ .) +   # <- one block per scenario

  scale_color_manual(values = method_colors) +

  labs(
    x = "Method",
    y = "Estimated Cross-Lagged Effect"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 11, face = "bold"),
    axis.text.x = element_text(angle = 25, hjust = 1)
  )

plot_estimates_scen

```

This means all can be created from a dataframe which has the following structure:

```{r final demo dataframe structure, echo=FALSE, message=FALSE, warning=FALSE}

library(kableExtra)

# show first rows of the final simulation dataset
demo_bias %>%
  head(10) %>%      # just the header
  kbl(booktabs = TRUE,
      caption = "Structure of the Final Simulation Output (First 10 Rows)") %>%
  kable_styling(full_width = FALSE)

```

Where in our final simulations, beta will not be one number, but a vector of coefficients for all confounders. Additionally, we will need to save the convergence information per method as well, which can simply be appended using an extra column per method with TRUE/FALSE values. 

### 1.4.2 Simulation study 2

* Plot 1: This plot looks exactly the same as for simulation study 1. 
* Plot 2: This plot is the same, except for one important caveat: we want to split the variance by non-linear and linear variance. We want to visualise this clearly. What we want to do is to 'add' the non-linear variance 'on top' of the linear variance. Note that for this plot to work we need both the linear and the non-linear coefficients. 

```{r visualising time-varying R2 non-linear addition, echo=FALSE, message=FALSE, warning=FALSE}

# assume 40% of explained variance is nonlinear
eta <- 0.40

# compute linear and nonlinear R² components
demo_data_r2 <- demo_data_xy %>%
  mutate(
    R2_total = beta_X^2,
    R2_nl    = eta * R2_total,
    R2_lin   = (1 - eta) * R2_total
  )

# long transform with correct stacking order
demo_data_r2_long <- demo_data_r2 %>%
  select(scenario, occasion, run, R2_lin, R2_nl) %>%
  pivot_longer(
    cols = c(R2_lin, R2_nl),
    names_to = "component",
    values_to = "R2"
  ) %>%
  mutate(
    component = factor(component, levels = c("R2_lin", "R2_nl"))
  )

# stacked area plot, no legend
plot_R2_stacked <- ggplot(demo_data_r2_long,
                          aes(x = occasion, y = R2,
                              fill = component,
                              group = interaction(run, component))) +
  geom_area(alpha = 0.65, position = "stack") +
  facet_wrap(~ scenario) +
  scale_fill_manual(
    values = c(
      R2_lin = "#7A70D1",   # bottom layer
      R2_nl  = "#EE3A8C"    # top layer
    )
  ) +
  labs(
    x = "Occasion",
    y = expression(R^2 ~ "Explained variance")
  ) +
  theme_minimal(base_size = 10) +
  theme(
    strip.text = element_text(size = 10),
    legend.position = "none"     # <— yeeted legend
  )

plot_R2_stacked

```

* Plot 3: Same as simulation study 1, but we now add some methods. 

```{r visualising relative bias example sim2, echo=FALSE, message=FALSE, warning=FALSE}

# reproducibility
set.seed(123)

# 9 methods for Simulation Study 2
methods <- c(
  "CLPM",
  "RI_CLPM",
  "DPM",
  "CLPM_LinAdj",
  "CLPM_BCA",
  "LASSO",
  "RF",
  "BART",
  "XGBoost"
)

# add true values
demo_bias <- demo_data_xy %>%
  mutate(
    true_cross = 0.10,
    true_auto  = 0.25
  )

# fake estimates for demonstration — replace later with real simulation outputs
demo_bias <- demo_bias %>%
  mutate(
    est_CLPM        = true_cross + rnorm(n(), 0, 0.04),
    est_RI_CLPM     = true_cross + rnorm(n(), 0, 0.04),
    est_DPM         = true_cross + rnorm(n(), 0, 0.04),
    est_CLPM_LinAdj = true_cross + rnorm(n(), 0, 0.04),
    est_CLPM_BCA    = true_cross + rnorm(n(), 0, 0.04),
    est_LASSO       = true_cross + rnorm(n(), 0, 0.04),
    est_RF          = true_cross + rnorm(n(), 0, 0.04),
    est_BART        = true_cross + rnorm(n(), 0, 0.04),
    est_XGBoost     = true_cross + rnorm(n(), 0, 0.04)
  )

# compute relative bias
demo_bias <- demo_bias %>%
  mutate(
    rb_CLPM        = abs(est_CLPM        - true_cross) / true_cross,
    rb_RI_CLPM     = abs(est_RI_CLPM     - true_cross) / true_cross,
    rb_DPM         = abs(est_DPM         - true_cross) / true_cross,
    rb_CLPM_LinAdj = abs(est_CLPM_LinAdj - true_cross) / true_cross,
    rb_CLPM_BCA    = abs(est_CLPM_BCA    - true_cross) / true_cross,
    rb_LASSO       = abs(est_LASSO       - true_cross) / true_cross,
    rb_RF          = abs(est_RF          - true_cross) / true_cross,
    rb_BART        = abs(est_BART        - true_cross) / true_cross,
    rb_XGBoost     = abs(est_XGBoost     - true_cross) / true_cross
  )

# reshape → mean RB per occasion × scenario × method
demo_plotdata <- demo_bias %>%
  select(scenario, occasion, run, starts_with("rb_")) %>%
  pivot_longer(
    cols = starts_with("rb_"),
    names_to = "method",
    values_to = "rel_bias"
  ) %>%
  mutate(
    method = gsub("rb_", "", method),
    method = factor(method, levels = methods)
  ) %>%
  group_by(scenario, method, occasion) %>%
  summarise(mean_bias = mean(rel_bias), .groups = "drop")

# method colors (9 distinct)
method_colors <- c(
  "CLPM"        = "#3B4CC0",
  "RI_CLPM"     = "#5E60CE",
  "DPM"         = "#7A70D1",
  "CLPM_LinAdj" = "#9775D3",
  "CLPM_BCA"    = "#B67BD2",
  "LASSO"       = "#D084CF",
  "RF"          = "#E28FC4",
  "BART"        = "#F29BC1",
  "XGBoost"     = "#EE3A8C"
)

# final plot: 9 methods × 5 scenarios grid
plot_bias_cross <- ggplot(demo_plotdata,
                          aes(x = occasion, y = mean_bias,
                              color = method, group = method)) +
  geom_point(size = 1.8) +
  geom_line(linewidth = 0.9) +
  facet_grid(method ~ scenario) +
  scale_color_manual(values = method_colors) +
  labs(
    x = "Occasion",
    y = "Mean Relative Bias (Cross-Lagged)"
  ) +
  theme_minimal(base_size = 9) +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 9)
  )

plot_bias_cross

```

* Plot 4: Same as simulation study 1, but with the added methods.

```{r visualising empirical SE example sim2, echo=FALSE, message=FALSE, warning=FALSE}

# empirical SD per occasion × scenario × method
demo_plotdata_se <- demo_bias %>%
  select(scenario, occasion, run, starts_with("est_")) %>%
  pivot_longer(
    cols = starts_with("est_"),
    names_to = "method",
    values_to = "estimate"
  ) %>%
  mutate(
    method = gsub("est_", "", method),
    method = factor(method, levels = methods)
  ) %>%
  group_by(scenario, method, occasion) %>%
  summarise(se_est = sd(estimate), .groups = "drop")

# same 9-method colors
method_colors <- c(
  "CLPM"        = "#3B4CC0",
  "RI_CLPM"     = "#5E60CE",
  "DPM"         = "#7A70D1",
  "CLPM_LinAdj" = "#9775D3",
  "CLPM_BCA"    = "#B67BD2",
  "LASSO"       = "#D084CF",
  "RF"          = "#E28FC4",
  "BART"        = "#F29BC1",
  "XGBoost"     = "#EE3A8C"
)

# empirical SD plot
plot_se_cross <- ggplot(demo_plotdata_se,
                        aes(x = occasion, y = se_est,
                            color = method, group = method)) +
  geom_point(size = 1.8) +
  geom_line(linewidth = 0.9) +
  facet_grid(method ~ scenario) +
  scale_color_manual(values = method_colors) +
  labs(
    x = "Occasion",
    y = "Empirical SD (Cross-Lagged)"
  ) +
  theme_minimal(base_size = 9) +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 9)
  )

plot_se_cross

```

* Plot 5: Same as simulation study 1, but with the added methods.

```{r visualising estimates summary sim2, echo=FALSE, message=FALSE, warning=FALSE}

# true cross-lagged value
true_beta <- 0.10

# build summary dataset (mean estimate + empirical SD)
summary_est <- demo_bias %>%
  select(scenario, occasion, run, starts_with("est_")) %>%
  pivot_longer(
    cols = starts_with("est_"),
    names_to = "method",
    values_to = "estimate"
  ) %>%
  mutate(
    method = gsub("est_", "", method),
    method = factor(method, levels = methods)   # 9-method vector
  ) %>%
  group_by(scenario, method) %>%
  summarise(
    mean_est = mean(estimate),   # average estimate across runs × occasions
    sd_est   = sd(estimate),     # empirical SD
    .groups = "drop"
  ) %>%
  mutate(true_val = true_beta)    # add truth line

# 9-method palette
method_colors <- c(
  "CLPM"        = "#3B4CC0",
  "RI_CLPM"     = "#5E60CE",
  "DPM"         = "#7A70D1",
  "CLPM_LinAdj" = "#9775D3",
  "CLPM_BCA"    = "#B67BD2",
  "LASSO"       = "#D084CF",
  "RF"          = "#E28FC4",
  "BART"        = "#F29BC1",
  "XGBoost"     = "#EE3A8C"
)

# final plot
plot_estimates_scen <- ggplot(summary_est,
                              aes(x = method, y = mean_est, color = method)) +

  geom_errorbar(aes(ymin = mean_est - sd_est,
                    ymax = mean_est + sd_est),
                width = 0.15, linewidth = 0.8, alpha = 0.85) +

  geom_point(size = 2) +

  geom_hline(aes(yintercept = true_val),
             linetype = "dashed", linewidth = 0.9) +

  facet_grid(scenario ~ .) +

  scale_color_manual(values = method_colors) +

  labs(
    x = "Method",
    y = "Estimated Cross-Lagged Effect"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 11, face = "bold"),
    axis.text.x = element_text(angle = 25, hjust = 1)
  )

plot_estimates_scen

```

Finally, the dataframe structure:

```{r final demo dataframe structure sim2, echo=FALSE, message=FALSE, warning=FALSE}

library(kableExtra)

# Build a template that includes linear + nonlinear pieces explicitly
simulation_output_template <- demo_bias %>%
  mutate(
    # Decompose beta_X and beta_Y into linear + nonlinear components.
    # For demonstration we assume eta = .40 as earlier.
    eta = 0.40,
    beta_X_lin = sqrt(1 - eta) * beta_X,
    beta_X_nl  = sqrt(eta)     * beta_X,
    beta_Y_lin = sqrt(1 - eta) * beta_Y,
    beta_Y_nl  = sqrt(eta)     * beta_Y
  ) %>%
  select(
    # identifiers
    scenario,
    occasion,
    run,

    # linear / nonlinear decomposition
    beta_X_lin,
    beta_X_nl,
    beta_Y_lin,
    beta_Y_nl,

    # total effect
    beta_X,
    beta_Y,

    # true values
    true_cross,
    true_auto,

    # all estimates (cross-lag)
    starts_with("est_"),

    # all estimates (autoregressive)
    starts_with("estA_")
  ) %>%
  head(10)

simulation_output_template %>%
  kbl(
    booktabs = TRUE,
    caption = "Template of Final Simulation Output Including Linear/Nonlinear Confounder Effects"
  ) %>%
  kable_styling(full_width = FALSE)

```

# 2. Simulation study 1. 
So now we will move to the first simulation study, where we want to show the inability of RI-CLPM and DPM to handle time-varying effects of baseline confounders. We will build the functions needed for this simulation study block by block, starting with the function to sample the B-matrix. 

## 2.1 Sample B-matrix function

```{r sample B linear function, eval=false, message=FALSE, warning=FALSE}

#| label: sample-B-linear
#| eval: true
#| message: false
#| warning: false

# ---------------------------------------------------------
# Function: sample_B_linear()
# ---------------------------------------------------------
# Purpose:
#   Sample a 2 × k matrix of *linear* confounder effects such that:
#       sum(b_x^2) = R2_1 / 2
#       sum(b_y^2) = R2_1 / 2
#   with random orientation but fixed magnitude.
#
# Arguments:
#   k        = number of linear confounders
#   R2_1     = total R² of confounders at t = 1
#   min_abs  = min allowed abs(beta)
#   max_abs  = max allowed abs(beta)
#   max_tries = maximum rejection attempts
#
# Returns:
#   B1 : 2 × k matrix (rows = X, Y)
# ---------------------------------------------------------

sample_B_linear <- function(
    k,
    R2_1,
    min_abs   = 0.01,
    max_abs   = 0.30,
    max_tries = 100000
) {

  # required variance explained per variable
  target_X <- R2_1 / 2
  target_Y <- R2_1 / 2

  for (i in seq_len(max_tries)) {

    # --- random direction for X row ---
    u_x <- rnorm(k)
    u_x <- u_x / sqrt(sum(u_x^2))
    b_x <- sqrt(target_X) * u_x

    # --- random direction for Y row ---
    u_y <- rnorm(k)
    u_y <- u_y / sqrt(sum(u_y^2))
    b_y <- sqrt(target_Y) * u_y

    # --- rejection condition ---
    if (all(abs(b_x) >= min_abs,
            abs(b_x) <= max_abs,
            abs(b_y) >= min_abs,
            abs(b_y) <= max_abs)) {

      B1 <- rbind(b_x, b_y)
      rownames(B1) <- c("X", "Y")
      colnames(B1) <- paste0("c", 1:k)
      return(B1)
    }
  }

  stop("Failed to sample a valid B matrix within max_tries.")
}

```

After we have created the beta-matrices for the first wave, we must generate their trajectory over the next waves. We will create a function that takes the initial beta-matrix and a scenario type, and returns the full list of beta-matrices for all occasions. It is of importance here that we will vary the confounder the same ammount in each scenario! 

## 2.2 Generate B trajectory function

```{r generate B trajectory function, eval=false, message=FALSE, warning=FALSE}

#| label: generate-B-trajectory
#| eval: true
#| message: false
#| warning: false

# ---------------------------------------------------------
# Function: generate_B_trajectory()
# ---------------------------------------------------------
# Purpose:
#   Starting from a baseline B1 (2 × k), generate a list of
#   time-varying B matrices B_t for t = 1...T according to:
#     - "constant"
#     - "linear"
#     - "sinusoidal"
#     - "stepwise"
#     - "random_walk"
#
#   Crucially: the *variance across occasions* of the beta
#   adjustments is equal across all scenarios.
#
# Arguments:
#   B1           = baseline B matrix (2 × k)
#   T            = number of occasions
#   scenario     = one of:
#                 "constant", "linear", "sinusoidal",
#                 "stepwise", "random_walk"
#   target_sd    = desired standard deviation of betas over time
#   rw_sd        = step size for random walk scenario
#
# Returns:
#   List of B_t matrices for t = 1...T
# ---------------------------------------------------------

generate_B_trajectory <- function(
    B1,
    T,
    scenario = c("constant","linear","sinusoidal","stepwise","random_walk"),
    target_sd = 0.10,   # sd of beta variation over occasions
    rw_sd     = 0.05
){

  scenario <- match.arg(scenario)
  k <- ncol(B1)

  # ---- Step 1: create unscaled variation pattern v_t ----
  tvec <- 1:T

  if (scenario == "constant") {
    v <- rep(0, T)

  } else if (scenario == "linear") {
    v <- seq(0, 1, length.out = T)

  } else if (scenario == "sinusoidal") {
    v <- sin(seq(0, 2*pi, length.out = T))

  } else if (scenario == "stepwise") {
    v <- c(rep(0, floor(T/2)), rep(1, T - floor(T/2)))

  } else if (scenario == "random_walk") {
    # zero start ensures comparability across reps
    steps <- rnorm(T - 1, mean = 0, sd = rw_sd)
    v <- c(0, cumsum(steps))
  }

  # ---- Step 2: scale v_t so Var(v_t) = target_sd^2 ----
  v_centered <- v - mean(v)
  current_sd <- sd(v_centered)

  if (current_sd > 0) {
    scaling_factor <- target_sd / current_sd
    v_scaled <- v_centered * scaling_factor
  } else {
    v_scaled <- rep(0, T)
  }

  # ---- Step 3: apply v_scaled to each beta coefficient ----
  B_list <- vector("list", T)
  names(B_list) <- paste0("t", 1:T)

  for (t in 1:T) {
    B_list[[t]] <- B1 + v_scaled[t]
  }

  return(B_list)
}

```

## 2.3 Simulate panel data function
Now that we have our beta-matrix for every occasion, we have all we need to simulate some juicy data. For these data, we do not take non-linear confounder effects into account, as this is reserved for simulation study 2. We will create a function that simulates data for one run of the simulation, given the B-matrices, autoregressive and cross-lagged parameters, number of individuals, and residual variances.

```{r simulate panel data function, eval=false, message=FALSE, warning=FALSE}

library(mvtnorm)

simulate_panel_data <- function(
    N,
    T,
    B_list,
    ar   = 0.25,
    cross = 0.10,
    rho   = 0.30
){
  # -------------------------------------------------------
  # 1. Dynamics matrix
  # -------------------------------------------------------
  A <- matrix(c(ar, cross,
                cross, ar), 2, 2, byrow = TRUE)

  k <- ncol(B_list[[1]])

  # -------------------------------------------------------
  # 2. Confounders
  # -------------------------------------------------------
  U <- mvtnorm::rmvnorm(N, sigma = diag(k))

  # output matrix
  df <- matrix(NA, N, 2*T + k)
  colnames(df) <- c(paste0("x",1:T), paste0("y",1:T), paste0("c",1:k))
  df[, (2*T+1):(2*T+k)] <- U

  # -------------------------------------------------------
  # Function to make innovation covariance Sigma_e
  # Ensures Var(X_t) = 1 and Var(Y_t) = 1
  # -------------------------------------------------------
  make_Sigma_e <- function(Bt, S_dyn, rho){
    # confounder contribution
    S_U <- Bt %*% t(Bt)

    # variance that must be supplied by the innovation errors
    d <- 1 - diag(S_dyn + S_U)

    # numerical floor
    d[d < 1e-12] <- 1e-12

    # correlation structure
    R <- matrix(c(1, rho,
                  rho, 1), 2, 2)

    # scale to desired variance
    D <- diag(sqrt(d))
    Sigma_e <- D %*% R %*% D
    Sigma_e
  }

  # -------------------------------------------------------
  # 3. Wave 1 (NO AR CONTRIBUTION)
  # -------------------------------------------------------
  S_dyn1 <- matrix(0, 2, 2)   # AR contributes nothing at t=1
  Sigma_e1 <- make_Sigma_e(B_list[[1]], S_dyn1, rho)

  Ddyn <- mvtnorm::rmvnorm(N, sigma = Sigma_e1)
  obs1 <- Ddyn + U %*% t(B_list[[1]])

  df[,1]   <- obs1[,1]
  df[,1+T] <- obs1[,2]

  # latent variance for next step
  S_prev <- Sigma_e1

  # -------------------------------------------------------
  # 4. Waves 2...T
  # -------------------------------------------------------
  for(t in 2:T){

    # AR contribution at this wave
    S_dyn_t <- A %*% S_prev %*% t(A)

    # innovation covariance required to keep Var = 1
    Sigma_et <- make_Sigma_e(B_list[[t]], S_dyn_t, rho)

    # update latent state
    Ddyn <- Ddyn %*% t(A) + mvtnorm::rmvnorm(N, sigma = Sigma_et)

    obs <- Ddyn + U %*% t(B_list[[t]])
    df[, t]    <- obs[,1]
    df[, t+T ] <- obs[,2]

    # update variance tracker
    S_prev <- S_dyn_t + Sigma_et
    S_prev <- (S_prev + t(S_prev))/2   # symmetrize
  }

  df
}

# example workflow
set.seed(123)

B1 <- sample_B_linear(
  k    = 3,
  R2_1 = 0.20   # total confounder R² at wave 1
)

B_list <- generate_B_trajectory(
  B1 = B1,
  T  = 5,
  scenario  = "sinusoidal",
  target_sd = 0.10
)

df <- simulate_panel_data(
  N = 10000000,
  T = 5,
  B_list = B_list,
  ar = 0.25,
  cross = 0.10,
  rho = 0.15
)

# round(apply(df[,1:(2*5)], 2, var), 3)

```

## 2.4 Build model functions


The next step is to build the models. We need functions to build those since the number of occasions and confounders is dynamic. The first model to build is the CLPM without any adjustment whatsoever. 

### 2.4.1 CLPM unadjusted

```{r build CLPM function, eval=false, message=FALSE, warning=FALSE}

# ---------------------------------------------------------
# Function: build_clpm()
# ---------------------------------------------------------
# Purpose:
#   Create a basic CLPM syntax string for lavaan.
#   Each occasion t is predicted by x(t-1) and y(t-1).
#   Residuals are allowed to covary at each time point.
#
# Arguments:
#   T  = number of occasions
#
# Returns:
#   A lavaan model string for the CLPM
# ---------------------------------------------------------

build_clpm <- function(T) {

  # regression paths over time
  regress_block <- paste(
    unlist(lapply(2:T, function(t){
      c(
        sprintf("x%d ~ x%d + y%d", t, t-1, t-1),
        sprintf("y%d ~ x%d + y%d", t, t-1, t-1)
      )
    })), collapse="\n"
  )

  # allow residual covariance x_t ~~ y_t
  resid_cov <- paste(sprintf("x%d ~~ y%d", 1:T, 1:T), collapse="\n")

  # allow variances x_t ~~ x_t and y_t ~~ y_t
  resid_vars <- paste(
    paste(sprintf("x%d ~~ x%d", 1:T, 1:T), collapse="\n"),
    paste(sprintf("y%d ~~ y%d", 1:T, 1:T), collapse="\n"),
    sep = "\n"
  )

  # freely estimate means
  means_block <- paste(
    paste(paste0("x",1:T), collapse=" + "), "~ 1\n",
    paste(paste0("y",1:T), collapse=" + "), "~ 1\n"
  )

  paste(regress_block, resid_cov, resid_vars, means_block, sep="\n")
}

```

##### 2.4.2 CLPM with confounder adjustment

Next we include the models with confounder adjustment built into lavaan:

```{r build CLPM with confounders function, eval=false, message=FALSE, warning=FALSE}

# ---------------------------------------------------------
# Function: build_clpm_with_Cs()
# ---------------------------------------------------------
# Purpose:
#   Create a CLPM that includes all confounders (c1...ck)
#   as additional predictors of every x_t and y_t.
#
# Arguments:
#   T = number of occasions
#   k = number of confounders
#
# Returns:
#   A lavaan model string for the adjusted CLPM
# ---------------------------------------------------------

build_clpm_with_Cs <- function(T, k) {

  # construct "c1 + c2 + ... + ck"
  C_names <- paste0("c", 1:k, collapse=" + ")

  # autoregressive and cross-lagged paths + confounders
  regress_block <- paste(
    unlist(lapply(2:T, function(t){
      c(
        sprintf("x%d ~ x%d + y%d + %s", t, t-1, t-1, C_names),
        sprintf("y%d ~ x%d + y%d + %s", t, t-1, t-1, C_names)
      )
    })), collapse="\n"
  )

  # residual covariance x_t ~~ y_t
  resid_cov <- paste(sprintf("x%d ~~ y%d", 1:T, 1:T), collapse="\n")

  # residual variances
  resid_vars <- paste(
    paste(sprintf("x%d ~~ x%d", 1:T, 1:T), collapse="\n"),
    paste(sprintf("y%d ~~ y%d", 1:T, 1:T), collapse="\n"),
    sep = "\n"
  )

  # means
  means_block <- paste(
    paste(paste0("x",1:T), collapse=" + "), "~ 1\n",
    paste(paste0("y",1:T), collapse=" + "), "~ 1\n"
  )

  paste(regress_block, resid_cov, resid_vars, means_block, sep="\n")
}

```

### 2.4.3 RI-CLPM 

Now onto the RI-CLPM:

```{r build RI-CLPM function, eval=false, message=FALSE, warning=FALSE}

# ---------------------------------------------------------
# Function: build_riclpm()
# ---------------------------------------------------------
# Purpose:
#   Create a RI-CLPM syntax string.
#   The model separates stable person-level traits (ri)
#   from within-person fluctuations (wx, wy).
#
# Arguments:
#   T = number of occasions
#
# Returns:
#   A lavaan model string for the RI-CLPM
# ---------------------------------------------------------

build_riclpm <- function(T) {

  # random intercept factors
  ri_block <- paste0(
    "rix =~ ", paste(sprintf("1*x%d", 1:T), collapse=" + "), "\n",
    "riy =~ ", paste(sprintf("1*y%d", 1:T), collapse=" + "), "\n",
    "rix ~~ rix\n riy ~~ riy\n rix ~~ riy\n"
  )

  # fix residual variances to zero at the observed level
  resid_fix <- paste0(
    paste(sprintf("x%d ~~ 0*x%d", 1:T, 1:T), collapse="; "), "\n",
    paste(sprintf("y%d ~~ 0*y%d", 1:T, 1:T), collapse="; "), "\n"
  )

  # within-person latent variables
  within_lat <- paste0(
    paste(sprintf("wx%d =~ 1*x%d", 1:T, 1:T), collapse="; "), "\n",
    paste(sprintf("wy%d =~ 1*y%d", 1:T, 1:T), collapse="; "), "\n"
  )

  # orthogonality: RI factors uncorrelated with within-person factors
  orth <- paste0(
    "rix ~~ ", paste(sprintf("0*wx%d", 1:T), collapse=" + "), "\n",
    "rix ~~ ", paste(sprintf("0*wy%d", 1:T), collapse=" + "), "\n",
    "riy ~~ ", paste(sprintf("0*wx%d", 1:T), collapse=" + "), "\n",
    "riy ~~ ", paste(sprintf("0*wy%d", 1:T), collapse=" + "), "\n"
  )

  # freely estimate within-person variances
  within_var <- paste0(
    paste(sprintf("wx%d ~~ wx%d", 1:T, 1:T), collapse="; "), "\n",
    paste(sprintf("wy%d ~~ wy%d", 1:T, 1:T), collapse="; "), "\n"
  )

  # within-person contemporaneous covariance
  within_cov <- paste0(
    paste(sprintf("wy%d ~~ wx%d", 1:T, 1:T), collapse="; "), "\n"
  )

  # dynamic (cross-lagged + autoregressive) paths
  regress <- paste(
    unlist(lapply(2:T, function(t){
      c(
        sprintf("wx%d ~ wx%d + wy%d", t, t-1, t-1),
        sprintf("wy%d ~ wx%d + wy%d", t, t-1, t-1)
      )
    })), collapse="\n"
  )

  # means across time
  means <- paste0(
    paste(paste0("x",1:T), collapse=" + "), " ~ mx*1\n",
    paste(paste0("y",1:T), collapse=" + "), " ~ my*1\n"
  )

  paste(ri_block, resid_fix, within_lat, orth,
        within_var, within_cov, regress, means, sep="\n")
}

```

### 2.4.4 DPM

and the DPM:

```{r build DPM function, eval=false, message=FALSE, warning=FALSE}

# ---------------------------------------------------------
# Function: build_dpm()
# ---------------------------------------------------------
# Purpose:
#   Build the dynamic panel model (DPM) used in Study 1.
#
#   FX loads on X2..XT
#   FY loads on Y2..YT
#
#   FX and FY are allowed to correlate with X1, Y1.
#   Each wave t > 1 is predicted from wave t-1:
#       Xt   ~ Xt-1 + Yt-1
#       Yt   ~ Xt-1 + Yt-1
#
#   Residual covariance X_t ~~ Y_t is freely estimated.
#   All means are freely estimated.
#
# Arguments:
#   T = number of waves (T ≥ 2)
#
# Returns:
#   A lavaan model string for the dynamic panel model
# ---------------------------------------------------------

build_dpm <- function(T) {

  # latent factors for slow processes (start at wave 2)
  FX_block <- paste0(
    "FX =~ ", paste(sprintf("1*x%d", 2:T), collapse=" + "), "\n"
  )

  FY_block <- paste0(
    "FY =~ ", paste(sprintf("1*y%d", 2:T), collapse=" + "), "\n"
  )

  # allow FX and FY to relate to the first wave
  fx_cov_block <- "FX ~~ x1 + y1\n"
  fy_cov_block <- "FY ~~ x1 + y1\n"

  # dynamic regression structure (lag-1)
  regress_block <- paste(
    unlist(lapply(2:T, function(t){
      c(
        sprintf("x%d ~ x%d + y%d", t, t-1, t-1),
        sprintf("y%d ~ x%d + y%d", t, t-1, t-1)
      )
    })), collapse="\n"
  )

  # residual covariance at each wave
  resid_cov_block <- paste(
    sprintf("x%d ~~ y%d", 1:T, 1:T),
    collapse="\n"
  )

  # latent factor covariances
  latent_cov_block <- paste(
    "FX ~~ FX",
    "FY ~~ FY",
    "FX ~~ FY",
    sep="\n"
  )

  # residual variances for all observed variables
  resid_var_block <- paste(
    paste(sprintf("x%d ~~ x%d", 1:T, 1:T), collapse="\n"),
    paste(sprintf("y%d ~~ y%d", 1:T, 1:T), collapse="\n"),
    sep="\n"
  )

  # freely estimate means at each wave
  means_block <- paste(
    paste(sprintf("x%d", 1:T), collapse=" + "), "~ 1\n",
    paste(sprintf("y%d", 1:T), collapse=" + "), "~ 1\n"
  )

  # combine all components into one model string
  paste(
    FX_block,
    FY_block,
    fx_cov_block,
    fy_cov_block,
    regress_block,
    resid_cov_block,
    latent_cov_block,
    resid_var_block,
    means_block,
    sep="\n"
  )
}

```


### 2.4.5 Linear BCA residualizer
and finally the linear BCA residualizer:

```{r build linear BCA function, eval=false, message=FALSE, warning=FALSE}

# ---------------------------------------------------------
# Function: residualise_panel_linearC()
# ---------------------------------------------------------
# Purpose:
#   Remove linear confounder effects from all observed
#   x_t and y_t scores by regressing them on c1...ck.
#   The residuals are then used in a standard CLPM.
#
# Arguments:
#   df        = data frame with x1..xT, y1..yT, c1..ck
#   x_prefix  = prefix for x variables
#   y_prefix  = prefix for y variables
#   c_prefix  = prefix for confounders
#
# Returns:
#   A data frame where x_t and y_t are replaced by residuals
# ---------------------------------------------------------

residualise_panel_linearC <- function(df,
                                      x_prefix = "x",
                                      y_prefix = "y",
                                      c_prefix = "c") {

  df <- as.data.frame(df)

  # detect names
  x_cols <- grep(paste0("^", x_prefix, "\\d+$"), names(df), value=TRUE)
  y_cols <- grep(paste0("^", y_prefix, "\\d+$"), names(df), value=TRUE)
  c_cols <- grep(paste0("^", c_prefix, "\\d+$"), names(df), value=TRUE)

  if (length(c_cols) == 0)
    stop("No confounder columns found.")

  # design matrix for confounders
  C <- as.matrix(df[c_cols])

  # residualise each x_t
  for (x in x_cols)
    df[[x]] <- resid(lm(df[[x]] ~ C))

  # residualise each y_t
  for (y in y_cols)
    df[[y]] <- resid(lm(df[[y]] ~ C))

  df
}

```

## 2.5 Next we need some functions that fit the models

The fitting functions for all aforementioned models are here all in one code chunck:

```{r model fitting functions, eval=false, message=FALSE, warning=FALSE}

# ---------------------------------------------------------
# Function: fit_clpm()
# ---------------------------------------------------------
# Purpose:
#   Fit the basic CLPM model using lavaan.
#
# Arguments:
#   data         = data frame with x1..xT, y1..yT
#   model_string = lavaan syntax from build_clpm()
#
# Returns:
#   A lavaan fit object, or NULL on failure
# ---------------------------------------------------------

fit_clpm <- function(data, model_string) {
  tryCatch(
    lavaan::lavaan(model_string,
                   data = as.data.frame(data),
                   estimator = "MLR",
                   warn = FALSE),
    error = function(e) NULL
  )
}

# ---------------------------------------------------------
# Function: fit_riclpm()
# ---------------------------------------------------------
# Purpose:
#   Fit the random-intercept CLPM.
#
# Arguments:
#   data         = panel data frame
#   model_string = lavaan syntax from build_riclpm()
#
# Returns:
#   A lavaan fit object, or NULL on failure
# ---------------------------------------------------------

fit_riclpm <- function(data, model_string) {
  tryCatch(
    lavaan::lavaan(model_string,
                   data = as.data.frame(data),
                   estimator = "MLR",
                   warn = FALSE),
    error = function(e) NULL
  )
}

# ---------------------------------------------------------
# Function: fit_dpm()
# ---------------------------------------------------------
# Purpose:
#   Fit the dynamic panel model (DPM) with slow processes FX, FY.
#
# Arguments:
#   data         = panel data frame
#   model_string = lavaan syntax from build_dpm()
#
# Returns:
#   A lavaan fit object, or NULL on failure
# ---------------------------------------------------------

fit_dpm <- function(data, model_string) {
  tryCatch(
    lavaan::lavaan(model_string,
                   data = as.data.frame(data),
                   estimator = "MLR",
                   warn = FALSE),
    error = function(e) NULL
  )
}

# ---------------------------------------------------------
# Function: fit_clpm_with_Cs()
# ---------------------------------------------------------
# Purpose:
#   Fit a CLPM where all confounders (c1..ck)
#   are included as predictors for x_t and y_t.
#
# Arguments:
#   data         = panel data with confounders
#   model_string = lavaan syntax from build_clpm_with_Cs()
#
# Returns:
#   A lavaan fit object, or NULL on failure
# ---------------------------------------------------------

fit_clpm_with_Cs <- function(data, model_string) {
  tryCatch(
    lavaan::lavaan(model_string,
                   data = as.data.frame(data),
                   estimator = "MLR",
                   warn = FALSE),
    error = function(e) NULL
  )
}

# ---------------------------------------------------------
# Function: fit_clpm_resid()
# ---------------------------------------------------------
# Purpose:
#   Residualise x_t and y_t on the confounders (linear model),
#   then fit a standard CLPM to the residualised data.
#
# Arguments:
#   data         = full data frame with x, y, c
#   model_string = lavaan syntax from build_clpm()
#
# Returns:
#   A lavaan fit object, or NULL on failure
# ---------------------------------------------------------

fit_clpm_resid <- function(data, model_string) {

  # remove confounder effects first
  df_resid <- tryCatch(
    residualise_panel_linearC(data),
    error = function(e) NULL
  )
  if (is.null(df_resid)) return(NULL)

  # fit the CLPM to the cleaned data
  tryCatch(
    lavaan::lavaan(model_string,
                   data = df_resid,
                   estimator = "MLR",
                   warn = FALSE),
    error = function(e) NULL
  )
}

```

## 2.6 Model output extraction function

Of course, now we need to also define a function which extracts the relevant outputs from this model.

```{r model output extraction function, eval=false, message=FALSE, warning=FALSE}

# ---------------------------------------------------------
# Function: extract_lagged_parameters()
# ---------------------------------------------------------
# Purpose:
#   Extract all AR and cross-lagged coefficients from a
#   fitted model (CLPM, RI-CLPM, DPM).
#
#   For CLPM / DPM:
#       x_t ~ x_{t-1} + y_{t-1}
#       y_t ~ x_{t-1} + y_{t-1}
#
#   For RI-CLPM:
#       wx_t ~ wx_{t-1} + wy_{t-1}
#       wy_t ~ wx_{t-1} + wy_{t-1}
#
# Arguments:
#   fit        = lavaan fit object
#   T          = number of waves
#   model_type = "clpm", "riclpm", or "dpm"
#
# Returns:
#   A list containing 4 numeric vectors:
#     ar_x[t]  = x_t <- x_{t-1}
#     ar_y[t]  = y_t <- y_{t-1}
#     xy[t]    = y_t <- x_{t-1}
#     yx[t]    = x_t <- y_{t-1}
#
#   Each vector has length T-1, indexed by t = 2..T.
# ---------------------------------------------------------

extract_lagged_parameters <- function(
    fit,
    T,
    model_type = c("clpm", "riclpm", "dpm")
){
  model_type <- match.arg(model_type)

  if (is.null(fit)) {
    return(list(
      ar_x = rep(NA, T-1),
      ar_y = rep(NA, T-1),
      xy   = rep(NA, T-1),
      yx   = rep(NA, T-1)
    ))
  }

  pe <- tryCatch(lavaan::parameterEstimates(fit), error=function(e) NULL)
  if (is.null(pe)) {
    return(list(
      ar_x = rep(NA, T-1),
      ar_y = rep(NA, T-1),
      xy   = rep(NA, T-1),
      yx   = rep(NA, T-1)
    ))
  }

  # variable prefixes: raw or RI-style
  if (model_type == "riclpm") {
    xvar <- "wx"
    yvar <- "wy"
  } else {
    xvar <- "x"
    yvar <- "y"
  }

  # helper to find an estimate in pe
  grab <- function(lhs, rhs){
    ix <- which(pe$lhs == lhs & pe$rhs == rhs)
    if (length(ix) == 0) return(NA)
    pe$est[ix[1]]
  }

  # initialize output vectors
  ar_x <- numeric(T-1)
  ar_y <- numeric(T-1)
  xy   <- numeric(T-1)
  yx   <- numeric(T-1)

  # loop over waves 2..T
  for (t in 2:T) {
    # AR coefficients
    ar_x[t-1] <- grab(paste0(xvar, t), paste0(xvar, t-1))
    ar_y[t-1] <- grab(paste0(yvar, t), paste0(yvar, t-1))

    # cross-lag X → Y
    xy[t-1]   <- grab(paste0(yvar, t), paste0(xvar, t-1))

    # cross-lag Y → X
    yx[t-1]   <- grab(paste0(xvar, t), paste0(yvar, t-1))
  }

  list(
    ar_x = ar_x,
    ar_y = ar_y,
    xy   = xy,
    yx   = yx
  )
}

```

## 2.7 wrappers

First we need a wrapper function to run one simulation iteration for a given scenario:

```{r simulation iteration wrapper function, eval=false, message=FALSE, warning=FALSE}

############################################################
##  LIBRARIES
############################################################

library(mvtnorm)
library(lavaan)
library(tidyverse)
library(parallel)
library(pbapply)

############################################################
##  ONE REPLICATION: STUDY 1 (ALL 5 SCENARIOS)
############################################################

run_one_rep_study1 <- function(
  rep_id,
  N,
  T,
  k,                 # number of linear confounders
  R2_1,              # total confounder R^2 at t = 1
  scenarios,         # vector of scenario names (lowercase)
  target_sd,         # SD of beta variation over time
  ar,                # true AR (x_t <- x_{t-1})
  cross,             # true cross-lag (y_t <- x_{t-1})
  rho,               # innovation correlation
  model_clpm,
  model_riclpm,
  model_dpm,
  model_clpm_with_Cs,
  base_seed = 12345
){

  set.seed(base_seed + rep_id)

  ## --- 1. Sample baseline B and re-use across all scenarios ----------------

  B1 <- sample_B_linear(
    k    = k,
    R2_1 = R2_1
  )

  # helper: pretty labels for plotting
  pretty_scen <- function(scen) {
    dplyr::recode(
      scen,
      "constant"   = "Constant",
      "linear"     = "Linear",
      "sinusoidal" = "Sinusoidal",
      "stepwise"   = "Stepwise",
      "random_walk"= "Random Walk",
      .default     = scen
    )
  }

  ## --- 2. Loop over scenarios and collect results --------------------------

  out_list <- vector("list", length(scenarios))

  for (j in seq_along(scenarios)) {

    scen <- scenarios[j]

    ## 2.1 Build B_t trajectory for this scenario
    B_list <- generate_B_trajectory(
      B1        = B1,
      T         = T,
      scenario  = scen,
      target_sd = target_sd
    )

    # beta as vectors (list-columns per occasion)
    beta_full <- lapply(B_list, function(Bt) as.numeric(Bt))      # length 2*k
    beta_X    <- lapply(B_list, function(Bt) as.numeric(Bt[1,]))  # length k
    beta_Y    <- lapply(B_list, function(Bt) as.numeric(Bt[2,]))  # length k

    ## 2.2 Simulate panel for this scenario

    df <- tryCatch(
      simulate_panel_data(
        N      = N,
        T      = T,
        B_list = B_list,
        ar     = ar,
        cross  = cross,
        rho    = rho
      ),
      error = function(e) NULL
    )

    # if simulation blew up completely → NA block for this scenario
    if (is.null(df)) {
      out_list[[j]] <- data.frame(
        run        = rep(rep_id, T),
        occasion   = 1:T,
        scenario   = pretty_scen(scen),

        beta       = I(beta_full),
        beta_X     = I(beta_X),
        beta_Y     = I(beta_Y),

        true_cross = cross,
        true_auto  = ar,

        est_CLPM        = NA_real_,
        est_RI_CLPM     = NA_real_,
        est_DPM         = NA_real_,
        est_CLPM_LBCA   = NA_real_,
        est_CLPM_Adj    = NA_real_,

        estA_CLPM       = NA_real_,
        estA_RI_CLPM    = NA_real_,
        estA_DPM        = NA_real_,
        estA_CLPM_LBCA  = NA_real_,
        estA_CLPM_Adj   = NA_real_,

        fail_CLPM       = TRUE,
        fail_RI_CLPM    = TRUE,
        fail_DPM        = TRUE,
        fail_CLPM_LBCA  = TRUE,
        fail_CLPM_Adj   = TRUE,

        is_na_run       = 1L
      )
      next
    }

    ## 2.3 Stationarity check: any Var(x_t) or Var(y_t) > 1? -----------------

    var_xy <- apply(df[, 1:(2*T)], 2, var)
    # small tolerance for numerical noise
    if (any(var_xy > 1 + 1e-3, na.rm = TRUE)) {

      out_list[[j]] <- data.frame(
        run        = rep(rep_id, T),
        occasion   = 1:T,
        scenario   = pretty_scen(scen),

        beta       = I(beta_full),
        beta_X     = I(beta_X),
        beta_Y     = I(beta_Y),

        true_cross = cross,
        true_auto  = ar,

        est_CLPM        = NA_real_,
        est_RI_CLPM     = NA_real_,
        est_DPM         = NA_real_,
        est_CLPM_LBCA   = NA_real_,
        est_CLPM_Adj    = NA_real_,

        estA_CLPM       = NA_real_,
        estA_RI_CLPM    = NA_real_,
        estA_DPM        = NA_real_,
        estA_CLPM_LBCA  = NA_real_,
        estA_CLPM_Adj   = NA_real_,

        fail_CLPM       = TRUE,
        fail_RI_CLPM    = TRUE,
        fail_DPM        = TRUE,
        fail_CLPM_LBCA  = TRUE,
        fail_CLPM_Adj   = TRUE,

        is_na_run       = 1L
      )
      next
    }

    ## 2.4 Fit all five models -----------------------------------------------

    # 1) CLPM (raw)
    fit_clpm_raw <- tryCatch(
      fit_clpm(df, model_clpm),
      error = function(e) NULL
    )

    # 2) RI-CLPM
    fit_ric <- tryCatch(
      fit_riclpm(df, model_riclpm),
      error = function(e) NULL
    )

    # 3) DPM
    fit_dpm0 <- tryCatch(
      fit_dpm(df, model_dpm),
      error = function(e) NULL
    )

    # 4) CLPM + Cs as covariates (per-occasion adjustment)
    fit_adj <- tryCatch(
      fit_clpm_with_Cs(df, model_clpm_with_Cs),
      error = function(e) NULL
    )

    # 5) CLPM after residualising on Cs (BCASM)
    fit_lbca <- tryCatch(
      fit_clpm_resid(df, model_clpm),
      error = function(e) NULL
    )

    ## 2.5 Extract AR + cross-lags per model --------------------------------

    lag_raw  <- extract_lagged_parameters(fit_clpm_raw, T, "clpm")
    lag_ric  <- extract_lagged_parameters(fit_ric,       T, "riclpm")
    lag_dpm0 <- extract_lagged_parameters(fit_dpm0,      T, "dpm")
    lag_adj  <- extract_lagged_parameters(fit_adj,       T, "clpm")
    lag_lbca <- extract_lagged_parameters(fit_lbca,      T, "clpm")

    fail_CLPM      <- is.null(fit_clpm_raw)
    fail_RI_CLPM   <- is.null(fit_ric)
    fail_DPM       <- is.null(fit_dpm0)
    fail_CLPM_Adj  <- is.null(fit_adj)
    fail_CLPM_LBCA <- is.null(fit_lbca)

    ## 2.6 Assemble long-format output (T rows, this scenario) --------------

    out_list[[j]] <- data.frame(
      run        = rep(rep_id, T),
      occasion   = 1:T,
      scenario   = pretty_scen(scen),

      beta       = I(beta_full),
      beta_X     = I(beta_X),
      beta_Y     = I(beta_Y),

      true_cross = cross,
      true_auto  = ar,

      # cross-lag X->Y (NA at t = 1)
      est_RI_CLPM   = c(NA, lag_ric$xy),
      est_DPM       = c(NA, lag_dpm0$xy),
      est_CLPM      = c(NA, lag_raw$xy),
      est_CLPM_LBCA = c(NA, lag_lbca$xy),
      est_CLPM_Adj  = c(NA, lag_adj$xy),

      # AR for X (NA at t = 1)
      estA_RI_CLPM   = c(NA, lag_ric$ar_x),
      estA_DPM       = c(NA, lag_dpm0$ar_x),
      estA_CLPM      = c(NA, lag_raw$ar_x),
      estA_CLPM_LBCA = c(NA, lag_lbca$ar_x),
      estA_CLPM_Adj  = c(NA, lag_adj$ar_x),

      fail_CLPM       = fail_CLPM,
      fail_RI_CLPM    = fail_RI_CLPM,
      fail_DPM        = fail_DPM,
      fail_CLPM_LBCA  = fail_CLPM_LBCA,
      fail_CLPM_Adj   = fail_CLPM_Adj,

      is_na_run = as.integer(
        all(is.na(c(lag_raw$xy,
                    lag_ric$xy,
                    lag_dpm0$xy,
                    lag_adj$xy,
                    lag_lbca$xy)))
      )
    )
  }

  # bind scenarios for this run
  dplyr::bind_rows(out_list)
}


```

and then a function to iteratively call this wrapper for multiple simulation repetitions:

```{r multiple replication wrapper function, eval=false, message=FALSE, warning=FALSE}

############################################################
##  MAIN SIMULATION: STUDY 1 (LINEAR ONLY, LONG OUTPUT)
############################################################

run_simulation_study1 <- function(
  reps      = 200,
  N         = 1000,
  T         = 5,
  k         = 3,
  R2_1      = 0.20,
  scenarios = c("constant","linear","sinusoidal","stepwise","random_walk"),
  target_sd = 0.10,
  ar        = 0.25,
  cross     = 0.10,
  rho       = 0.15,
  cores     = max(1, parallel::detectCores() - 1),
  base_seed = 12345
){

  # match/validate scenario names (allows subsets if you want)
  scenarios <- match.arg(scenarios,
                         choices = c("constant","linear","sinusoidal",
                                     "stepwise","random_walk"),
                         several.ok = TRUE)

  start_time <- Sys.time()

  ## 1. Build model strings once (on the master) -----------------------------

  model_clpm          <- build_clpm(T)
  model_riclpm        <- build_riclpm(T)
  model_dpm           <- build_dpm(T)
  model_clpm_with_Cs  <- build_clpm_with_Cs(T, k)

  tasks <- 1:reps

  ## 2. Start cluster --------------------------------------------------------

  cl <- parallel::makeCluster(cores)

  ## 3. Export everything needed to workers ---------------------------------

  parallel::clusterExport(
    cl,
    c(
      "sample_B_linear",
      "generate_B_trajectory",
      "simulate_panel_data",

      "build_clpm",
      "build_riclpm",
      "build_dpm",
      "build_clpm_with_Cs",

      "fit_clpm",
      "fit_riclpm",
      "fit_dpm",
      "fit_clpm_with_Cs",
      "fit_clpm_resid",

      "residualise_panel_linearC",
      "extract_lagged_parameters",
      "run_one_rep_study1",

      # fixed scalar arguments:
      "N","T","k","R2_1","scenarios","target_sd",
      "ar","cross","rho",
      "model_clpm","model_riclpm","model_dpm","model_clpm_with_Cs",
      "base_seed"
    ),
    envir = environment()
  )

  ## 4. Load packages on workers --------------------------------------------

  parallel::clusterEvalQ(cl, {
    library(mvtnorm)
    library(lavaan)
  })

  ## 5. Parallel loop with pbapply progress bar -----------------------------

  res_list <- pbapply::pblapply(
    X  = tasks,
    cl = cl,
    FUN = function(rid) {
      run_one_rep_study1(
        rep_id           = rid,
        N                = N,
        T                = T,
        k                = k,
        R2_1             = R2_1,
        scenarios        = scenarios,
        target_sd        = target_sd,
        ar               = ar,
        cross            = cross,
        rho              = rho,
        model_clpm       = model_clpm,
        model_riclpm     = model_riclpm,
        model_dpm        = model_dpm,
        model_clpm_with_Cs = model_clpm_with_Cs,
        base_seed        = base_seed
      )
    }
  )

  parallel::stopCluster(cl)

  raw_df  <- dplyr::bind_rows(res_list)
  runtime <- Sys.time() - start_time

  attr(raw_df, "runtime") <- runtime
  raw_df
}


```

## 2.8 Running a simulation function

Even though we have defined all code above, debugging that is a huge pain in the ass (mainly caused by the parralell exectution). The true code that generates the results is stored in a script. Do note that it works the same. 

```{r run simulation example}

# get the functions
source("C:/Users/Admin/Desktop/UU/Thesis umbrella/Thesis/Code/Qmd_Rmd_R_Pyth/Sim Studies/Sim_Study_1_forqmd.R")

# create the output
out_long <- run_simulation_study1(
  reps      = 250,
  N         = 1000,
  T         = 5,
  k         = 3,
  R2_1      = 0.25,
  target_sd = 0.15,
  ar        = 0.2,
  cross     = 0.10,
  rho       = 0.15
)

# save the output as a dataframe
saveRDS(
  out_long,
  file = "C:/Users/Admin/Desktop/UU/Thesis umbrella/Thesis/Code/Data/out_long_study1.rds"
)

```

## 2.9 Making the plots

Now we make the plots we discussed earlier:

* Plot 1: 

```{r }

# read the rds file
df <- readRDS("C:/Users/Admin/Desktop/UU/Thesis umbrella/Thesis/Code/Data/out_long_study1.rds")
str(df)


# colours
blue_magenta <- c(
  "#3B4CC0",  
  "#5E60CE",
  "#7A70D1",
  "#9775D3",
  "#B67BD2",
  "#D084CF",
  "#E28FC4",
  "#F29BC1",
  "#FF9ACD"
)

method_colors <- c(
  "RI_CLPM"   = "#3B4CC0",
  "DPM"       = "#5E60CE",
  "CLPM"      = "#7A70D1",
  "CLPM_LBCA" = "#B67BD2",
  "CLPM_Adj"  = "#EE3A8C"
)

# extend colours to number of runs
n_runs <- length(unique(df$run))
blue_magenta_extended <- rep(blue_magenta, length.out = n_runs)
plot_beta_X <- df %>%
  mutate(beta_X = map_dbl(beta_X, ~ .x)) %>%
  ggplot(aes(x = occasion, y = beta_X,
             group = run, color = factor(run))) +
  geom_line(linewidth = 0.7, alpha = 0.9) +
  facet_wrap(~ scenario) +
  scale_color_manual(values = blue_magenta_extended) +
  guides(color = "none") +
  labs(x = "Occasion", y = "Effect of Confounder on X") +
  theme_minimal()

plot_beta_X

plot_beta_Y <- df %>%
  mutate(beta_Y = map_dbl(beta_Y, ~ .x)) %>%
  ggplot(aes(x = occasion, y = beta_Y,
             group = run, color = factor(run))) +
  geom_line(linewidth = 0.7, alpha = 0.9) +
  facet_wrap(~ scenario) +
  scale_color_manual(values = blue_magenta_extended) +
  guides(color = "none") +
  labs(x = "Occasion", y = "Effect of Confounder on Y") +
  theme_minimal()

plot_beta_Y

plot_R2_X <- df %>%
  mutate(beta_X = map_dbl(beta_X, ~ .x),
         R2_X = beta_X^2) %>%
  ggplot(aes(x = occasion, y = R2_X,
             group = run, color = factor(run))) +
  geom_line(linewidth = 0.7, alpha = 0.9) +
  facet_wrap(~ scenario) +
  scale_color_manual(values = blue_magenta_extended) +
  guides(color = "none") +
  labs(x = "Occasion", y = expression(R^2 ~ "(confounder → X)")) +
  theme_minimal()

plot_R2_X

plot_bias_cross <- df %>%
  mutate(
    rb_RI_CLPM   = abs(est_RI_CLPM   - true_cross) / true_cross,
    rb_DPM       = abs(est_DPM       - true_cross) / true_cross,
    rb_CLPM      = abs(est_CLPM      - true_cross) / true_cross,
    rb_CLPM_LBCA = abs(est_CLPM_LBCA - true_cross) / true_cross,
    rb_CLPM_Adj  = abs(est_CLPM_Adj  - true_cross) / true_cross
  ) %>%
  pivot_longer(
    starts_with("rb_"),
    names_to = "method",
    values_to = "rel_bias"
  ) %>%
  mutate(method = gsub("rb_", "", method),
         method = factor(method, levels = names(method_colors))) %>%
  group_by(scenario, method, occasion) %>%
  summarise(mean_bias = mean(rel_bias, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = occasion, y = mean_bias,
             color = method, group = method)) +
  geom_point(size = 2) +
  geom_line(linewidth = 0.8) +
  facet_grid(method ~ scenario) +
  scale_color_manual(values = method_colors) +
  theme_minimal() +
  labs(x = "Occasion", y = "Mean Relative Bias (Cross-Lagged)")

plot_bias_cross

plot_se_cross <- df %>%
  mutate(
    rb_RI_CLPM   = abs(est_RI_CLPM   - true_cross) / true_cross,
    rb_DPM       = abs(est_DPM       - true_cross) / true_cross,
    rb_CLPM      = abs(est_CLPM      - true_cross) / true_cross,
    rb_CLPM_LBCA = abs(est_CLPM_LBCA - true_cross) / true_cross,
    rb_CLPM_Adj  = abs(est_CLPM_Adj  - true_cross) / true_cross
  ) %>%
  pivot_longer(
    starts_with("rb_"),
    names_to = "method",
    values_to = "rel_bias"
  ) %>%
  mutate(method = gsub("rb_", "", method),
         method = factor(method, levels = names(method_colors))) %>%
  group_by(scenario, method, occasion) %>%
  summarise(se_bias = sd(rel_bias, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = occasion, y = se_bias,
             color = method, group = method)) +
  geom_point(size = 2) +
  geom_line(linewidth = 0.8) +
  facet_grid(method ~ scenario) +
  scale_color_manual(values = method_colors) +
  theme_minimal() +
  labs(x = "Occasion", y = "Empirical SE (Cross-Lagged)")

plot_se_cross

plot_summary <- df %>%
  select(scenario, run, starts_with("est_"), true_cross) %>%
  pivot_longer(
    starts_with("est_"),
    names_to = "method",
    values_to = "estimate"
  ) %>%
  mutate(method = gsub("est_", "", method),
         method = factor(method, levels = names(method_colors))) %>%
  group_by(scenario, method) %>%
  summarise(
    mean_est = mean(estimate, na.rm = TRUE),
    sd_est   = sd(estimate,   na.rm = TRUE),
    true_val = unique(true_cross),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = method, y = mean_est, color = method)) +
  geom_point(size = 2.5) +
  geom_errorbar(aes(ymin = mean_est - sd_est,
                    ymax = mean_est + sd_est),
                width = 0.15, linewidth = 0.8) +
  geom_hline(aes(yintercept = true_val), linetype = "dashed") +
  facet_grid(scenario ~ .) +
  scale_color_manual(values = method_colors) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 25, hjust = 1),
    legend.position = "none"
  ) +
  labs(x = "Method", y = "Estimated Cross-Lagged Effect")

plot_summary

```

# 3. SImulation study 2

## 3.8 Running the simulation function

```{r}

# load the simulation engine
source("C:/Users/Admin/Desktop/UU/Thesis umbrella/Thesis/Code/Qmd_Rmd_R_Pyth/Sim Studies/Sim_Study_2_forqmd.R")

out_long <- run_simulation_study2(
  reps      = 10,
  N         = 1000,
  T         = 5,
  k_lin     = 3,
  ax        = 0.25,
  ay        = 0.25,
  bx        = 0.10,
  by        = 0.10,
  rho       = 0.15,
  R2_1      = 0.25,
  eta_1     = 0.40,
  target_sd = 0.10,
  rw_sd     = 0.10,
  cores     = 6,
  base_seed = 12345
)

# save the results
saveRDS(
  out_long,
  file = "C:/Users/Admin/Desktop/UU/Thesis umbrella/Thesis/Code/Data/out_long_study2.rds"
)

```

## 3.9 Making the plots

```{r}

df <- readRDS("C:/Users/Admin/Desktop/UU/Thesis umbrella/Thesis/Code/Data/out_long_study2.rds")
str(df)

library(tidyverse)




```
