\documentclass{article}
\usepackage{graphicx}
\usepackage{amsmath}

\begin{document}

\section*{Notes on Variance Calculations}
The mathematical derivations showing how we can achieve the following in our panel data simulations:

\begin{itemize}
    \item The variance of $X$, $Y$ and $C$ is 1.
    \item Each variable has a mean 0.
    \item The autoregressive effects of $X_{t-1}$ on $X_t$ is 0.25 (same for $Y_{t-1}$ on $Y_t$).
    \item The lag-1 cross effect of $X_{t-1}$ on $Y_t$ is 0.10 (same for $Y_{t-1}$ on $X_t$).
    \item There are 3 confounders, and hence $k=3$.
    \item Their effects differ for X and Y, such that X is affected at [0.30, 0.20, 0.10] and Y is affected at [0.30, 0.20, 0.10].
    \item The innovations have residual correlations of 0.3 at each time-point, but not across time-points.
    \item At the first time-point ($T=0$), X and Y are correlated due to the confounding variables. 
    \item At the first time-point, we can assume that the dynamic process was going on before the data started being collected. Hence,
    the 'observed' covariance matrix S between X and Y should be the same for the first measurement occassion. But since we do not have 
    any previous time-point creating feed forward effects, X and Y should be simulated with 'extra' covariance as if they were 'caused' 
    by previous variables. 
    \item One possible extention: confounders in reality are with a very high probability not uncorrelated. In the event that these are 
    correlated, we get a true matrix Psi instead of the indentity matrix I. Calculations are not affected with respect to unknown values. 
\end{itemize}

What we need to do, is to find the right variance for the innovations such that the total variance of X and Y becomes 1. We also need to find
out what correlation X and Y are supposed to have at each time-point. As such we have three unknowns: a variance of the innovation for X,
a variance of the innovation for Y, and a correlation between X and Y. We must pick these values such that all our demands are met. In essence,
this is the purpose of all math done here. 

\newpage

\subsection*{Defining matrices}

\noindent
We have variables $X$ over occasions $T$ and $Y$ over occasions $T$. This yields the vector:

\[
Z_t = \begin{bmatrix} X_t \\ Y_t \end{bmatrix}
\]

\noindent
For one individual \textit{i}, the vector looks like:

\[
Z_i =
\begin{bmatrix}
X_{i1} \\
Y_{i1} \\
X_{i2} \\
Y_{i2} \\
X_{i3} \\
Y_{i3} \\
X_{i4} \\
Y_{i4} \\
X_{i5} \\
Y_{i5}
\end{bmatrix}.
\]

\noindent
Expanding for $N$ individuals:

\[
Z =
\begin{bmatrix}
X_{1,1} & Y_{1,1} & X_{1,2} & Y_{1,2} & X_{1,3} & Y_{1,3} & X_{1,4} & Y_{1,4} & X_{1,5} & Y_{1,5} \\
X_{2,1} & Y_{2,1} & X_{2,2} & Y_{2,2} & X_{2,3} & Y_{2,3} & X_{2,4} & Y_{2,4} & X_{2,5} & Y_{2,5} \\
\vdots  & \vdots  & \vdots  & \vdots  & \vdots  & \vdots  & \vdots  & \vdots  & \vdots  & \vdots  \\
X_{N,1} & Y_{N,1} & X_{N,2} & Y_{N,2} & X_{N,3} & Y_{N,3} & X_{N,4} & Y_{N,4} & X_{N,5} & Y_{N,5}
\end{bmatrix}.
\]

\noindent
Now for time-point $t$, $X_t$ and $Y_t$ are made up of some part of the confounder effect, and some part by the dynamic efects. With
the dynamic effects here we mean every effect that 'feeds forward' to the next time-point. For the first time-point, there are no dynamic
effects at all, since there are no previous time-points. Since we assume the process does not start with measurement, the observed correlation
matrix between X and Y should be unnafected though. Another way of phrasing this is that some part of the variance in X (or Y) is explained 
by the confounders, some part is explained by the dynamic effects, and some part is explained by the innovations. We can capture this in the formula:

\[
Z_t = A Z_{t-1} + B U + E_t,
\]

\noindent
where

\begin{itemize}
    \item $A$ is the matrix of autoregressive and cross-lagged effects,
    \item $B$ is the matrix of confounder effects,
    \item $U$ is the vector of time-invariant confounders,
    \item $E_t$ is the vector of innovations at time-point $t$.
\end{itemize}

\noindent
Here we can clearly see that if there is no previous time-point, the dynamic effects part ($A Z_{t-1}$) will be zero. If $Z_t$ is then to have
a similar covariance matrix, the part of variance received from the previous time-points and the lagged effects should be 'compensated'. 

\noindent
If we choose our effects as denoted above we get the following $A$ matrix:

\[
A = \begin{bmatrix} a_x & b_y \\ b_x & a_y \end{bmatrix} = \begin{bmatrix} 0.25 & 0.10 \\ 0.10 & 0.25 \end{bmatrix}
\]

\noindent
We get the following beta matrix:

\[
B = \begin{bmatrix} c_{x1} & c_{x2} & c_{x3} \\ c_{y1} & c_{y2} & c_{y3} \end{bmatrix} = \begin{bmatrix} 0.30 & 0.20 & 0.10 \\ 0.30 & 0.20 & 0.10 \end{bmatrix}
\]

\noindent
And matrix U:

\[
U =
\begin{bmatrix}
c_{1,1} & c_{1,2} & c_{1,3} \\
c_{2,1} & c_{2,2} & c_{2,3} \\
c_{3,1} & c_{3,2} & c_{3,3} \\
\vdots  & \vdots  & \vdots  \\
c_{N,1} & c_{N,2} & c_{N,3}
\end{bmatrix}
\]

\noindent
Finally, we can specify the innovations matrix as:

\[E_t = \begin{bmatrix} e_{X_t} \\ e_{Y_t} \end{bmatrix}\]

\newpage

\subsection*{Confounder variance calculations}

\noindent
As we just denoted, each measurement occasion takes some variance from the confounders and some from the dynamic effect. The variance received from
the confounders is time-invariant, which means we can calculate it directly. Since the variances of all variables (and thus also the confounders) is 1,
and the confoudners are not correlated with each other, their covariance matrix is the identity matrix $I$. Note that including confounder covariance
does not change very much about the calculations, since its a value we just pick and insert using psi. Hence, the variance received from the confounders is:

\[S_U = B \Psi B^T = B I B^T = B B^T\]

\noindent
If we use the values from above, we get: 

\[S_U = \begin{bmatrix} 0.30 & 0.20 & 0.10 \\ 0.30 & 0.20 & 0.10 \end{bmatrix} \begin{bmatrix} 0.30 & 0.30 \\ 0.20 & 0.20 \\ 0.10 & 0.10 \end{bmatrix} = \begin{bmatrix} 0.14 & 0.14 \\ 0.14 & 0.14 \end{bmatrix}\]

\newpage

\subsection*{The innovation covariance matrix}

\noindent
The confounders add a variance of 0.14 to both X and Y, and a covariance of 0.14 between X and Y at each time-point. We wanted each variable to have a variance of 1, 
which means that the dynamic effect still needs to add 1-0.14 = 0.86 to the variance of each variable. We also want the innovations to have a correlation of 0.3
at each time-point, which means that the part of X and Y that cannot be explained by $X_{t-1}$ and $Y_{t-1}$ or C needs to be correlated at 0.3. In math notation, we want:

\[Corr(e_{X_t}, e_{Y_t}) = \rho = 0.3\]

\noindent
We can denote the variance of the innovations as:

\[Var(e_{X_t}) = v_1\] and \[Var(e_{Y_t}) = v_2\]

\noindent
The covariance between the innovations is then their correlation multiplied by their standard deviations:

\[Cov(e_{X_t}, e_{Y_t}) = \rho \sqrt{v_1} \sqrt{v_2}\]

\noindent
The covariance matrix of the innovations can then be written as:

\[S_E = \begin{bmatrix} v_1 & \rho \sqrt{v_1} \sqrt{v_2} \\ \rho \sqrt{v_1} \sqrt{v_2} & v_2 \end{bmatrix}\]

\noindent 
Since we know that rho is 0.3, we can rewrite this as:

\[S_E = \begin{bmatrix} v_1 & 0.3 \sqrt{v_1} \sqrt{v_2} \\ 0.3 \sqrt{v_1} \sqrt{v_2} & v_2 \end{bmatrix}\]

\noindent
Which leaves us with two unknown values: $v_1$ and $v_2$. We need to choose these such that the total variance of X and Y becomes 1. 
Another way of phrasing this requirement is that $S$, which is the total variance-covariance matrix of $Z_t$, needs to have 1s on the diagonal.

\newpage

\subsection*{The covariance matrix S (of X and Y)}

\noindent
Now first we need to specify what $S$ looks like. We know that:

\[S = var(Z_t) = \begin{bmatrix} var(X_t) & cov(X_t, Y_t) \\ cov(X_t, Y_t) & var(Y_t) \end{bmatrix}\]

\noindent
We want the variance of X and Y to be 1, which means that we want:

\[var(X_t) = var(Y_t) = 1\], yielding:

\[S = \begin{bmatrix} 1 & cov(X_t, Y_t) \\ cov(X_t, Y_t) & 1 \end{bmatrix}\]

\noindent
We can now denote $cov(X_t, Y_t)$ as $c$. This yields:

\[S = \begin{bmatrix} 1 & c \\ c & 1 \end{bmatrix}\]

\noindent
Note that this covariance matrix is statiorary when it is the same for all time-points. In order to satisfy this constraint, the following equation must hold:

\[S = A S A^T + S_U + S_E\]

\noindent
The covariance matrix of the current time-point, is equal to that of the previous time-point, multiplied by the cross-lagged and autoregressive effects, 
plus the variance received from the confounders, plus the variance received from the innovations. We must solve this equation for $v_1$, $v_2$ and $c$. 
$v1$ and $v_2$ are only present in $S_E$, while $c$ is only present in $S$. Expanding the equation yields:

\[
\begin{bmatrix}
1 & c \\
c & 1
\end{bmatrix}
=
\begin{bmatrix}
(ASA^\top)_{11} & (ASA^\top)_{12} \\
(ASA^\top)_{21} & (ASA^\top)_{22}
\end{bmatrix}
+
\begin{bmatrix}
0.14 & 0.14 \\
0.14 & 0.14
\end{bmatrix}
+
\begin{bmatrix}
v_1 & \rho\sqrt{v_1 v_2} \\
\rho\sqrt{v_1 v_2} & v_2
\end{bmatrix}.
\]

\noindent
Note that on the left side is our desired stationary variance covariance matrix between X and Y, containing unknown $c$. On the right 
side we have the 






\end{document}

% pdflatex Notes_on_variance_calculations.tex
% Remove-Item "*.aux","*.log","*.toc","*.out","*.synctex.gz" -Force
