---
title: "Research Rapport Results"
format: html
---

First, let's generate the results for simulation study 1. Let's grab our simulation engine from our source file. 

```{r sim study 1}

# load the required functions and dependencies
source("Thesis\\Code\\Qmd_Rmd_R_Pyth\\Sim Studies\\Sim_Study_1_RR.R")

# autoregressive + cross-lag structure
A <- matrix(c(
  0.20, 0,                         # autoregressive and cross-lag for Y
  0.10, 0.20                       # cross-lag and autoregressive for X
), nrow = 2, byrow = TRUE)

# confounder covariance (3 confounders)
Psi <- diag(3)                     # uncorrelated confounders with var=1

# run the simulation
results_s1 <- run_simulation_study1(
  reps          = 1000,                        # number of replications
  N             = 1000,                        # sample size
  T             = 5,                           # number of waves
  k             = 3,                           # number of confounders
  R2_1          = 0.15,                        # confounder R2 at wave 1
  target_sd     = 0.5,                        # SD of B(t) variation
  scenarios     = c("constant", "stepwise"),   # scenarios to run
  A             = A,                           # 2×2 AR + cross-lag matrix
  Psi           = Psi,                         # confounder covariance
  rho_extra     = 0.1,                         # extra X–Y covariance
  models_to_run = c("clpm","riclpm",           # 5 modeling approaches 
                    "dpm","lbca","adj"),
  cores         = 6,                           # number of cores to use
  base_seed     = 12345                        # reproducible base seed
)

```

Now, let's visualize the results for simulation study 1.

```{r sim study 1 visualization}

# representation of the dataframe 
df <- results_s1
head(df)

# custom colour palette
blue_magenta <- c(
  "#3B4CC0", "#5E60CE", "#7A70D1", "#9775D3", "#B67BD2",
  "#D084CF", "#E28FC4", "#F29BC1", "#FF9ACD"
)

# dxtend colours to number of runs
n_runs <- length(unique(df$run))
blue_magenta_extended <- rep(blue_magenta, length.out = n_runs)

###############################################################################
# plot 1: visualize how beta_X and beta_Y change over occasions
###############################################################################

# the effect of the confounder on X and Y over occasions, by scenario
plot_beta_X <- df %>%

  # the occasion on the X-axis, beta_X on the Y-axis, grouped and coloured by run
  # note that beta_X is the average over the 3 confounders
  ggplot(aes(x = occasion, y = beta_X,
             group = run, color = factor(run))) +

  # determines the look of the line
  geom_line(linewidth = 0.7, alpha = 0.9) +
  
  # separate plots for each scenario
  facet_wrap(~ scenario) +
  
  # use the custom colour palette
  scale_color_manual(values = blue_magenta_extended) +

  # remove the legend
  guides(color = "none") +

  # labels for the axes
  labs(x = "Occasion", y = "Effect of Confounder C on X") +

  # minimal theme for the plot
  theme_minimal()

# precisely the same for beta_Y
plot_beta_Y <- df %>%
  ggplot(aes(x = occasion, y = beta_Y,
             group = run, color = factor(run))) +
  geom_line(linewidth = 0.7, alpha = 0.9) +
  facet_wrap(~ scenario) +
  scale_color_manual(values = blue_magenta_extended) +
  guides(color = "none") +
  labs(x = "Occasion", y = "Effect of Confounder on Y") +
  theme_minimal()

###############################################################################
# plot 2: visualise how the explained variance changes over occasions
###############################################################################
plot_R2_X <- df %>%
  
  # take the square of beta_X to get R2_X
  mutate(R2_X = beta_X^2) %>%
  
  # the occasion on the X-axis, R2_X on the Y-axis, grouped and coloured by run
  ggplot(aes(
    x = occasion,
    y = R2_X,
    group = run,
    color = factor(run)
  )) +

  # determines the look of the line
  geom_line(linewidth = 0.7, alpha = 0.9) +

  # separate plots for each scenario
  facet_wrap(~ scenario) +

  # use the custom colour palette
  scale_color_manual(values = blue_magenta_extended) +

  # remove the legend
  guides(color = "none") +

  # labels for the axes
  labs(
    x = "Occasion",
    y = "Explained Variance of Confounder C on X"
  ) +

  # minimal theme for the plot
  theme_minimal()

# same for R2_Y
plot_R2_Y <- df %>%
  mutate(R2_Y = beta_Y^2) %>%
  ggplot(aes(
    x = occasion,
    y = R2_Y,
    group = run,
    color = factor(run)
  )) +
  geom_line(linewidth = 0.7, alpha = 0.9) +
  facet_wrap(~ scenario) +
  scale_color_manual(values = blue_magenta_extended) +
  guides(color = "none") +
  labs(
    x = "Occasion",
    y = "Explained Variance of Confounder C on Y"
  ) +
  theme_minimal()

###############################################################################
# plot 3: show the mean absolute bias of the different models 
###############################################################################

# true effects extracted from A
true_xy <- A[2,1]   # XY (CL)
true_yx <- A[1,2]   # YX (CL)
true_ax <- A[1,1]   # X (AR)
true_ay <- A[2,2]   # Y (AR)

# cross-lagged estimates mean absolute bias X on Y 
bias_XY <- df %>%
  select(run, occasion, scenario, starts_with("estXY_")) %>%
  
  # long format
  pivot_longer(
    cols      = starts_with("estXY_"),
    names_to  = "model",
    values_to = "estimate"
  ) %>%
  
  # compute absolute bias
  mutate(abs_bias = abs(estimate - true_xy)) %>%
  
  # compute mean absolute bias per scenario × model × occasion
  group_by(scenario, model, occasion) %>%
  summarise(mean_abs_bias = mean(abs_bias, na.rm = TRUE),
            .groups = "drop")

# plot mean abs bias for XY
plot_bias_XY <- bias_XY %>%
  ggplot(aes(x = occasion, y = mean_abs_bias, color = model)) +
  geom_line(linewidth = 0.9) +
  facet_grid(rows = vars(scenario), cols = vars(model), scales = "free_y") +
  labs(x = "Occasion",
       y = "Mean Absolute Bias",
       title = "Bias: Cross-Lag XY") +
  theme_minimal() +
  theme(strip.text = element_text(size = 10),
        legend.position = "none")

# same but for YX
bias_YX <- df %>%
  select(run, occasion, scenario, starts_with("estYX_")) %>%
  pivot_longer(
    cols      = starts_with("estYX_"),
    names_to  = "model",
    values_to = "estimate"
  ) %>%
  mutate(abs_bias = abs(estimate - true_yx)) %>%
  group_by(scenario, model, occasion) %>%
  summarise(mean_abs_bias = mean(abs_bias, na.rm = TRUE),
            .groups = "drop")

plot_bias_YX <- bias_YX %>%
  ggplot(aes(x = occasion, y = mean_abs_bias, color = model)) +
  geom_line(linewidth = 0.9) +
  facet_grid(rows = vars(scenario), cols = vars(model), scales = "free_y") +
  labs(x = "Occasion",
       y = "Mean Absolute Bias",
       title = "Bias: Cross-Lag YX") +
  theme_minimal() +
  theme(strip.text = element_text(size = 10),
        legend.position = "none")


# compute mean absolute bias for AR(X)
bias_AX <- df %>%
  select(run, occasion, scenario,
         estA_CLPM, estA_RI_CLPM, estA_DPM,
         estA_CLPM_Adj, estA_CLPM_LBCA) %>%
  
  pivot_longer(
    cols      = starts_with("estA_"),
    names_to  = "model",
    values_to = "estimate"
  ) %>%
  
  mutate(abs_bias = abs(estimate - true_ax)) %>%
  
  group_by(scenario, model, occasion) %>%
  summarise(mean_abs_bias = mean(abs_bias, na.rm = TRUE),
            .groups = "drop")

plot_bias_AX <- bias_AX %>%
  ggplot(aes(x = occasion, y = mean_abs_bias, color = model)) +
  geom_line(linewidth = 0.9) +
  facet_grid(rows = vars(scenario),
             cols = vars(model),
             scales = "free_y") +
  labs(
    x = "Occasion",
    y = "Mean Absolute Bias",
    title = "Bias of AX Estimates"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 10),
    legend.position = "none"
  )

bias_AY <- df %>%
  select(run, occasion, scenario,
         estAY_CLPM, estAY_RI_CLPM, estAY_DPM,
         estAY_CLPM_Adj, estAY_CLPM_LBCA) %>%
  
  pivot_longer(
    cols      = starts_with("estAY_"),
    names_to  = "model",
    values_to = "estimate"
  ) %>%
  
  mutate(abs_bias = abs(estimate - true_ay)) %>%
  
  group_by(scenario, model, occasion) %>%
  summarise(mean_abs_bias = mean(abs_bias, na.rm = TRUE),
            .groups = "drop")

plot_bias_AY <- bias_AY %>%
  ggplot(aes(x = occasion, y = mean_abs_bias, color = model)) +
  geom_line(linewidth = 0.9) +
  facet_grid(rows = vars(scenario),
             cols = vars(model),
             scales = "free_y") +
  labs(
    x = "Occasion",
    y = "Mean Absolute Bias",
    title = "Bias of AY Estimates"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 10),
    legend.position = "none"
  )

###############################################################################
# plot 4: the empirical standard error of the estimates over time
###############################################################################

# reshape and compute empirical standard error for cross-lag X→Y
se_crosslag_XY <- df %>%
  
  # select cross-lag X→Y estimates
  select(run, occasion, scenario,
         estXY_CLPM, estXY_RI_CLPM, estXY_DPM,
         estXY_CLPM_Adj, estXY_CLPM_LBCA) %>%
  
  # pivot to long format
  pivot_longer(
    cols = starts_with("estXY_"),
    names_to = "model",
    values_to = "estimate"
  ) %>%
  
  # compute empirical SE: sd of estimates over runs
  group_by(scenario, model, occasion) %>%
  summarise(
    emp_se = sd(estimate, na.rm = TRUE),
    .groups = "drop"
  )

# plot empirical SE (scenarios = rows, methods = columns)
plot_se_crosslag_XY <- se_crosslag_XY %>%
  ggplot(aes(x = occasion, y = emp_se, color = model)) +
  geom_line(linewidth = 0.9) +
  facet_grid(rows = vars(scenario), cols = vars(model), scales = "free_y") +
  labs(
    x = "Occasion",
    y = "Empirical SE",
    title = "Empirical Standard Error: Cross-lag XY"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 10),
    legend.position = "none"
  )

# reshape and compute empirical standard error for cross-lag Y→X
se_crosslag_YX <- df %>%
  
  # select cross-lag Y→X estimates
  select(run, occasion, scenario,
         estYX_CLPM, estYX_RI_CLPM, estYX_DPM,
         estYX_CLPM_Adj, estYX_CLPM_LBCA) %>%
  
  pivot_longer(
    cols = starts_with("estYX_"),
    names_to = "model",
    values_to = "estimate"
  ) %>%
  
  group_by(scenario, model, occasion) %>%
  summarise(
    emp_se = sd(estimate, na.rm = TRUE),
    .groups = "drop"
  )

plot_se_crosslag_YX <- se_crosslag_YX %>%
  ggplot(aes(x = occasion, y = emp_se, color = model)) +
  geom_line(linewidth = 0.9) +
  facet_grid(rows = vars(scenario), cols = vars(model), scales = "free_y") +
  labs(
    x = "Occasion",
    y = "Empirical SE",
    title = "Empirical Standard Error: Cross-lag YX"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 10),
    legend.position = "none"
  )

# reshape and compute empirical standard error for AR(X→X)
se_AR_X <- df %>%
  
  # select AR(X→X) estimates
  select(run, occasion, scenario,
         estA_CLPM, estA_RI_CLPM, estA_DPM,
         estA_CLPM_Adj, estA_CLPM_LBCA) %>%
  
  pivot_longer(
    cols = starts_with("estA_"),
    names_to = "model",
    values_to = "estimate"
  ) %>%
  
  group_by(scenario, model, occasion) %>%
  summarise(
    emp_se = sd(estimate, na.rm = TRUE),
    .groups = "drop"
  )

plot_se_AR_X <- se_AR_X %>%
  ggplot(aes(x = occasion, y = emp_se, color = model)) +
  geom_line(linewidth = 0.9) +
  facet_grid(rows = vars(scenario), cols = vars(model), scales = "free_y") +
  labs(
    x = "Occasion",
    y = "Empirical SE",
    title = "Empirical Standard Error: AX"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 10),
    legend.position = "none"
  )

# reshape and compute empirical standard error for AR(Y→Y)
se_AR_Y <- df %>%
  
  # select AR(Y→Y) estimates
  select(run, occasion, scenario,
         estAY_CLPM, estAY_RI_CLPM, estAY_DPM,
         estAY_CLPM_Adj, estAY_CLPM_LBCA) %>%
  
  pivot_longer(
    cols = starts_with("estAY_"),
    names_to = "model",
    values_to = "estimate"
  ) %>%
  
  group_by(scenario, model, occasion) %>%
  summarise(
    emp_se = sd(estimate, na.rm = TRUE),
    .groups = "drop"
  )

plot_se_AR_Y <- se_AR_Y %>%
  ggplot(aes(x = occasion, y = emp_se, color = model)) +
  geom_line(linewidth = 0.9) +
  facet_grid(rows = vars(scenario), cols = vars(model), scales = "free_y") +
  labs(
    x = "Occasion",
    y = "Empirical SE",
    title = "Empirical Standard Error: AR(Y → Y)"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 10),
    legend.position = "none"
  )

###############################################################################
# plot 5: summary plot 
###############################################################################

# true cross-lag effect X on Y
true_xy <- A[2, 1]

# compute per-run average estimate and mean absolute bias
box_crosslag_X <- df %>%
  
  # pull cross-lag X estimates
  select(run, occasion, scenario,
         estXY_CLPM, estXY_RI_CLPM, estXY_DPM,
         estXY_CLPM_Adj, estXY_CLPM_LBCA) %>%
  
  # reshape to long format
  pivot_longer(
    cols = starts_with("estXY_"),
    names_to = "model",
    values_to = "estimate"
  ) %>%
  
  # per-run averages (per model x scenario)
  group_by(run, scenario, model) %>%
  summarise(
    mean_est = mean(estimate, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  
  # absolute bias relative to true parameter
  mutate(abs_bias = abs(mean_est - true_xy))

# boxplots: mean absolute bias across runs
plot_box_crosslag_X <- box_crosslag_X %>%
  ggplot(aes(x = model, y = abs_bias, fill = model)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.3) +
  
  facet_grid(rows = vars(scenario), cols = vars(model), scales = "free_y") +
  
  labs(
    x = "",
    y = "Mean absolute bias",
    title = "Cross-lag X on Y: mean absolute bias"
  ) +
  
  theme_minimal() +
  theme(
    strip.text = element_text(size = 10),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "none"
  )

# true cross-lag effect Y on X
true_yx <- A[1, 2]

# compute per-run average estimate and mean absolute bias
box_crosslag_Y <- df %>%
  
  # pull cross-lag Y estimates
  select(run, occasion, scenario,
         estYX_CLPM, estYX_RI_CLPM, estYX_DPM,
         estYX_CLPM_Adj, estYX_CLPM_LBCA) %>%
  
  pivot_longer(
    cols = starts_with("estYX_"),
    names_to = "model",
    values_to = "estimate"
  ) %>%
  
  group_by(run, scenario, model) %>%
  summarise(
    mean_est = mean(estimate, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  
  mutate(abs_bias = abs(mean_est - true_yx))

# boxplots
plot_box_crosslag_Y <- box_crosslag_Y %>%
  ggplot(aes(x = model, y = abs_bias, fill = model)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.3) +
  
  facet_grid(rows = vars(scenario), cols = vars(model), scales = "free_y") +
  
  labs(
    x = "",
    y = "Mean absolute bias",
    title = "Cross-lag Y on X: mean absolute bias"
  ) +
  
  theme_minimal() +
  theme(
    strip.text = element_text(size = 10),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "none"
  )

# true AR effect for X
true_ar_x <- A[1, 1]

box_ar_X <- df %>%
  select(run, occasion, scenario,
         estA_CLPM, estA_RI_CLPM, estA_DPM,
         estA_CLPM_Adj, estA_CLPM_LBCA) %>%
  
  pivot_longer(
    cols = starts_with("estA_"),
    names_to = "model",
    values_to = "estimate"
  ) %>%
  
  group_by(run, scenario, model) %>%
  summarise(
    mean_est = mean(estimate, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  
  mutate(abs_bias = abs(mean_est - true_ar_x))

plot_box_ar_X <- box_ar_X %>%
  ggplot(aes(x = model, y = abs_bias, fill = model)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.3) +
  
  facet_grid(rows = vars(scenario), cols = vars(model), scales = "free_y") +
  
  labs(
    x = "",
    y = "Mean absolute bias",
    title = "Autoregressive X: mean absolute bias"
  ) +
  
  theme_minimal() +
  theme(
    strip.text = element_text(size = 10),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "none"
  )

# true AR effect for Y
true_ar_y <- A[2, 2]

box_ar_Y <- df %>%
  select(run, occasion, scenario,
         estAY_CLPM, estAY_RI_CLPM, estAY_DPM,
         estAY_CLPM_Adj, estAY_CLPM_LBCA) %>%
  
  pivot_longer(
    cols = starts_with("estAY_"),
    names_to = "model",
    values_to = "estimate"
  ) %>%
  
  group_by(run, scenario, model) %>%
  summarise(
    mean_est = mean(estimate, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  
  mutate(abs_bias = abs(mean_est - true_ar_y))

plot_box_ar_Y <- box_ar_Y %>%
  ggplot(aes(x = model, y = abs_bias, fill = model)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.3) +
  
  facet_grid(rows = vars(scenario), cols = vars(model), scales = "free_y") +
  
  labs(
    x = "",
    y = "Mean absolute bias",
    title = "Autoregressive Y: mean absolute bias"
  ) +
  
  theme_minimal() +
  theme(
    strip.text = element_text(size = 10),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "none"
  )

###############################################################################
# displaying all plots
###############################################################################

# effect of confounder on X and Y over time
plot_beta_X
plot_beta_Y

# variance explained by confounder on X and Y over time
plot_R2_X
plot_R2_Y

# mean absolute bias of cross-lagged estimates over time 
plot_bias_XY
plot_bias_YX

# mean absolute bias of autoregressive estimates over time
plot_bias_AX
plot_bias_AY

# mean empirical SE of cross-lagged estimates over time
plot_se_crosslag_XY
plot_se_crosslag_YX

# mean empirical SE of autoregressive estimates over time
plot_se_AR_X
plot_se_AR_Y

# summary plot of mean absolute bias of cross-lagged estimates
plot_box_crosslag_X # (on Y)
plot_box_crosslag_Y # (on X)

# summary plot of mean absolute bias of autoregressive estimates
plot_box_ar_X
plot_box_ar_Y

```

Now lets proceed to simulation study 2. We will again source the required functions and run the simulation.

```{r}

# load the required functions and dependencies
source("Thesis\\Code\\Qmd_Rmd_R_Pyth\\Sim Studies\\Sim_Study_2_RR.R")

# autoregressive + cross-lag structure
A <- matrix(c(
  0.20, 0,                         # autoregressive and cross-lag for Y
  0.10, 0.20                       # cross-lag and autoregressive for X
), nrow = 2, byrow = TRUE)

# confounder covariance (3 confounders)
Psi <- diag(3)                     # uncorrelated confounders with var=1

# run the simulation
results_s2 <- run_simulation_study2(
  reps          = 1000,                        # number of replications
  N             = 1000,                        # sample size
  T             = 5,                           # number of waves
  k             = 3,                           # number of confounders
  R2_1          = 0.15,                        # confounder R2 at wave 1
  eta_1         = 0.5,                         # share of R2 due to non-linear confounding
  target_sd     = 0.5,                         # SD of B(t) variation
  scenarios     = c("constant", "stepwise"),   # scenarios to run
  A             = A,                           # 2×2 AR + cross-lag matrix
  Psi           = Psi,                         # confounder covariance
  rho_extra     = 0.1,                         # extra X–Y covariance
  models_to_run = c("clpm","riclpm",           # 5 modeling approaches 
                    "dpm","lbca","adj",
                    "xgb"),
  cores         = 6,                           # number of cores to use
  base_seed     = 12345                        # reproducible base seed
)

```

And now visualise the results for simulation study 2.

```{r}

df <- results_s2
str(df)

# custom colour palette
blue_magenta <- c(
  "#3B4CC0", "#5E60CE", "#7A70D1", "#9775D3", "#B67BD2",
  "#D084CF", "#E28FC4", "#F29BC1", "#FF9ACD"
)

# dxtend colours to number of runs
n_runs <- length(unique(df$run))
blue_magenta_extended <- rep(blue_magenta, length.out = n_runs)

###############################################################################
# plot 1: visualize how beta_X and beta_Y change over occasions
###############################################################################

# the effect of the confounder on X and Y over occasions, by scenario
plot_beta_X_s2 <- df %>%

  # the occasion on the X-axis, beta_X on the Y-axis, grouped and coloured by run
  # note that beta_X is the average over the 3 confounders
  ggplot(aes(x = occasion, y = beta_X,
             group = run, color = factor(run))) +

  # determines the look of the line
  geom_line(linewidth = 0.7, alpha = 0.9) +
  
  # separate plots for each scenario
  facet_wrap(~ scenario) +
  
  # use the custom colour palette
  scale_color_manual(values = blue_magenta_extended) +

  # remove the legend
  guides(color = "none") +

  # labels for the axes
  labs(x = "Occasion", y = "Effect of Confounder C on X") +

  # minimal theme for the plot
  theme_minimal()

# precisely the same for beta_Y
plot_beta_Y_s2 <- df %>%
  ggplot(aes(x = occasion, y = beta_Y,
             group = run, color = factor(run))) +
  geom_line(linewidth = 0.7, alpha = 0.9) +
  facet_wrap(~ scenario) +
  scale_color_manual(values = blue_magenta_extended) +
  guides(color = "none") +
  labs(x = "Occasion", y = "Effect of Confounder on Y") +
  theme_minimal()

###############################################################################
# plot 2: visualise how the explained variance changes over occasions
###############################################################################

# set equal to simulation parameter
eta_1 <- 0.50

plot_R2_X_s2 <- df %>%
  
  # compute total R2 at each occasion from time-varying beta
  mutate(
    R2_X_tot = beta_X^2,
    
    # split total R2 into linear and non-linear parts using eta_1
    R2_X_lin = (1 - eta_1) * R2_X_tot,
    R2_X_non = eta_1 * R2_X_tot
  ) %>%
  
  # reshape to long format for stacked plotting
  pivot_longer(
    cols = c(R2_X_lin, R2_X_non),
    names_to = "component",
    values_to = "R2_X"
  ) %>%
  
  # the occasion on the X-axis, R2_X on the Y-axis, grouped by run
  ggplot(aes(
    x = occasion,
    y = R2_X,
    group = interaction(run, component),
    fill  = component
  )) +

  # stacked area shows total magnitude and decomposition
  geom_area(alpha = 0.8, color = NA) +

  # separate plots for each scenario
  facet_wrap(~ scenario) +

  # use the custom colour palette
  scale_fill_manual(
    values = c(
      R2_X_lin = blue_magenta_extended[1],
      R2_X_non = blue_magenta_extended[3]
    ),
    labels = c(
      R2_X_lin = "Linear component",
      R2_X_non = "Non-linear component"
    )
  ) +

  # labels for the axes
  labs(
    x = "Occasion",
    y = "Explained Variance of Confounder C on X",
    fill = ""
  ) +

  # minimal theme for the plot
  theme_minimal() +
  theme(
    legend.position = "bottom"
  )


# same for R2_Y
plot_R2_Y_s2 <- df %>%
  
  # compute total R2 at each occasion from time-varying beta
  mutate(
    R2_Y_tot = beta_Y^2,
    
    # split total R2 into linear and non-linear parts using eta_1
    R2_Y_lin = (1 - eta_1) * R2_Y_tot,
    R2_Y_non = eta_1 * R2_Y_tot
  ) %>%
  
  # reshape to long format for stacked plotting
  pivot_longer(
    cols = c(R2_Y_lin, R2_Y_non),
    names_to = "component",
    values_to = "R2_Y"
  ) %>%
  
  ggplot(aes(
    x = occasion,
    y = R2_Y,
    group = interaction(run, component),
    fill  = component
  )) +
  geom_area(alpha = 0.8, color = NA) +
  facet_wrap(~ scenario) +
  scale_fill_manual(
    values = c(
      R2_Y_lin = blue_magenta_extended[1],
      R2_Y_non = blue_magenta_extended[3]
    ),
    labels = c(
      R2_Y_lin = "Linear component",
      R2_Y_non = "Non-linear component"
    )
  ) +
  labs(
    x = "Occasion",
    y = "Explained Variance of Confounder C on Y",
    fill = ""
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom"
  )

###############################################################################
# plot 3: show the mean absolute bias of the different models 
###############################################################################

# true effects extracted from A
true_xy <- A[2,1]   # XY (CL)
true_yx <- A[1,2]   # YX (CL)
true_ax <- A[1,1]   # X (AR)
true_ay <- A[2,2]   # Y (AR)

# cross-lagged estimates mean absolute bias X on Y 
bias_XY_s2 <- df %>%
  select(run, occasion, scenario,
         estXY_CLPM, estXY_RI_CLPM, estXY_DPM,
         estXY_CLPM_Adj, estXY_CLPM_LBCA, estXY_CLPM_XGB) %>%
  
  # long format
  pivot_longer(
    cols      = starts_with("estXY_"),
    names_to  = "model",
    values_to = "estimate"
  ) %>%
  
  # compute absolute bias
  mutate(abs_bias = abs(estimate - true_xy)) %>%
  
  # compute mean absolute bias per scenario × model × occasion
  group_by(scenario, model, occasion) %>%
  summarise(mean_abs_bias = mean(abs_bias, na.rm = TRUE),
            .groups = "drop")

# plot mean abs bias for XY
plot_bias_XY_s2 <- bias_XY_s2 %>%
  ggplot(aes(x = occasion, y = mean_abs_bias, color = model)) +
  geom_line(linewidth = 0.9) +
  facet_grid(rows = vars(scenario), cols = vars(model), scales = "free_y") +
  labs(x = "Occasion",
       y = "Mean Absolute Bias",
       title = "Bias: Cross-Lag XY") +
  theme_minimal() +
  theme(strip.text = element_text(size = 10),
        legend.position = "none")

# same but for YX
bias_YX_s2 <- df %>%
  select(run, occasion, scenario,
         estYX_CLPM, estYX_RI_CLPM, estYX_DPM,
         estYX_CLPM_Adj, estYX_CLPM_LBCA, estYX_CLPM_XGB) %>%
  
  pivot_longer(
    cols      = starts_with("estYX_"),
    names_to  = "model",
    values_to = "estimate"
  ) %>%
  
  mutate(abs_bias = abs(estimate - true_yx)) %>%
  
  group_by(scenario, model, occasion) %>%
  summarise(mean_abs_bias = mean(abs_bias, na.rm = TRUE),
            .groups = "drop")

plot_bias_YX_s2 <- bias_YX_s2 %>%
  ggplot(aes(x = occasion, y = mean_abs_bias, color = model)) +
  geom_line(linewidth = 0.9) +
  facet_grid(rows = vars(scenario), cols = vars(model), scales = "free_y") +
  labs(x = "Occasion",
       y = "Mean Absolute Bias",
       title = "Bias: Cross-Lag YX") +
  theme_minimal() +
  theme(strip.text = element_text(size = 10),
        legend.position = "none")

# compute mean absolute bias for AR(X)
bias_AX_s2 <- df %>%
  select(run, occasion, scenario,
         estA_CLPM, estA_RI_CLPM, estA_DPM,
         estA_CLPM_Adj, estA_CLPM_LBCA, estA_CLPM_XGB) %>%
  
  pivot_longer(
    cols      = starts_with("estA_"),
    names_to  = "model",
    values_to = "estimate"
  ) %>%
  
  mutate(abs_bias = abs(estimate - true_ax)) %>%
  
  group_by(scenario, model, occasion) %>%
  summarise(mean_abs_bias = mean(abs_bias, na.rm = TRUE),
            .groups = "drop")

plot_bias_AX_s2 <- bias_AX_s2 %>%
  ggplot(aes(x = occasion, y = mean_abs_bias, color = model)) +
  geom_line(linewidth = 0.9) +
  facet_grid(rows = vars(scenario),
             cols = vars(model),
             scales = "free_y") +
  labs(
    x = "Occasion",
    y = "Mean Absolute Bias",
    title = "Bias of AX Estimates"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 10),
    legend.position = "none"
  )

# compute mean absolute bias for AR(Y)
bias_AY_s2 <- df %>%
  select(run, occasion, scenario,
         estAY_CLPM, estAY_RI_CLPM, estAY_DPM,
         estAY_CLPM_Adj, estAY_CLPM_LBCA, estAY_CLPM_XGB) %>%
  
  pivot_longer(
    cols      = starts_with("estAY_"),
    names_to  = "model",
    values_to = "estimate"
  ) %>%
  
  mutate(abs_bias = abs(estimate - true_ay)) %>%
  
  group_by(scenario, model, occasion) %>%
  summarise(mean_abs_bias = mean(abs_bias, na.rm = TRUE),
            .groups = "drop")

plot_bias_AY_s2 <- bias_AY_s2 %>%
  ggplot(aes(x = occasion, y = mean_abs_bias, color = model)) +
  geom_line(linewidth = 0.9) +
  facet_grid(rows = vars(scenario),
             cols = vars(model),
             scales = "free_y") +
  labs(
    x = "Occasion",
    y = "Mean Absolute Bias",
    title = "Bias of AY Estimates"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 10),
    legend.position = "none"
  )

###############################################################################
# plot 4: the empirical standard error of the estimates over time
###############################################################################

# reshape and compute empirical standard error for cross-lag XY
se_crosslag_XY_s2 <- df %>%
  
  # select cross-lag XY estimates
  select(run, occasion, scenario,
         estXY_CLPM, estXY_RI_CLPM, estXY_DPM,
         estXY_CLPM_Adj, estXY_CLPM_LBCA, estXY_CLPM_XGB) %>%
  
  # pivot to long format
  pivot_longer(
    cols = starts_with("estXY_"),
    names_to = "model",
    values_to = "estimate"
  ) %>%
  
  # compute empirical SE: sd of estimates over runs
  group_by(scenario, model, occasion) %>%
  summarise(
    emp_se = sd(estimate, na.rm = TRUE),
    .groups = "drop"
  )

# plot empirical SE (scenarios = rows, methods = columns)
plot_se_crosslag_XY_s2 <- se_crosslag_XY_s2 %>%
  ggplot(aes(x = occasion, y = emp_se, color = model)) +
  geom_line(linewidth = 0.9) +
  facet_grid(rows = vars(scenario), cols = vars(model), scales = "free_y") +
  labs(
    x = "Occasion",
    y = "Empirical SE",
    title = "Empirical Standard Error: Cross-lag XY"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 10),
    legend.position = "none"
  )

# reshape and compute empirical standard error for cross-lag YX
se_crosslag_YX_s2 <- df %>%
  
  # select cross-lag YX estimates
  select(run, occasion, scenario,
         estYX_CLPM, estYX_RI_CLPM, estYX_DPM,
         estYX_CLPM_Adj, estYX_CLPM_LBCA, estYX_CLPM_XGB) %>%
  
  pivot_longer(
    cols = starts_with("estYX_"),
    names_to = "model",
    values_to = "estimate"
  ) %>%
  
  group_by(scenario, model, occasion) %>%
  summarise(
    emp_se = sd(estimate, na.rm = TRUE),
    .groups = "drop"
  )

plot_se_crosslag_YX_s2 <- se_crosslag_YX_s2 %>%
  ggplot(aes(x = occasion, y = emp_se, color = model)) +
  geom_line(linewidth = 0.9) +
  facet_grid(rows = vars(scenario), cols = vars(model), scales = "free_y") +
  labs(
    x = "Occasion",
    y = "Empirical SE",
    title = "Empirical Standard Error: Cross-lag YX"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 10),
    legend.position = "none"
  )

# reshape and compute empirical standard error for AR(X)
se_AR_X_s2 <- df %>%
  
  # select AR(X) estimates
  select(run, occasion, scenario,
         estA_CLPM, estA_RI_CLPM, estA_DPM,
         estA_CLPM_Adj, estA_CLPM_LBCA, estA_CLPM_XGB) %>%
  
  pivot_longer(
    cols = starts_with("estA_"),
    names_to = "model",
    values_to = "estimate"
  ) %>%
  
  group_by(scenario, model, occasion) %>%
  summarise(
    emp_se = sd(estimate, na.rm = TRUE),
    .groups = "drop"
  )

plot_se_AR_X_s2 <- se_AR_X_s2 %>%
  ggplot(aes(x = occasion, y = emp_se, color = model)) +
  geom_line(linewidth = 0.9) +
  facet_grid(rows = vars(scenario), cols = vars(model), scales = "free_y") +
  labs(
    x = "Occasion",
    y = "Empirical SE",
    title = "Empirical Standard Error: AX"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 10),
    legend.position = "none"
  )

# reshape and compute empirical standard error for AR(Y)
se_AR_Y_s2 <- df %>%
  
  # select AR(Y) estimates
  select(run, occasion, scenario,
         estAY_CLPM, estAY_RI_CLPM, estAY_DPM,
         estAY_CLPM_Adj, estAY_CLPM_LBCA, estAY_CLPM_XGB) %>%
  
  pivot_longer(
    cols = starts_with("estAY_"),
    names_to = "model",
    values_to = "estimate"
  ) %>%
  
  group_by(scenario, model, occasion) %>%
  summarise(
    emp_se = sd(estimate, na.rm = TRUE),
    .groups = "drop"
  )

plot_se_AR_Y_s2 <- se_AR_Y_s2 %>%
  ggplot(aes(x = occasion, y = emp_se, color = model)) +
  geom_line(linewidth = 0.9) +
  facet_grid(rows = vars(scenario), cols = vars(model), scales = "free_y") +
  labs(
    x = "Occasion",
    y = "Empirical SE",
    title = "Empirical Standard Error: AY"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 10),
    legend.position = "none"
  )

# plot 5: summary plot 
# true cross-lag effect X on Y
true_xy <- A[2, 1]

# compute per-run average estimate and mean absolute bias
box_crosslag_X_s2 <- df %>%
  
  # pull cross-lag X estimates
  select(run, occasion, scenario,
         estXY_CLPM, estXY_RI_CLPM, estXY_DPM,
         estXY_CLPM_Adj, estXY_CLPM_LBCA, estXY_CLPM_XGB) %>%
  
  # reshape to long format
  pivot_longer(
    cols = starts_with("estXY_"),
    names_to = "model",
    values_to = "estimate"
  ) %>%
  
  # per-run averages (per model x scenario)
  group_by(run, scenario, model) %>%
  summarise(
    mean_est = mean(estimate, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  
  # absolute bias relative to true parameter
  mutate(abs_bias = abs(mean_est - true_xy))

# boxplots: mean absolute bias across runs
plot_box_crosslag_X_s2 <- box_crosslag_X_s2 %>%
  ggplot(aes(x = model, y = abs_bias, fill = model)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.3) +
  
  facet_grid(rows = vars(scenario), cols = vars(model), scales = "free_y") +
  
  labs(
    x = "",
    y = "Mean absolute bias",
    title = "Cross-lag XY: mean absolute bias"
  ) +
  
  theme_minimal() +
  theme(
    strip.text = element_text(size = 10),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "none"
  )

# true cross-lag effect Y on X
true_yx <- A[1, 2]

# compute per-run average estimate and mean absolute bias
box_crosslag_Y_s2 <- df %>%
  
  # pull cross-lag Y estimates
  select(run, occasion, scenario,
         estYX_CLPM, estYX_RI_CLPM, estYX_DPM,
         estYX_CLPM_Adj, estYX_CLPM_LBCA, estYX_CLPM_XGB) %>%
  
  pivot_longer(
    cols = starts_with("estYX_"),
    names_to = "model",
    values_to = "estimate"
  ) %>%
  
  group_by(run, scenario, model) %>%
  summarise(
    mean_est = mean(estimate, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  
  mutate(abs_bias = abs(mean_est - true_yx))

# boxplots
plot_box_crosslag_Y_s2 <- box_crosslag_Y_s2 %>%
  ggplot(aes(x = model, y = abs_bias, fill = model)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.3) +
  
  facet_grid(rows = vars(scenario), cols = vars(model), scales = "free_y") +
  
  labs(
    x = "",
    y = "Mean absolute bias",
    title = "Cross-lag YX: mean absolute bias"
  ) +
  
  theme_minimal() +
  theme(
    strip.text = element_text(size = 10),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "none"
  )

# true AR effect for X
true_ar_x <- A[1, 1]

box_ar_X_s2 <- df %>%
  
  # pull AR(X) estimates
  select(run, occasion, scenario,
         estA_CLPM, estA_RI_CLPM, estA_DPM,
         estA_CLPM_Adj, estA_CLPM_LBCA, estA_CLPM_XGB) %>%
  
  pivot_longer(
    cols = starts_with("estA_"),
    names_to = "model",
    values_to = "estimate"
  ) %>%
  
  group_by(run, scenario, model) %>%
  summarise(
    mean_est = mean(estimate, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  
  mutate(abs_bias = abs(mean_est - true_ar_x))

plot_box_ar_X_s2 <- box_ar_X_s2 %>%
  ggplot(aes(x = model, y = abs_bias, fill = model)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.3) +
  
  facet_grid(rows = vars(scenario), cols = vars(model), scales = "free_y") +
  
  labs(
    x = "",
    y = "Mean absolute bias",
    title = "Autoregressive X: mean absolute bias"
  ) +
  
  theme_minimal() +
  theme(
    strip.text = element_text(size = 10),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "none"
  )

# true AR effect for Y
true_ar_y <- A[2, 2]

box_ar_Y_s2 <- df %>%
  
  # pull AR(Y) estimates
  select(run, occasion, scenario,
         estAY_CLPM, estAY_RI_CLPM, estAY_DPM,
         estAY_CLPM_Adj, estAY_CLPM_LBCA, estAY_CLPM_XGB) %>%
  
  pivot_longer(
    cols = starts_with("estAY_"),
    names_to = "model",
    values_to = "estimate"
  ) %>%
  
  group_by(run, scenario, model) %>%
  summarise(
    mean_est = mean(estimate, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  
  mutate(abs_bias = abs(mean_est - true_ar_y))

plot_box_ar_Y_s2 <- box_ar_Y_s2 %>%
  ggplot(aes(x = model, y = abs_bias, fill = model)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.3) +
  
  facet_grid(rows = vars(scenario), cols = vars(model), scales = "free_y") +
  
  labs(
    x = "",
    y = "Mean absolute bias",
    title = "Autoregressive Y: mean absolute bias"
  ) +
  
  theme_minimal() +
  theme(
    strip.text = element_text(size = 10),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "none"
  )

###############################################################################
# displaying all plots
###############################################################################

# effect of confounder on X and Y over time
plot_beta_X_s2
plot_beta_Y_s2

# variance explained by confounder on X and Y over time
plot_R2_X_s2
plot_R2_Y_s2

# mean absolute bias of cross-lagged estimates over time 
plot_bias_XY_s2
plot_bias_YX_s2

# mean absolute bias of autoregressive estimates over time
plot_bias_AX_s2
plot_bias_AY_s2

# mean empirical SE of cross-lagged estimates over time
plot_se_crosslag_XY_s2
plot_se_crosslag_YX_s2

# mean empirical SE of autoregressive estimates over time
plot_se_AR_X_s2
plot_se_AR_Y_s2

# summary plot of mean absolute bias of cross-lagged estimates
plot_box_crosslag_X_s2 # (on Y)
plot_box_crosslag_Y_s2 # (on X)

# summary plot of mean absolute bias of autoregressive estimates
plot_box_ar_X_s2
plot_box_ar_Y_s2


```

Extra bit where we finalize the two chosen plots for the research rapport

Simulation Study 1: Boxplots of Mean Absolute Relative Bias for Cross-Lagged Effect Y on X

```{r}

library(ggh4x)
library(colorspace)

df1 <- results_s1

# --- true cross-lag effect X -> Y ---
true_xy <- A[2, 1]

# --- build abs. relative bias data (per run x scenario x model) ---
box_crosslag_X_s1_absrel <- df1 %>%
  select(run, occasion, scenario,
         estXY_CLPM, estXY_CLPM_Adj, estXY_CLPM_LBCA,
         estXY_DPM, estXY_RI_CLPM) %>%
  pivot_longer(
    cols = starts_with("estXY_"),
    names_to = "model",
    values_to = "estimate"
  ) %>%
  group_by(run, scenario, model) %>%
  summarise(mean_est = mean(estimate, na.rm = TRUE), .groups = "drop") %>%
  mutate(
    abs_rel_bias = abs((mean_est - true_xy) / true_xy),

    model = recode(model,
      estXY_CLPM      = "CLPM unadjusted",
      estXY_CLPM_Adj  = "CLPM adjusted",
      estXY_CLPM_LBCA = "CLPM linear BCA",
      estXY_DPM       = "DPM",
      estXY_RI_CLPM   = "RI-CLPM"
    ),
    model = factor(model, levels = c(
      "CLPM unadjusted",
      "CLPM adjusted",
      "CLPM linear BCA",
      "DPM",
      "RI-CLPM"
    )),
    scenario = factor(scenario, levels = c("constant", "stepwise"))
  )

# --- runs per scenario (for facet strips) ---
runs_by_scenario <- box_crosslag_X_s1_absrel %>%
  summarise(n_runs = n_distinct(run), .by = scenario)

box_crosslag_X_s1_absrel <- box_crosslag_X_s1_absrel %>%
  left_join(runs_by_scenario, by = "scenario") %>%
  mutate(
    scenario_pretty = recode(as.character(scenario),
      constant = "Constant",
      stepwise = "Stepwise"
    ),
    scenario_facet = paste0("bold('", scenario_pretty, "')*' (runs = ", n_runs, ")'")
  )

# fixed y-axis across scenarios
y_lim <- range(box_crosslag_X_s1_absrel$abs_rel_bias, na.rm = TRUE)

# --- muted, colorblind-friendly palette (stable mapping to model levels) ---
pal <- colorspace::qualitative_hcl(n = nlevels(box_crosslag_X_s1_absrel$model), palette = "Set 2")
pal <- setNames(pal, levels(box_crosslag_X_s1_absrel$model))

# --- plot ---
plot_box_crosslag_X_s1_absrel <- ggplot(
  box_crosslag_X_s1_absrel,
  aes(x = model, y = abs_rel_bias, fill = model)
) +
  geom_boxplot(width = 0.55, alpha = 0.85, outlier.alpha = 0.3, linewidth = 0.35) +
  ggh4x::facet_wrap2(
    ~ scenario_facet,
    nrow = 1,
    labeller = label_parsed,
    axes = "y"   # repeat y-axis on each facet (axis between Constant and Stepwise)
  ) +
  coord_cartesian(ylim = y_lim) +
  scale_x_discrete(expand = expansion(mult = c(0.14, 0.03))) +
  scale_fill_manual(values = pal) +
  labs(
    x = "",
    y = "Mean absolute relative error"
  ) +
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text = element_text(size = 14),
    axis.text.x = element_text(
      size = 13, angle = 90, vjust = 0.5, hjust = 1,
      margin = margin(t = 8)
    ),
    axis.title.y = element_text(size = 13),
    axis.ticks.x = element_blank(),
    legend.position = "none",
    plot.margin = margin(t = 8, r = 8, b = 12, l = 10)
  )

plot_box_crosslag_X_s1_absrel

```

Simulation Study 2: Boxplots of Mean Absolute Relative Bias for Cross-Lagged Effect X on Y
```{r}

library(ggh4x)
library(colorspace)

df2 <- results_s2

# --- true cross-lag effect X -> Y ---
true_xy <- A[2, 1]

# --- build abs. relative bias data (per run x scenario x model) ---
box_crosslag_X_s2_absrel <- df2 %>%
  select(run, occasion, scenario,
         estXY_CLPM, estXY_CLPM_Adj, estXY_CLPM_LBCA, estXY_CLPM_XGB,
         estXY_DPM, estXY_RI_CLPM) %>%
  pivot_longer(
    cols = starts_with("estXY_"),
    names_to = "model",
    values_to = "estimate"
  ) %>%
  group_by(run, scenario, model) %>%
  summarise(mean_est = mean(estimate, na.rm = TRUE), .groups = "drop") %>%
  mutate(
    abs_rel_bias = abs((mean_est - true_xy) / true_xy),

    model = recode(model,
      estXY_CLPM      = "CLPM unadjusted",
      estXY_CLPM_Adj  = "CLPM adjusted",
      estXY_CLPM_LBCA = "CLPM linear BCA",
      estXY_CLPM_XGB  = "CLPM XGB BCA",
      estXY_DPM       = "DPM",
      estXY_RI_CLPM   = "RI-CLPM"
    ),
    model = factor(model, levels = c(
      "CLPM unadjusted",
      "CLPM adjusted",
      "CLPM linear BCA",
      "CLPM XGB BCA",
      "DPM",
      "RI-CLPM"
    )),
    scenario = factor(scenario, levels = c("constant", "stepwise"))
  )

# --- runs per scenario (for facet strips) ---
runs_by_scenario <- box_crosslag_X_s2_absrel %>%
  summarise(n_runs = n_distinct(run), .by = scenario)

box_crosslag_X_s2_absrel <- box_crosslag_X_s2_absrel %>%
  left_join(runs_by_scenario, by = "scenario") %>%
  mutate(
    scenario_pretty = recode(as.character(scenario),
      constant = "Constant",
      stepwise = "Stepwise"
    ),
    scenario_facet = paste0("bold('", scenario_pretty, "')*' (runs = ", n_runs, ")'")
  )

# fixed y-axis across scenarios
y_lim <- range(box_crosslag_X_s2_absrel$abs_rel_bias, na.rm = TRUE)

# --- muted, colorblind-friendly palette (stable mapping to model levels) ---
pal <- colorspace::qualitative_hcl(n = nlevels(box_crosslag_X_s2_absrel$model), palette = "Set 2")
pal <- setNames(pal, levels(box_crosslag_X_s2_absrel$model))

# --- plot ---
plot_box_crosslag_X_s2_absrel <- ggplot(
  box_crosslag_X_s2_absrel,
  aes(x = model, y = abs_rel_bias, fill = model)
) +
  geom_boxplot(width = 0.55, alpha = 0.85, outlier.alpha = 0.3, linewidth = 0.35) +
  ggh4x::facet_wrap2(
    ~ scenario_facet,
    nrow = 1,
    labeller = label_parsed,
    axes = "y"   # repeat y-axis on each facet (axis between Constant and Stepwise)
  ) +
  coord_cartesian(ylim = y_lim) +
  scale_x_discrete(expand = expansion(mult = c(0.14, 0.03))) +
  scale_fill_manual(values = pal) +
  labs(
    x = "",
    y = "Mean absolute relative error"
  ) +
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text = element_text(size = 14),
    axis.text.x = element_text(
      size = 13, angle = 90, vjust = 0.5, hjust = 1,
      margin = margin(t = 8)
    ),
    axis.title.y = element_text(size = 13),
    axis.ticks.x = element_blank(),
    legend.position = "none",
    plot.margin = margin(t = 8, r = 8, b = 12, l = 10)
  )

plot_box_crosslag_X_s2_absrel

```

saving our pictures: 

```{r}

ggsave(
  "C:/Users/Admin/Desktop/plot_crosslag_X_s1_absrel_HQ.png",
  plot = plot_box_crosslag_X_s1_absrel,
  width = 215, height = 150, units = "mm",
  dpi = 600,
  bg = "white"
)

ggsave(
  "C:/Users/Admin/Desktop/plot_crosslag_X_s2_absrel_HQ.png",
  plot = plot_box_crosslag_X_s2_absrel,
  width = 215, height = 150, units = "mm",
  dpi = 600,
  bg = "white"
)

```