---
title: "Research Rapport Results 2"
format: html
---

# Simulation Study 1 Results

This script provides the revised results for the research rapport simulations. Lets start by running simulation one from script.

### Helpers

```{r helpers, message=FALSE, warning=FALSE}

sample_B1 <- function(
  k,
  R2_1,
  eta_1 = 0.0,
  min_abs = 0.01,
  max_abs = 0.60,
  max_tries = 100000
) {

  if (eta_1 < 0 || eta_1 > 1)
    stop("eta_1 must be between 0 and 1")

  k_lin <- k
  k_non <- if (eta_1 > 0) {
    sum(sapply(2:k, function(m) choose(k, m)))
  } else 0

  R2_lin <- (1 - eta_1) * R2_1
  R2_non <- eta_1 * R2_1

  for (i in seq_len(max_tries)) {

    # linear effects
    bx_lin <- if (k_lin > 0 && R2_lin > 0) {
      v <- rnorm(k_lin)
      sqrt(R2_lin) * v / sqrt(sum(v^2))
    } else numeric(0)

    by_lin <- if (k_lin > 0 && R2_lin > 0) {
      v <- rnorm(k_lin)
      sqrt(R2_lin) * v / sqrt(sum(v^2))
    } else numeric(0)

    # non-linear effects
    bx_non <- by_non <- numeric(0)

    if (k_non > 0 && R2_non > 0) {
      v <- rnorm(k_non)
      bx_non <- sqrt(R2_non) * v / sqrt(sum(v^2))

      v <- rnorm(k_non)
      by_non <- sqrt(R2_non) * v / sqrt(sum(v^2))
    }

    bx <- c(bx_lin, bx_non)
    by <- c(by_lin, by_non)

    if (all(abs(c(bx, by)) >= min_abs & abs(c(bx, by)) <= max_abs)) {

      B1 <- rbind(bx, by)
      rownames(B1) <- c("X", "Y")

      # column names (MATCH simulator exactly)
      if (k_non == 0) {
        colnames(B1) <- paste0("c", 1:k)
      } else {
        int_names <- unlist(
          lapply(2:k, function(m)
            combn(1:k, m, FUN = function(ix)
              paste0("c", paste(ix, collapse=":")))
          )
        )
        colnames(B1) <- c(paste0("c", 1:k), int_names)
      }

      return(B1)
    }
  }

  stop("Failed to sample B1 within max_tries")
}

```

### Generating Results Simulation Study 1

```{r simulation study 1}

# load the function and dependencies
source("Thesis\\Code\\Qmd_Rmd_R_Pyth\\Sim Studies\\Sim_Study_1_RR_no_sampling.R")

# set a seed for reproducibility
set.seed(1234)

# autoregressive + cross-lag structure
A <- matrix(c(
  0.20, 0,                         # autoregressive and cross-lag for Y
  0.10, 0.20                       # cross-lag and autoregressive for X
), nrow = 2, byrow = TRUE)

# confounder covariance (3 confounders)
Psi <- diag(3)                     # uncorrelated confounders with var=1

# sample B1 matrix
B1 <- sample_B1(
  k = 3,
  R2_1 = 0.15,
  eta_1 = 0,
  min_abs = 0.001,
  max_abs = 0.40
)

# for the constant scenario, replicate B1 across time points
B_constant <- lapply(1:5, function(t) B1)

# scaling factor such that R_2 becomes twice as large: scaling factor = sqrt(old R2 / new R2)
scale_factor <- sqrt(0.40 / 0.15)              # = 1.414214
                                  
# create beta list for increasing R2 scenario
B_high <- B1 * scale_factor

# fill the list
B_stepwise <- list(
  t1 = B1,
  t2 = B1,
  t3 = B1,
  t4 = B_high,
  t5 = B_high
)

# run Simulation Study 1
results_sim1 <- run_simulation_study1(
  reps        = 1000,                                  # replications
  N           = 1000,                                  # sample size
  T           = 5,                                     # number of time points
  k           = 3,                                     # number of confounders
  scenarios   = c("constant", "stepwise"),             # B scenarios
  B_scenarios = list(                                  # beta trajectories
    constant = B_constant,
    stepwise = B_stepwise
  ),
  A           = A,                                     # autoregressive + cross-lag matrix
  Psi         = Psi,                                   # confounder covariance matrix
  rho_extra  = 0.1,                                    # extra correlation among X_t and Y_t
  models_to_run = c(                                   # models to run
    "clpm",
    "riclpm",
    "dpm",
    "adj",
    "lbca"
  ),
  cores       = 6,                                     # number of cores for parallel processing
  base_seed  = 12345                                   # base seed for reproducibility
)

```

### Plot 0.1: Beta trajectory plots Simulation Study 1

This plot shows how the confounder effects (betas) change over time in Simulation Study 1. Each line represents one confounder. In the constant scenario the effects stay the same across occasions, while in the stepwise scenario they increase at later occasions by design.

```{r}

# libraries
library(tidyverse)
library(ggh4x)
library(colorspace)

# first load the data
df1 <- results_sim1 

# how many runs were not successful?
sum(df1$is_na_run)

# custom colour palette (scientific, colorblind-friendly)
pal_beta <- qualitative_hcl(
  n = ncol(B1),
  palette = "Set 2"
)

# number of occasions
T <- length(B_stepwise)              # = 5

# number of confounders
k <- ncol(B1)                        # = 3

# build dataframe for beta trajectories
df_beta <- bind_rows(

  # constant betas (same B at each occasion)
  lapply(1:T, function(t) {
    tibble(
      occasion   = t,                        # time point
      confounder = factor(1:k),               # confounder index
      beta_X     = B_constant[[t]][1, ],      # effect of C on X
      beta_Y     = B_constant[[t]][2, ],      # effect of C on Y
      scenario   = "constant"                 # scenario label
    )
  }),

  # stepwise betas (increase at later occasions)
  lapply(1:T, function(t) {
    tibble(
      occasion   = t,                        # time point
      confounder = factor(1:k),               # confounder index
      beta_X     = B_stepwise[[t]][1, ],      # effect of C on X
      beta_Y     = B_stepwise[[t]][2, ],      # effect of C on Y
      scenario   = "stepwise"                 # scenario label
    )
  })
)

# plot the effect of the confounders on X over time
plot_beta_X <- df_beta %>%

  # ensure ordered labels for legend
  mutate(confounder = factor(
    confounder,
    levels = 1:k,
    labels = paste0("C", 1:k)
  )) %>%

  # occasion on X-axis, beta_X on Y-axis
  # one line per confounder
  ggplot(aes(
    x     = occasion,
    y     = beta_X,
    group = confounder,
    color = confounder
  )) +

  # draw beta trajectories
  geom_line(linewidth = 0.9, alpha = 0.9) +

  # separate plots for each scenario
  ggh4x::facet_wrap2(
    ~ scenario,
    nrow = 1,
    axes = "y"
  ) +

  # use scientific colour palette
  scale_color_manual(
    values = pal_beta,
    guide = guide_legend(nrow = 1)
  ) +

  # axis labels
  labs(
    x = "Occasion",
    y = "Effect of Confounder on X",
    color = NULL
  ) +

  # scientific theme
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12),
    plot.margin     = margin(t = 8, r = 8, b = 12, l = 10)
  )


# plot the effect of the confounders on Y over time
plot_beta_Y <- df_beta %>%

  # ensure ordered labels for legend
  mutate(confounder = factor(
    confounder,
    levels = 1:k,
    labels = paste0("C", 1:k)
  )) %>%

  # occasion on X-axis, beta_Y on Y-axis
  # one line per confounder
  ggplot(aes(
    x     = occasion,
    y     = beta_Y,
    group = confounder,
    color = confounder
  )) +

  # draw beta trajectories
  geom_line(linewidth = 0.9, alpha = 0.9) +

  # separate plots for each scenario
  ggh4x::facet_wrap2(
    ~ scenario,
    nrow = 1,
    axes = "y"
  ) +

  # use scientific colour palette
  scale_color_manual(
    values = pal_beta,
    guide = guide_legend(nrow = 1)
  ) +

  # axis labels
  labs(
    x = "Occasion",
    y = "Effect of Confounder on Y",
    color = NULL
  ) +

  # scientific theme
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12),
    plot.margin     = margin(t = 8, r = 8, b = 12, l = 10)
  )

# display the plots
plot_beta_X
plot_beta_Y

```

### Plot 0.2: R squared trajectory plots of the confounders Simulation Study 1

This plot shows how the explained variance (R²) of the confounders changes over time in Simulation Study 1. Each line represents one confounder, and the total line shows their combined explained variance. In the constant scenario the explained variance stays stable across occasions, while in the stepwise scenario it increases at later occasions by design.

```{r R squared plots}

# name the palette to match confounder labels
pal_beta_named <- setNames(pal_beta, paste0("C", 1:k))

# plot the R^2 of confounders on X over time
plot_R2_X <- df_beta %>%

  # compute R^2 and relabel confounders
  mutate(
    R2 = beta_X^2,
    confounder = factor(
      confounder,
      levels = 1:k,
      labels = paste0("C", 1:k)
    )
  ) %>%

  # add total R^2 per occasion and scenario
  bind_rows(
    df_beta %>%
      mutate(R2 = beta_X^2) %>%
      group_by(occasion, scenario) %>%
      summarise(
        R2 = sum(R2),
        confounder = "Total",
        .groups = "drop"
      )
  ) %>%

  # occasion on X-axis, R^2 on Y-axis
  ggplot(aes(
    x     = occasion,
    y     = R2,
    group = confounder,
    color = confounder
  )) +

  # draw R^2 trajectories
  geom_line(linewidth = 0.9, alpha = 0.9) +

  # separate plots for each scenario
  ggh4x::facet_wrap2(
    ~ scenario,
    nrow = 1,
    axes = "y"
  ) +

  # correctly mapped colours + legend order
  scale_color_manual(
    values = c(pal_beta_named, Total = "black"),
    breaks = c(paste0("C", 1:k), "Total"),
    guide  = guide_legend(nrow = 1)
  ) +

  # axis labels
  labs(
    x = "Occasion",
    y = "Explained variance (R^2) of confounders on X",
    color = NULL
  ) +

  # scientific theme
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12),
    plot.margin     = margin(t = 8, r = 8, b = 12, l = 10)
  )


# plot the R^2 of confounders on Y over time
plot_R2_Y <- df_beta %>%

  # compute R^2 and relabel confounders
  mutate(
    R2 = beta_Y^2,
    confounder = factor(
      confounder,
      levels = 1:k,
      labels = paste0("C", 1:k)
    )
  ) %>%

  # add total R^2 per occasion and scenario
  bind_rows(
    df_beta %>%
      mutate(R2 = beta_Y^2) %>%
      group_by(occasion, scenario) %>%
      summarise(
        R2 = sum(R2),
        confounder = "Total",
        .groups = "drop"
      )
  ) %>%

  # occasion on X-axis, R^2 on Y-axis
  ggplot(aes(
    x     = occasion,
    y     = R2,
    group = confounder,
    color = confounder
  )) +

  # draw R^2 trajectories
  geom_line(linewidth = 0.9, alpha = 0.9) +

  # separate plots for each scenario
  ggh4x::facet_wrap2(
    ~ scenario,
    nrow = 1,
    axes = "y"
  ) +

  # correctly mapped colours + legend order
  scale_color_manual(
    values = c(pal_beta_named, Total = "black"),
    breaks = c(paste0("C", 1:k), "Total"),
    guide  = guide_legend(nrow = 1)
  ) +

  # axis labels
  labs(
    x = "Occasion",
    y = "Explained variance (R^2) of confounders on Y",
    color = NULL
  ) +

  # scientific theme
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12),
    plot.margin     = margin(t = 8, r = 8, b = 12, l = 10)
  )

# display the plots
plot_R2_X
plot_R2_Y

```

### Plot 1.1: Relative Bias Simulation Study 1

These plots show the relative bias of the estimated parameters across occasions in Simulation Study 1. Relative bias is computed as the difference between the estimated and true parameter value, divided by the true value. Because the true cross-lagged effect of Y on X equals zero, relative bias is not defined for this parameter and is therefore omitted from the figure.

```{r relative bias plots}

library(tidyverse)
library(ggh4x)
library(viridis)

#####################################################
# Build relative bias data for all lagged parameters
#####################################################

# true parameters from the data-generating model
true_params <- tibble(
  estimand = c("XY", "YX", "AX", "AY"),
  true     = c(
    A[2, 1],  # XY
    A[1, 2],  # YX
    A[1, 1],  # AX
    A[2, 2]   # AY
  )
)

# build relative bias data frame
relbias_df <- results_sim1 %>%

  # only occasions with lagged parameters
  filter(occasion >= 2) %>%

  # reshape estimates to long format
  pivot_longer(
    cols = matches("^est(XY|YX|A|AY)_"),
    names_to  = "param",
    values_to = "estimate"
  ) %>%

  # drop failed fits
  filter(!is.na(estimate)) %>%

  # identify estimand type and method
  mutate(
    estimand = case_when(
      str_detect(param, "^estXY_") ~ "XY",
      str_detect(param, "^estYX_") ~ "YX",
      str_detect(param, "^estA_")  ~ "AX",
      str_detect(param, "^estAY_") ~ "AY"
    ),

    method = param %>%
      str_remove("^est(XY|YX|A|AY)_") %>%
      recode(
        CLPM      = "CLPM",
        CLPM_Adj  = "True model",
        CLPM_LBCA = "CLPM linear BCA",
        DPM       = "DPM",
        RI_CLPM   = "RI-CLPM"
      )
  ) %>%

  # attach true parameter values
  left_join(true_params, by = "estimand") %>%

  # compute relative bias and Monte Carlo SE
  group_by(scenario, occasion, estimand, method) %>%
  reframe(
    nsim = n(),

    mean_est = mean(estimate),

    # relative bias
    rel_bias = (mean_est - true) / true,

    # Monte Carlo SE of relative bias
    mcse_rel_bias =
      sqrt(
        sum((estimate - mean_est)^2) /
        (nsim * (nsim - 1))
      ) / abs(true)
  ) %>%

  # set clean factor ordering
  mutate(
    scenario = factor(scenario, levels = c("constant", "stepwise")),
    method   = factor(
      method,
      levels = c(
        "CLPM",
        "True model",
        "CLPM linear BCA",
        "DPM",
        "RI-CLPM"
      )
    )
  )

#####################################################
# Colour palette for methods (shared across plots)
#####################################################

pal_method <- viridis(
  n = nlevels(relbias_df$method),
  option = "viridis",
  begin  = 0.1,
  end    = 0.9
)
names(pal_method) <- levels(relbias_df$method)

#####################################################
# Plot 1: cross-lagged effect of X on Y (XY)
#####################################################

plot_relbias_XY <- relbias_df %>%
  filter(estimand == "XY") %>%

  ggplot(aes(
    x     = occasion,
    y     = rel_bias,
    color = method,
    group = method
  )) +

  # relative bias trajectories
  geom_line(linewidth = 0.9) +

  # add points at each occasion
  geom_point(size = 2) +

  # Monte Carlo SE error bars
  geom_errorbar(
    aes(
      ymin = rel_bias - mcse_rel_bias,
      ymax = rel_bias + mcse_rel_bias
    ),
    width = 0.15,
    linewidth = 0.4
  ) +

  # zero-bias reference line
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.4) +

  # facet by scenario
  ggh4x::facet_wrap2(
    ~ scenario,
    nrow = 1,
    axes = "y"
  ) +

  # colours
  scale_color_manual(values = pal_method) +

  # labels
  labs(
    x = "Occasion",
    y = "Relative bias",
    color = NULL
  ) +

  # scientific theme
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12)
  )

#####################################################
# Plot 2: autoregressive effect of X on X (AX)
#####################################################

plot_relbias_AX <- relbias_df %>%
  filter(estimand == "AX") %>%

  ggplot(aes(
    x     = occasion,
    y     = rel_bias,
    color = method,
    group = method
  )) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  geom_errorbar(
    aes(
      ymin = rel_bias - mcse_rel_bias,
      ymax = rel_bias + mcse_rel_bias
    ),
    width = 0.15,
    linewidth = 0.4
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.4) +
  ggh4x::facet_wrap2(~ scenario, nrow = 1, axes = "y") +
  scale_color_manual(values = pal_method) +
  labs(x = "Occasion", y = "Relative bias", color = NULL) +
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12)
  )

#####################################################
# Plot 3: autoregressive effect of Y on Y (AY)
#####################################################

plot_relbias_AY <- relbias_df %>%
  filter(estimand == "AY") %>%

  ggplot(aes(
    x     = occasion,
    y     = rel_bias,
    color = method,
    group = method
  )) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  geom_errorbar(
    aes(
      ymin = rel_bias - mcse_rel_bias,
      ymax = rel_bias + mcse_rel_bias
    ),
    width = 0.15,
    linewidth = 0.4
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.4) +
  ggh4x::facet_wrap2(~ scenario, nrow = 1, axes = "y") +
  scale_color_manual(values = pal_method) +
  labs(x = "Occasion", y = "Relative bias", color = NULL) +
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12)
  )

#####################################################
# Display all plots
#####################################################

plot_relbias_XY
plot_relbias_AX
plot_relbias_AY

```

### Plot 1.2: Bias Simulation Study 1

These plots show the bias of the estimated parameters across occasions in Simulation Study 1. Bias is computed as the difference between the estimated and true parameter value and illustrates whether, and in which direction, the estimators systematically deviate from the data-generating values over time.

```{r}

#####################################################
# Build bias data for all lagged parameters
#####################################################

# true parameters from the data-generating model
true_params <- tibble(
  estimand = c("XY", "YX", "AX", "AY"),
  true     = c(
    A[2, 1],  # XY
    A[1, 2],  # YX (true = 0)
    A[1, 1],  # AX
    A[2, 2]   # AY
  )
)

# build bias data frame
bias_df <- results_sim1 %>%

  # only occasions with lagged parameters
  filter(occasion >= 2) %>%

  # reshape estimates to long format
  pivot_longer(
    cols = matches("^est(XY|YX|A|AY)_"),
    names_to  = "param",
    values_to = "estimate"
  ) %>%

  # drop failed fits
  filter(!is.na(estimate)) %>%

  # identify estimand type and method
  mutate(
    estimand = case_when(
      str_detect(param, "^estXY_") ~ "XY",
      str_detect(param, "^estYX_") ~ "YX",
      str_detect(param, "^estA_")  ~ "AX",
      str_detect(param, "^estAY_") ~ "AY"
    ),

    method = param %>%
      str_remove("^est(XY|YX|A|AY)_") %>%
      recode(
        CLPM      = "CLPM",
        CLPM_Adj  = "True model",
        CLPM_LBCA = "CLPM linear BCA",
        DPM       = "DPM",
        RI_CLPM   = "RI-CLPM"
      )
  ) %>%

  # attach true parameter values
  left_join(true_params, by = "estimand") %>%

  # compute bias and Monte Carlo SE
  group_by(scenario, occasion, estimand, method) %>%
  reframe(
    nsim = n(),

    mean_est = mean(estimate),

    # bias
    bias = mean_est - true,

    # Monte Carlo SE of the bias
    mcse_bias =
      sqrt(
        sum((estimate - mean_est)^2) /
        (nsim * (nsim - 1))
      )
  ) %>%

  # clean factor ordering
  mutate(
    scenario = factor(scenario, levels = c("constant", "stepwise")),
    method   = factor(
      method,
      levels = c(
        "CLPM",
        "True model",
        "CLPM linear BCA",
        "DPM",
        "RI-CLPM"
      )
    )
  )

#####################################################
# Colour palette for methods (shared across plots)
#####################################################

pal_method <- qualitative_hcl(
  n = nlevels(bias_df$method),
  palette = "Set 2"
)
names(pal_method) <- levels(bias_df$method)

#####################################################
# Plot 1: cross-lagged effect of X on Y (XY)
#####################################################

plot_bias_XY <- bias_df %>%
  filter(estimand == "XY") %>%

  ggplot(aes(
    x     = occasion,
    y     = bias,
    color = method,
    group = method
  )) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  geom_errorbar(
    aes(
      ymin = bias - mcse_bias,
      ymax = bias + mcse_bias
    ),
    width = 0.15,
    linewidth = 0.4
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.4) +
  ggh4x::facet_wrap2(~ scenario, nrow = 1, axes = "y") +
  scale_color_manual(values = pal_method) +
  labs(
    x = "Occasion",
    y = "Bias",
    color = NULL
  ) +
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12)
  )

#####################################################
# Plot 2: cross-lagged effect of Y on X (YX)
#####################################################

plot_bias_YX <- bias_df %>%
  filter(estimand == "YX") %>%

  ggplot(aes(
    x     = occasion,
    y     = bias,
    color = method,
    group = method
  )) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  geom_errorbar(
    aes(
      ymin = bias - mcse_bias,
      ymax = bias + mcse_bias
    ),
    width = 0.15,
    linewidth = 0.4
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.4) +
  ggh4x::facet_wrap2(~ scenario, nrow = 1, axes = "y") +
  scale_color_manual(values = pal_method) +
  labs(
    x = "Occasion",
    y = "Bias",
    color = NULL
  ) +
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12)
  )

#####################################################
# Plot 3: autoregressive effect of X on X (AX)
#####################################################

plot_bias_AX <- bias_df %>%
  filter(estimand == "AX") %>%

  ggplot(aes(
    x     = occasion,
    y     = bias,
    color = method,
    group = method
  )) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  geom_errorbar(
    aes(
      ymin = bias - mcse_bias,
      ymax = bias + mcse_bias
    ),
    width = 0.15,
    linewidth = 0.4
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.4) +
  ggh4x::facet_wrap2(~ scenario, nrow = 1, axes = "y") +
  scale_color_manual(values = pal_method) +
  labs(
    x = "Occasion",
    y = "Bias",
    color = NULL
  ) +
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12)
  )

#####################################################
# Plot 4: autoregressive effect of Y on Y (AY)
#####################################################

plot_bias_AY <- bias_df %>%
  filter(estimand == "AY") %>%

  ggplot(aes(
    x     = occasion,
    y     = bias,
    color = method,
    group = method
  )) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  geom_errorbar(
    aes(
      ymin = bias - mcse_bias,
      ymax = bias + mcse_bias
    ),
    width = 0.15,
    linewidth = 0.4
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.4) +
  ggh4x::facet_wrap2(~ scenario, nrow = 1, axes = "y") +
  scale_color_manual(values = pal_method) +
  labs(
    x = "Occasion",
    y = "Bias",
    color = NULL
  ) +
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12)
  )

#####################################################
# Display all plots
#####################################################

plot_bias_XY
plot_bias_YX
plot_bias_AX
plot_bias_AY

```

### Plot 2: RMSE Simulation Study 1

These plots show the root mean squared error (RMSE) of the estimated parameters across occasions in Simulation Study 1. RMSE combines both bias and variability, providing an overall measure of estimation accuracy over time for each method.

```{r RMSE plots}
library(tidyverse)
library(ggh4x)
library(viridis)

#####################################################
# Build RMSE data for all lagged parameters
#####################################################

# true parameters from the data-generating model
true_params <- tibble(
  estimand = c("XY", "YX", "AX", "AY"),
  true     = c(
    A[2, 1],  # XY
    A[1, 2],  # YX (true = 0)
    A[1, 1],  # AX
    A[2, 2]   # AY
  )
)

# build RMSE data frame
rmse_df <- results_sim1 %>%

  # only occasions with lagged parameters
  filter(occasion >= 2) %>%

  # reshape estimates to long format
  pivot_longer(
    cols = matches("^est(XY|YX|A|AY)_"),
    names_to  = "param",
    values_to = "estimate"
  ) %>%

  # drop failed fits
  filter(!is.na(estimate)) %>%

  # identify estimand type and method
  mutate(
    estimand = case_when(
      str_detect(param, "^estXY_") ~ "XY",
      str_detect(param, "^estYX_") ~ "YX",
      str_detect(param, "^estA_")  ~ "AX",
      str_detect(param, "^estAY_") ~ "AY"
    ),

    method = param %>%
      str_remove("^est(XY|YX|A|AY)_") %>%
      recode(
        CLPM      = "CLPM",
        CLPM_Adj  = "True model",
        CLPM_LBCA = "CLPM linear BCA",
        DPM       = "DPM",
        RI_CLPM   = "RI-CLPM"
      )
  ) %>%

  # attach true parameter values
  left_join(true_params, by = "estimand") %>%

  # compute RMSE and Monte Carlo SE
  group_by(scenario, occasion, estimand, method) %>%
  reframe(
    nsim = n(),

    # RMSE
    rmse = sqrt(mean((estimate - true)^2)),

    # Monte Carlo SE of RMSE
    mcse_rmse =
      sqrt(
        sum(
          ((estimate - true)^2 - mean((estimate - true)^2))^2
        ) /
        (nsim * (nsim - 1))
      ) / (2 * rmse)
  ) %>%

  # clean factor ordering
  mutate(
    scenario = factor(scenario, levels = c("constant", "stepwise")),
    method   = factor(
      method,
      levels = c(
        "CLPM",
        "True model",
        "CLPM linear BCA",
        "DPM",
        "RI-CLPM"
      )
    )
  )

#####################################################
# Colour palette for methods (shared across plots)
#####################################################

pal_method <- viridis(
  n = nlevels(rmse_df$method),
  option = "viridis",
  begin  = 0.1,
  end    = 0.9
)
names(pal_method) <- levels(rmse_df$method)

#####################################################
# Plot 1: cross-lagged effect of X on Y (XY)
#####################################################

plot_rmse_XY <- rmse_df %>%
  filter(estimand == "XY") %>%

  ggplot(aes(
    x     = occasion,
    y     = rmse,
    color = method,
    group = method
  )) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  geom_errorbar(
    aes(
      ymin = rmse - mcse_rmse,
      ymax = rmse + mcse_rmse
    ),
    width = 0.15,
    linewidth = 0.4
  ) +
  ggh4x::facet_wrap2(~ scenario, nrow = 1, axes = "y") +
  scale_color_manual(values = pal_method) +
  labs(
    x = "Occasion",
    y = "RMSE",
    color = NULL
  ) +
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12)
  )

#####################################################
# Plot 2: cross-lagged effect of Y on X (YX)
#####################################################

plot_rmse_YX <- rmse_df %>%
  filter(estimand == "YX") %>%

  ggplot(aes(
    x     = occasion,
    y     = rmse,
    color = method,
    group = method
  )) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  geom_errorbar(
    aes(
      ymin = rmse - mcse_rmse,
      ymax = rmse + mcse_rmse
    ),
    width = 0.15,
    linewidth = 0.4
  ) +
  ggh4x::facet_wrap2(~ scenario, nrow = 1, axes = "y") +
  scale_color_manual(values = pal_method) +
  labs(
    x = "Occasion",
    y = "RMSE",
    color = NULL
  ) +
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12)
  )

#####################################################
# Plot 3: autoregressive effect of X on X (AX)
#####################################################

plot_rmse_AX <- rmse_df %>%
  filter(estimand == "AX") %>%

  ggplot(aes(
    x     = occasion,
    y     = rmse,
    color = method,
    group = method
  )) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  geom_errorbar(
    aes(
      ymin = rmse - mcse_rmse,
      ymax = rmse + mcse_rmse
    ),
    width = 0.15,
    linewidth = 0.4
  ) +
  ggh4x::facet_wrap2(~ scenario, nrow = 1, axes = "y") +
  scale_color_manual(values = pal_method) +
  labs(
    x = "Occasion",
    y = "RMSE",
    color = NULL
  ) +
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12)
  )

#####################################################
# Plot 4: autoregressive effect of Y on Y (AY)
#####################################################

plot_rmse_AY <- rmse_df %>%
  filter(estimand == "AY") %>%

  ggplot(aes(
    x     = occasion,
    y     = rmse,
    color = method,
    group = method
  )) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  geom_errorbar(
    aes(
      ymin = rmse - mcse_rmse,
      ymax = rmse + mcse_rmse
    ),
    width = 0.15,
    linewidth = 0.4
  ) +
  ggh4x::facet_wrap2(~ scenario, nrow = 1, axes = "y") +
  scale_color_manual(values = pal_method) +
  labs(
    x = "Occasion",
    y = "RMSE",
    color = NULL
  ) +
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12)
  )

#####################################################
# Display all plots
#####################################################

plot_rmse_XY
plot_rmse_YX
plot_rmse_AX
plot_rmse_AY

```

### Plot 3: Empirical SE Simulation Study 1

These plots show the empirical standard error of the estimated parameters across occasions in Simulation Study 1. The empirical SE reflects the variability of the estimates across simulation runs and captures sampling uncertainty independent of bias.

```{r Empirical SE plots}

#####################################################
# Build empirical SE data for all lagged parameters
#####################################################

# build empirical SE data frame
empse_df <- results_sim1 %>%

  # only occasions with lagged parameters
  filter(occasion >= 2) %>%

  # reshape estimates to long format
  pivot_longer(
    cols = matches("^est(XY|YX|A|AY)_"),
    names_to  = "param",
    values_to = "estimate"
  ) %>%

  # drop failed fits
  filter(!is.na(estimate)) %>%

  # identify estimand type and method
  mutate(
    estimand = case_when(
      str_detect(param, "^estXY_") ~ "XY",
      str_detect(param, "^estYX_") ~ "YX",
      str_detect(param, "^estA_")  ~ "AX",
      str_detect(param, "^estAY_") ~ "AY"
    ),

    method = param %>%
      str_remove("^est(XY|YX|A|AY)_") %>%
      recode(
        CLPM      = "CLPM",
        CLPM_Adj  = "True model",
        CLPM_LBCA = "CLPM linear BCA",
        DPM       = "DPM",
        RI_CLPM   = "RI-CLPM"
      )
  ) %>%

  # compute empirical SE and Monte Carlo SE
  group_by(scenario, occasion, estimand, method) %>%
  reframe(
    nsim = n(),

    # empirical standard error
    emp_se = sd(estimate),

    # Monte Carlo SE of empirical SE
    mcse_emp_se =
      emp_se / sqrt(2 * (nsim - 1))
  ) %>%

  # clean factor ordering
  mutate(
    scenario = factor(scenario, levels = c("constant", "stepwise")),
    method   = factor(
      method,
      levels = c(
        "CLPM",
        "True model",
        "CLPM linear BCA",
        "DPM",
        "RI-CLPM"
      )
    )
  )

#####################################################
# Colour palette for methods (shared across plots)
#####################################################

pal_method <- qualitative_hcl(
  n = nlevels(empse_df$method),
  palette = "Set 2"
)
names(pal_method) <- levels(empse_df$method)

#####################################################
# Plot 1: cross-lagged effect of X on Y (XY)
#####################################################

plot_empse_XY <- empse_df %>%
  filter(estimand == "XY") %>%

  ggplot(aes(
    x     = occasion,
    y     = emp_se,
    color = method,
    group = method
  )) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  geom_errorbar(
    aes(
      ymin = emp_se - mcse_emp_se,
      ymax = emp_se + mcse_emp_se
    ),
    width = 0.15,
    linewidth = 0.4
  ) +
  ggh4x::facet_wrap2(~ scenario, nrow = 1, axes = "y") +
  scale_color_manual(values = pal_method) +
  labs(
    x = "Occasion",
    y = "Empirical SE",
    color = NULL
  ) +
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12)
  )

#####################################################
# Plot 2: cross-lagged effect of Y on X (YX)
#####################################################

plot_empse_YX <- empse_df %>%
  filter(estimand == "YX") %>%

  ggplot(aes(
    x     = occasion,
    y     = emp_se,
    color = method,
    group = method
  )) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  geom_errorbar(
    aes(
      ymin = emp_se - mcse_emp_se,
      ymax = emp_se + mcse_emp_se
    ),
    width = 0.15,
    linewidth = 0.4
  ) +
  ggh4x::facet_wrap2(~ scenario, nrow = 1, axes = "y") +
  scale_color_manual(values = pal_method) +
  labs(
    x = "Occasion",
    y = "Empirical SE",
    color = NULL
  ) +
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12)
  )

#####################################################
# Plot 3: autoregressive effect of X on X (AX)
#####################################################

plot_empse_AX <- empse_df %>%
  filter(estimand == "AX") %>%

  ggplot(aes(
    x     = occasion,
    y     = emp_se,
    color = method,
    group = method
  )) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  geom_errorbar(
    aes(
      ymin = emp_se - mcse_emp_se,
      ymax = emp_se + mcse_emp_se
    ),
    width = 0.15,
    linewidth = 0.4
  ) +
  ggh4x::facet_wrap2(~ scenario, nrow = 1, axes = "y") +
  scale_color_manual(values = pal_method) +
  labs(
    x = "Occasion",
    y = "Empirical SE",
    color = NULL
  ) +
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12)
  )

#####################################################
# Plot 4: autoregressive effect of Y on Y (AY)
#####################################################

plot_empse_AY <- empse_df %>%
  filter(estimand == "AY") %>%

  ggplot(aes(
    x     = occasion,
    y     = emp_se,
    color = method,
    group = method
  )) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  geom_errorbar(
    aes(
      ymin = emp_se - mcse_emp_se,
      ymax = emp_se + mcse_emp_se
    ),
    width = 0.15,
    linewidth = 0.4
  ) +
  ggh4x::facet_wrap2(~ scenario, nrow = 1, axes = "y") +
  scale_color_manual(values = pal_method) +
  labs(
    x = "Occasion",
    y = "Empirical SE",
    color = NULL
  ) +
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12)
  )

#####################################################
# Display all plots
#####################################################

plot_empse_XY
plot_empse_YX
plot_empse_AX
plot_empse_AY

```

### Plot 4.1 : Coverage Simulation Study 1

These plots show the coverage of the nominal 95% confidence intervals across occasions in Simulation Study 1. Coverage represents the proportion of simulation runs in which the true parameter value falls within the estimated confidence interval, indicating how well uncertainty is quantified by each method

```{r Coverage plots}

#####################################################
# Build coverage data for all lagged parameters
#####################################################

# true parameters from the data-generating model
true_params <- tibble(
  estimand = c("XY", "YX", "AX", "AY"),
  true     = c(
    A[2, 1],  # XY
    A[1, 2],  # YX
    A[1, 1],  # AX
    A[2, 2]   # AY
  )
)

# --- estimates ---
est_df <- results_sim1 %>%

  # only occasions with lagged parameters
  filter(occasion >= 2) %>%

  pivot_longer(
    cols = matches("^est(XY|YX|A|AY)_"),
    names_to  = "param",
    values_to = "estimate"
  ) %>%

  filter(!is.na(estimate)) %>%

  mutate(
    estimand = case_when(
      str_detect(param, "^estXY_") ~ "XY",
      str_detect(param, "^estYX_") ~ "YX",
      str_detect(param, "^estA_")  ~ "AX",
      str_detect(param, "^estAY_") ~ "AY"
    ),
    method = param %>%
      str_remove("^est(XY|YX|A|AY)_") %>%
      recode(
        CLPM      = "CLPM",
        CLPM_Adj  = "True model",
        CLPM_LBCA = "CLPM linear BCA",
        DPM       = "DPM",
        RI_CLPM   = "RI-CLPM"
      )
  ) %>%
  select(run, occasion, scenario, estimand, method, estimate)

# --- CI lower bounds ---
ciL_df <- results_sim1 %>%
  filter(occasion >= 2) %>%

  pivot_longer(
    cols = matches("^ciL_(XY|YX|A|AY)_"),
    names_to  = "param",
    values_to = "ciL"
  ) %>%

  mutate(
    estimand = case_when(
      str_detect(param, "^ciL_XY_") ~ "XY",
      str_detect(param, "^ciL_YX_") ~ "YX",
      str_detect(param, "^ciL_A_")  ~ "AX",
      str_detect(param, "^ciL_AY_") ~ "AY"
    ),
    method = param %>%
      str_remove("^ciL_(XY|YX|A|AY)_") %>%
      recode(
        CLPM      = "CLPM",
        CLPM_Adj  = "True model",
        CLPM_LBCA = "CLPM linear BCA",
        DPM       = "DPM",
        RI_CLPM   = "RI-CLPM"
      )
  ) %>%
  select(run, occasion, scenario, estimand, method, ciL)

# --- CI upper bounds ---
ciU_df <- results_sim1 %>%
  filter(occasion >= 2) %>%

  pivot_longer(
    cols = matches("^ciU_(XY|YX|A|AY)_"),
    names_to  = "param",
    values_to = "ciU"
  ) %>%

  mutate(
    estimand = case_when(
      str_detect(param, "^ciU_XY_") ~ "XY",
      str_detect(param, "^ciU_YX_") ~ "YX",
      str_detect(param, "^ciU_A_")  ~ "AX",
      str_detect(param, "^ciU_AY_") ~ "AY"
    ),
    method = param %>%
      str_remove("^ciU_(XY|YX|A|AY)_") %>%
      recode(
        CLPM      = "CLPM",
        CLPM_Adj  = "True model",
        CLPM_LBCA = "CLPM linear BCA",
        DPM       = "DPM",
        RI_CLPM   = "RI-CLPM"
      )
  ) %>%
  select(run, occasion, scenario, estimand, method, ciU)

# --- combine everything ---
coverage_df <- est_df %>%
  left_join(ciL_df, by = c("run","occasion","scenario","estimand","method")) %>%
  left_join(ciU_df, by = c("run","occasion","scenario","estimand","method")) %>%
  left_join(true_params, by = "estimand") %>%

  filter(!is.na(ciL), !is.na(ciU)) %>%

  mutate(
    covered = (ciL <= true & true <= ciU)
  ) %>%

  group_by(scenario, occasion, estimand, method) %>%
  reframe(
    nsim = n(),
    coverage = mean(covered),
    mcse_coverage = sqrt(coverage * (1 - coverage) / nsim)
  ) %>%

  mutate(
    scenario = factor(scenario, levels = c("constant", "stepwise")),
    method   = factor(
      method,
      levels = c(
        "CLPM",
        "True model",
        "CLPM linear BCA",
        "DPM",
        "RI-CLPM"
      )
    )
  )

#####################################################
# Colour palette for methods
#####################################################

pal_method <- qualitative_hcl(
  n = nlevels(coverage_df$method),
  palette = "Set 2"
)
names(pal_method) <- levels(coverage_df$method)

#####################################################
# Plot helper
#####################################################

plot_coverage <- function(df, estimand_label) {
  df %>%
    filter(estimand == estimand_label) %>%
    ggplot(aes(
      x = occasion,
      y = coverage,
      color = method,
      group = method
    )) +
    geom_line(linewidth = 0.9) +
    geom_point(size = 2) +
    geom_errorbar(
      aes(
        ymin = coverage - mcse_coverage,
        ymax = coverage + mcse_coverage
      ),
      width = 0.15,
      linewidth = 0.4
    ) +
    geom_hline(yintercept = 0.95, linetype = "dashed", linewidth = 0.4) +
    ggh4x::facet_wrap2(~ scenario, nrow = 1, axes = "y") +
    scale_color_manual(values = pal_method) +
    labs(
      x = "Occasion",
      y = "Coverage",
      color = NULL
    ) +
    theme_classic(base_size = 13) +
    theme(
      panel.spacing.x = unit(1.2, "cm"),
      strip.text      = element_text(size = 14),
      axis.title      = element_text(size = 13),
      axis.text       = element_text(size = 12),
      legend.position = "bottom",
      legend.text     = element_text(size = 12)
    )
}

#####################################################
# Build and display all four plots
#####################################################

plot_cov_XY <- plot_coverage(coverage_df, "XY")
plot_cov_YX <- plot_coverage(coverage_df, "YX")
plot_cov_AX <- plot_coverage(coverage_df, "AX")
plot_cov_AY <- plot_coverage(coverage_df, "AY")

plot_cov_XY
plot_cov_YX
plot_cov_AX
plot_cov_AY

```

### plot 4.2: type 1 error Simulation Study 1

These plots show the Type I error rates for the null-true parameters in Simulation Study 1. Type I error is the proportion of simulation runs in which a true null hypothesis is incorrectly rejected, indicating how well each method controls false positives over time. 

```{r Type 1 error plots}

#####################################################
# Build type I error data for all lagged parameters
#####################################################

# true parameters from the data-generating model
true_params <- tibble(
  estimand = c("XY", "YX", "AX", "AY"),
  true     = c(
    A[2, 1],  # XY
    A[1, 2],  # YX (typically 0)
    A[1, 1],  # AX
    A[2, 2]   # AY
  )
)

# --- p-values ---
pval_df <- results_sim1 %>%

  # only occasions with lagged parameters
  filter(occasion >= 2) %>%

  # reshape p-values to long format
  pivot_longer(
    cols = matches("^p(XY|YX|A|AY)_"),
    names_to  = "param",
    values_to = "pval"
  ) %>%

  # drop failed fits
  filter(!is.na(pval)) %>%

  # identify estimand type and method
  mutate(
    estimand = case_when(
      str_detect(param, "^pXY_") ~ "XY",
      str_detect(param, "^pYX_") ~ "YX",
      str_detect(param, "^pA_")  ~ "AX",
      str_detect(param, "^pAY_") ~ "AY"
    ),

    method = param %>%
      str_remove("^p(XY|YX|A|AY)_") %>%
      recode(
        CLPM      = "CLPM",
        CLPM_Adj  = "True model",
        CLPM_LBCA = "CLPM linear BCA",
        DPM       = "DPM",
        RI_CLPM   = "RI-CLPM"
      )
  ) %>%
  select(run, occasion, scenario, estimand, method, pval)

# build type I error data frame
type1_df <- pval_df %>%

  # attach true parameter values
  left_join(true_params, by = "estimand") %>%

  # only estimands where the null is true (theta = 0)
  filter(true == 0) %>%

  # indicator: reject at alpha
  mutate(
    reject = (pval <= 0.05)
  ) %>%

  # type I error rate and Monte Carlo SE
  group_by(scenario, occasion, estimand, method) %>%
  reframe(
    nsim = n(),

    # type I error rate
    type1 = mean(reject),

    # Monte Carlo SE of type I error rate
    mcse_type1 =
      sqrt(type1 * (1 - type1) / nsim)
  ) %>%

  # clean factor ordering
  mutate(
    scenario = factor(scenario, levels = c("constant", "stepwise")),
    method   = factor(
      method,
      levels = c(
        "CLPM",
        "True model",
        "CLPM linear BCA",
        "DPM",
        "RI-CLPM"
      )
    )
  )

#####################################################
# Colour palette for methods (shared across plots)
#####################################################

pal_method <- qualitative_hcl(
  n = nlevels(type1_df$method),
  palette = "Set 2"
)
names(pal_method) <- levels(type1_df$method)

#####################################################
# Plot: type I error (only for null-true estimands)
#####################################################

plot_type1 <- type1_df %>%

  ggplot(aes(
    x     = occasion,
    y     = type1,
    color = method,
    group = method
  )) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  geom_errorbar(
    aes(
      ymin = type1 - mcse_type1,
      ymax = type1 + mcse_type1
    ),
    width = 0.15,
    linewidth = 0.4
  ) +

  # nominal alpha reference line
  geom_hline(yintercept = 0.05, linetype = "dashed", linewidth = 0.4) +

  # facet by scenario
  ggh4x::facet_wrap2(
    ~ scenario,
    nrow = 1,
    axes = "y"
  ) +

  # colours
  scale_color_manual(values = pal_method) +

  # labels
  labs(
    x = "Occasion",
    y = "Type I error",
    color = NULL
  ) +

  # scientific theme
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12)
  )

#####################################################
# Display plot
#####################################################

plot_type1

```

### Ploit 4.3: Power (type 2 error) Simulation Study 1

These plots show the statistical power for the non-null parameters in Simulation Study 1. Power is the proportion of simulation runs in which a false null hypothesis is correctly rejected, and reflects how well each method detects true effects over time.

```{r Power plots}

#####################################################
# Build power data for all lagged parameters
#####################################################

# true parameters from the data-generating model
true_params <- tibble(
  estimand = c("XY", "YX", "AX", "AY"),
  true     = c(
    A[2, 1],  # XY
    A[1, 2],  # YX (usually 0)
    A[1, 1],  # AX
    A[2, 2]   # AY
  )
)

# --- p-values ---
pval_df <- results_sim1 %>%

  # only occasions with lagged parameters
  filter(occasion >= 2) %>%

  # reshape p-values to long format
  pivot_longer(
    cols = matches("^p(XY|YX|A|AY)_"),
    names_to  = "param",
    values_to = "pval"
  ) %>%

  # drop failed fits
  filter(!is.na(pval)) %>%

  # identify estimand type and method
  mutate(
    estimand = case_when(
      str_detect(param, "^pXY_") ~ "XY",
      str_detect(param, "^pYX_") ~ "YX",
      str_detect(param, "^pA_")  ~ "AX",
      str_detect(param, "^pAY_") ~ "AY"
    ),

    method = param %>%
      str_remove("^p(XY|YX|A|AY)_") %>%
      recode(
        CLPM      = "CLPM",
        CLPM_Adj  = "True model",
        CLPM_LBCA = "CLPM linear BCA",
        DPM       = "DPM",
        RI_CLPM   = "RI-CLPM"
      )
  ) %>%
  select(run, occasion, scenario, estimand, method, pval)

# build power data frame
power_df <- pval_df %>%

  # attach true parameter values
  left_join(true_params, by = "estimand") %>%

  # only estimands where the null is false
  filter(true != 0) %>%

  # indicator: reject at alpha
  mutate(
    reject = (pval <= 0.05)
  ) %>%

  # power and Monte Carlo SE
  group_by(scenario, occasion, estimand, method) %>%
  reframe(
    nsim = n(),

    # power
    power = mean(reject),

    # Monte Carlo SE of power
    mcse_power =
      sqrt(power * (1 - power) / nsim)
  ) %>%

  # clean factor ordering
  mutate(
    scenario = factor(scenario, levels = c("constant", "stepwise")),
    method   = factor(
      method,
      levels = c(
        "CLPM",
        "True model",
        "CLPM linear BCA",
        "DPM",
        "RI-CLPM"
      )
    )
  )

#####################################################
# Colour palette for methods (shared across plots)
#####################################################

pal_method <- qualitative_hcl(
  n = nlevels(power_df$method),
  palette = "Set 2"
)
names(pal_method) <- levels(power_df$method)

#####################################################
# Plot helper (used for all three estimands)
#####################################################

plot_power <- function(df, estimand_label) {

  df %>%
    filter(estimand == estimand_label) %>%

    ggplot(aes(
      x     = occasion,
      y     = power,
      color = method,
      group = method
    )) +
    geom_line(linewidth = 0.9) +
    geom_point(size = 2) +
    geom_errorbar(
      aes(
        ymin = power - mcse_power,
        ymax = power + mcse_power
      ),
      width = 0.15,
      linewidth = 0.4
    ) +

    # reference line for high power
    geom_hline(yintercept = 0.80, linetype = "dashed", linewidth = 0.4) +

    # facet by scenario
    ggh4x::facet_wrap2(
      ~ scenario,
      nrow = 1,
      axes = "y"
    ) +

    # colours
    scale_color_manual(values = pal_method) +

    # labels
    labs(
      x = "Occasion",
      y = "Power",
      color = NULL
    ) +

    # scientific theme
    theme_classic(base_size = 13) +
    theme(
      panel.spacing.x = unit(1.2, "cm"),
      strip.text      = element_text(size = 14),
      axis.title      = element_text(size = 13),
      axis.text       = element_text(size = 12),
      legend.position = "bottom",
      legend.text     = element_text(size = 12)
    )
}

#####################################################
# Build and display all three power plots
#####################################################

plot_power_XY <- plot_power(power_df, "XY")
plot_power_AX <- plot_power(power_df, "AX")
plot_power_AY <- plot_power(power_df, "AY")

plot_power_XY
plot_power_AX
plot_power_AY

```

concatenate these results into a single object for later use

```{r}
library(patchwork)

# create legendless plots
plot_relbias_XY_noleg <- plot_relbias_XY +
  theme(legend.position = "none")

# combine the two
combined_XY <- plot_relbias_XY_noleg / plot_rmse_XY

ggsave(
  filename = "combined_XY_relbias_RMSE_1.png",
  plot     = combined_XY,
  path     = "C:/Users/Admin/Desktop",
  width    = 10,
  height   = 8,
  dpi      = 300
)

```

# Simulation Study 2 Results

```{r}

# load the function and dependencies
source("Thesis\\Code\\Qmd_Rmd_R_Pyth\\Sim Studies\\Sim_Study_2_RR_no_sampling.R")

# set a seed for reproducibility
set.seed(183718)

# autoregressive + cross-lag structure
A <- matrix(c(
  0.20, 0,                         # autoregressive and cross-lag for Y
  0.10, 0.20                       # cross-lag and autoregressive for X
), nrow = 2, byrow = TRUE)

# confounder covariance (3 confounders)
Psi <- diag(3)                     # uncorrelated confounders with var=1

# sample B1 matrix
B1 <- sample_B1(
  k = 3,
  R2_1 = 0.15,
  eta_1 = 0.50,
  min_abs = 0.001,
  max_abs = 0.40
)

# for the constant scenario, replicate B1 across time points
B_constant <- lapply(1:5, function(t) B1)

# scaling factor such that R_2 becomes twice as large: scaling factor = sqrt(old R2 / new R2)
scale_factor <- sqrt(0.40 / 0.15)              # = 1.414214
                                  
# create beta list for increasing R2 scenario
B_high <- B1 * scale_factor

# fill the list
B_stepwise <- list(
  t1 = B1,
  t2 = B1,
  t3 = B1,
  t4 = B_high,
  t5 = B_high
)

# call the simulation function
results_sim2 <- run_simulation_study2(
  reps          = 1000,                                # replications
  N             = 1000,                                # sample size
  T             = 5,                                   # number of time points
  k             = 3,                                   # number of linear confounders
  scenarios     = c("constant", "stepwise"),           # B scenarios
  B_scenarios   = list(                                # beta trajectories
    constant = B_constant,
    stepwise = B_stepwise
  ),
  A             = A,                                   # AR + cross-lag matrix
  Psi           = Psi,                                 # confounder covariance matrix
  rho_extra     = 0.1,                                 # extra residual correlation
  models_to_run = c(                                   # models to run
    "clpm",
    "riclpm",
    "dpm",
    "adj",
    "lbca",
    "xgb",
    "true"
  ),
  cores         = 6,                                   # parallel cores
  base_seed     = 12345                                # base seed
)

```

### Plot 0.1: Beta trajectory plots Simulation Study 2

```{r}

# libraries
library(tidyverse)
library(ggh4x)
library(colorspace)

# first load the data
df2 <- results_sim2

# how many runs were not successful?
sum(df2$is_na_run)

# custom colour palette (scientific, colorblind-friendly)
pal_beta <- qualitative_hcl(
  n = ncol(B1),
  palette = "Set 2"
)

# number of occasions
T <- length(B_stepwise)

# confounder names
conf_names <- colnames(B1)

# identify linear vs non-linear terms
term_type <- tibble(
  term = conf_names,
  type = if_else(str_detect(term, ":"), "Non-linear", "Linear")
)

# colour palette (2 colours only)
pal_beta <- qualitative_hcl(
  n = 2,
  palette = "Set 2"
)
names(pal_beta) <- c("Linear", "Non-linear")

# build beta trajectory dataframe
df_beta <- bind_rows(

  # constant scenario
  lapply(1:T, function(t) {
    tibble(
      occasion = t,
      term     = conf_names,
      beta_X   = B_constant[[t]][1, ],
      beta_Y   = B_constant[[t]][2, ],
      scenario = "constant"
    )
  }),

  # stepwise scenario
  lapply(1:T, function(t) {
    tibble(
      occasion = t,
      term     = conf_names,
      beta_X   = B_stepwise[[t]][1, ],
      beta_Y   = B_stepwise[[t]][2, ],
      scenario = "stepwise"
    )
  })
) %>%
  left_join(term_type, by = "term") %>%
  mutate(
    scenario = factor(scenario, levels = c("constant", "stepwise")),
    type     = factor(type, levels = c("Linear", "Non-linear"))
  )

plot_beta_X <- df_beta %>%
  ggplot(aes(
    x     = occasion,
    y     = beta_X,
    group = term,
    color = type
  )) +
  geom_line(linewidth = 0.9, alpha = 0.9) +
  ggh4x::facet_wrap2(
    ~ scenario,
    nrow = 1,
    axes = "y"
  ) +
  scale_color_manual(values = pal_beta) +
  labs(
    x = "Occasion",
    y = "Effect of confounders on X",
    color = NULL
  ) +
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12)
  )

# same for Y
plot_beta_Y <- df_beta %>%
  ggplot(aes(
    x     = occasion,
    y     = beta_Y,
    group = term,
    color = type
  )) +
  geom_line(linewidth = 0.9, alpha = 0.9) +
  ggh4x::facet_wrap2(
    ~ scenario,
    nrow = 1,
    axes = "y"
  ) +
  scale_color_manual(values = pal_beta) +
  labs(
    x = "Occasion",
    y = "Effect of confounders on Y",
    color = NULL
  ) +
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12)
  )

# display the plots
plot_beta_X
plot_beta_Y

```

### Plot 0.2: Variance Explained Simulation Study 2

```{r}

# R-squared / variance explained trajectories
# (since all vars have Var = 1, you said we can use beta^2)

df_r2 <- df_beta %>%
  mutate(
    r2_X = beta_X^2,
    r2_Y = beta_Y^2
  )

plot_r2_X <- df_r2 %>%
  ggplot(aes(
    x     = occasion,
    y     = r2_X,
    group = term,
    color = type
  )) +
  geom_line(linewidth = 0.9, alpha = 0.9) +
  ggh4x::facet_wrap2(
    ~ scenario,
    nrow = 1,
    axes = "y"
  ) +
  scale_color_manual(values = pal_beta) +
  labs(
    x = "Occasion",
    y = expression(R^2*" (variance explained) in X"),
    color = NULL
  ) +
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12)
  )

plot_r2_Y <- df_r2 %>%
  ggplot(aes(
    x     = occasion,
    y     = r2_Y,
    group = term,
    color = type
  )) +
  geom_line(linewidth = 0.9, alpha = 0.9) +
  ggh4x::facet_wrap2(
    ~ scenario,
    nrow = 1,
    axes = "y"
  ) +
  scale_color_manual(values = pal_beta) +
  labs(
    x = "Occasion",
    y = expression(R^2*" (variance explained) in Y"),
    color = NULL
  ) +
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12)
  )

# display
plot_r2_X
plot_r2_Y

```

### Plot 1: relative bias Simulation Study 2

```{r}

library(tidyverse)
library(ggh4x)
library(viridis)

#####################################################
# True parameters
#####################################################

true_params <- tibble(
  estimand = c("XY", "YX", "AX", "AY"),
  true     = c(
    A[2, 1],  # XY
    A[1, 2],  # YX
    A[1, 1],  # AX
    A[2, 2]   # AY
  )
)

#####################################################
# Build relative bias data for all lagged parameters
#####################################################

relbias_df <- results_sim2 %>%

  # only occasions with lagged parameters
  filter(occasion >= 2) %>%

  # reshape estimates to long format
  pivot_longer(
    cols = matches("^est(XY|YX|A|AY)_"),
    names_to  = "param",
    values_to = "estimate"
  ) %>%

  # drop failed fits
  filter(!is.na(estimate)) %>%

  # identify estimand type and method
  mutate(
    estimand = case_when(
      str_detect(param, "^estXY_") ~ "XY",
      str_detect(param, "^estYX_") ~ "YX",
      str_detect(param, "^estA_")  ~ "AX",
      str_detect(param, "^estAY_") ~ "AY"
    ),

    method = param %>%
      str_remove("^est(XY|YX|A|AY)_") %>%
      recode(
        CLPM      = "CLPM",
        CLPM_Adj  = "Linear adj CLPM",
        CLPM_LBCA = "CLPM linear BCA",
        DPM       = "DPM",
        RI_CLPM   = "RI-CLPM",
        CLPM_XGB  = "CLPM XGB BCA"
      )
  ) %>%

  # attach true parameter values
  left_join(true_params, by = "estimand") %>%

  # compute relative bias and Monte Carlo SE
  group_by(scenario, occasion, estimand, method) %>%
  reframe(
    nsim = n(),
    mean_est = mean(estimate),

    # relative bias
    rel_bias = (mean_est - true) / true,

    # Monte Carlo SE of relative bias
    mcse_rel_bias =
      sqrt(
        sum((estimate - mean_est)^2) /
        (nsim * (nsim - 1))
      ) / abs(true)
  ) %>%

  # clean factor ordering
  mutate(
    scenario = factor(scenario, levels = c("constant", "stepwise")),
    method   = factor(
      method,
      levels = c(
        "CLPM",
        "Linear adj CLPM",
        "CLPM linear BCA",
        "CLPM XGB BCA",
        "DPM",
        "RI-CLPM"
      )
    )
  )

#####################################################
# Colour palette for methods (shared across plots)
#####################################################

pal_method <- viridis(
  n = nlevels(relbias_df$method),
  option = "viridis",
  begin  = 0.1,
  end    = 0.9
)
names(pal_method) <- levels(relbias_df$method)

#####################################################
# Plot 1: cross-lagged effect of X on Y (XY)
#####################################################

plot_relbias_XY <- relbias_df %>%
  filter(estimand == "XY") %>%

  ggplot(aes(
    x     = occasion,
    y     = rel_bias,
    color = method,
    group = method
  )) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  geom_errorbar(
    aes(
      ymin = rel_bias - mcse_rel_bias,
      ymax = rel_bias + mcse_rel_bias
    ),
    width = 0.15,
    linewidth = 0.4
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.4) +
  ggh4x::facet_wrap2(~ scenario, nrow = 1, axes = "y") +
  scale_color_manual(values = pal_method) +
  labs(x = "Occasion", y = "Relative bias", color = NULL) +
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12)
  )

#####################################################
# Plot 2: autoregressive effect of X on X (AX)
#####################################################

plot_relbias_AX <- relbias_df %>%
  filter(estimand == "AX") %>%

  ggplot(aes(occasion, rel_bias, color = method, group = method)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  geom_errorbar(
    aes(
      ymin = rel_bias - mcse_rel_bias,
      ymax = rel_bias + mcse_rel_bias
    ),
    width = 0.15,
    linewidth = 0.4
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.4) +
  ggh4x::facet_wrap2(~ scenario, nrow = 1, axes = "y") +
  scale_color_manual(values = pal_method) +
  labs(x = "Occasion", y = "Relative bias", color = NULL) +
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    legend.position = "bottom"
  )

#####################################################
# Plot 3: autoregressive effect of Y on Y (AY)
#####################################################

plot_relbias_AY <- relbias_df %>%
  filter(estimand == "AY") %>%

  ggplot(aes(occasion, rel_bias, color = method, group = method)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  geom_errorbar(
    aes(
      ymin = rel_bias - mcse_rel_bias,
      ymax = rel_bias + mcse_rel_bias
    ),
    width = 0.15,
    linewidth = 0.4
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.4) +
  ggh4x::facet_wrap2(~ scenario, nrow = 1, axes = "y") +
  scale_color_manual(values = pal_method) +
  labs(x = "Occasion", y = "Relative bias", color = NULL) +
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    legend.position = "bottom"
  )

#####################################################
# Display all plots
#####################################################

plot_relbias_XY
plot_relbias_AX
plot_relbias_AY

```

### Plot 1.2: Bias Simulation Study 2

```{r}

#####################################################
# Build bias data for all lagged parameters
#####################################################

# true parameters from the data-generating model
true_params <- tibble(
  estimand = c("XY", "YX", "AX", "AY"),
  true     = c(
    A[2, 1],  # XY
    A[1, 2],  # YX (true = 0)
    A[1, 1],  # AX
    A[2, 2]   # AY
  )
)

# build bias data frame
bias_df <- results_sim2 %>%

  # only occasions with lagged parameters
  filter(occasion >= 2) %>%

  # reshape estimates to long format
  pivot_longer(
    cols = matches("^est(XY|YX|A|AY)_"),
    names_to  = "param",
    values_to = "estimate"
  ) %>%

  # drop failed fits
  filter(!is.na(estimate)) %>%

  # identify estimand type and method
  mutate(
    estimand = case_when(
      str_detect(param, "^estXY_") ~ "XY",
      str_detect(param, "^estYX_") ~ "YX",
      str_detect(param, "^estA_")  ~ "AX",
      str_detect(param, "^estAY_") ~ "AY"
    ),

    method = param %>%
      str_remove("^est(XY|YX|A|AY)_") %>%
      recode(
        CLPM      = "CLPM",
        CLPM_Adj  = "Lin adj",
        CLPM_LBCA = "CLPM linear BCA",
        DPM       = "DPM",
        RI_CLPM   = "RI-CLPM",
        CLPM_XGB  = "CLPM XGB"
      )
  ) %>%

  # attach true parameter values
  left_join(true_params, by = "estimand") %>%

  # compute bias and Monte Carlo SE
  group_by(scenario, occasion, estimand, method) %>%
  reframe(
    nsim = n(),

    mean_est = mean(estimate),

    # bias
    bias = mean_est - true,

    # Monte Carlo SE of the bias
    mcse_bias =
      sqrt(
        sum((estimate - mean_est)^2) /
        (nsim * (nsim - 1))
      )
  ) %>%

  # clean factor ordering
  mutate(
    scenario = factor(scenario, levels = c("constant", "stepwise")),
    method   = factor(
      method,
      levels = c(
        "CLPM",
        "Lin adj",
        "CLPM linear BCA",
        "CLPM XGB",
        "DPM",
        "RI-CLPM"
      )
    )
  )

#####################################################
# Colour palette for methods (shared across plots)
#####################################################

pal_method <- qualitative_hcl(
  n = nlevels(bias_df$method),
  palette = "Set 2"
)
names(pal_method) <- levels(bias_df$method)

#####################################################
# Plot 1: cross-lagged effect of X on Y (XY)
#####################################################

plot_bias_XY <- bias_df %>%
  filter(estimand == "XY") %>%

  ggplot(aes(
    x     = occasion,
    y     = bias,
    color = method,
    group = method
  )) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  geom_errorbar(
    aes(
      ymin = bias - mcse_bias,
      ymax = bias + mcse_bias
    ),
    width = 0.15,
    linewidth = 0.4
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.4) +
  ggh4x::facet_wrap2(~ scenario, nrow = 1, axes = "y") +
  scale_color_manual(values = pal_method) +
  labs(
    x = "Occasion",
    y = "Bias",
    color = NULL
  ) +
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12)
  )

#####################################################
# Plot 2: cross-lagged effect of Y on X (YX)
#####################################################

plot_bias_YX <- bias_df %>%
  filter(estimand == "YX") %>%

  ggplot(aes(
    x     = occasion,
    y     = bias,
    color = method,
    group = method
  )) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  geom_errorbar(
    aes(
      ymin = bias - mcse_bias,
      ymax = bias + mcse_bias
    ),
    width = 0.15,
    linewidth = 0.4
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.4) +
  ggh4x::facet_wrap2(~ scenario, nrow = 1, axes = "y") +
  scale_color_manual(values = pal_method) +
  labs(
    x = "Occasion",
    y = "Bias",
    color = NULL
  ) +
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12)
  )

#####################################################
# Plot 3: autoregressive effect of X on X (AX)
#####################################################

plot_bias_AX <- bias_df %>%
  filter(estimand == "AX") %>%

  ggplot(aes(
    x     = occasion,
    y     = bias,
    color = method,
    group = method
  )) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  geom_errorbar(
    aes(
      ymin = bias - mcse_bias,
      ymax = bias + mcse_bias
    ),
    width = 0.15,
    linewidth = 0.4
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.4) +
  ggh4x::facet_wrap2(~ scenario, nrow = 1, axes = "y") +
  scale_color_manual(values = pal_method) +
  labs(
    x = "Occasion",
    y = "Bias",
    color = NULL
  ) +
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12)
  )

#####################################################
# Plot 4: autoregressive effect of Y on Y (AY)
#####################################################

plot_bias_AY <- bias_df %>%
  filter(estimand == "AY") %>%

  ggplot(aes(
    x     = occasion,
    y     = bias,
    color = method,
    group = method
  )) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  geom_errorbar(
    aes(
      ymin = bias - mcse_bias,
      ymax = bias + mcse_bias
    ),
    width = 0.15,
    linewidth = 0.4
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.4) +
  ggh4x::facet_wrap2(~ scenario, nrow = 1, axes = "y") +
  scale_color_manual(values = pal_method) +
  labs(
    x = "Occasion",
    y = "Bias",
    color = NULL
  ) +
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12)
  )

#####################################################
# Display all plots
#####################################################

plot_bias_XY
plot_bias_YX
plot_bias_AX
plot_bias_AY

```

### Plot 2: RMSE Simulation Study 2

```{r}

library(tidyverse)
library(ggh4x)
library(viridis)

#####################################################
# Build RMSE data for all lagged parameters
#####################################################

# true parameters from the data-generating model
true_params <- tibble(
  estimand = c("XY", "YX", "AX", "AY"),
  true     = c(
    A[2, 1],  # XY
    A[1, 2],  # YX (true = 0)
    A[1, 1],  # AX
    A[2, 2]   # AY
  )
)

# build RMSE data frame
rmse_df <- results_sim2 %>%

  # only occasions with lagged parameters
  filter(occasion >= 2) %>%

  # reshape estimates to long format
  pivot_longer(
    cols = matches("^est(XY|YX|A|AY)_"),
    names_to  = "param",
    values_to = "estimate"
  ) %>%

  # drop failed fits
  filter(!is.na(estimate)) %>%

  # identify estimand type and method
  mutate(
    estimand = case_when(
      str_detect(param, "^estXY_") ~ "XY",
      str_detect(param, "^estYX_") ~ "YX",
      str_detect(param, "^estA_")  ~ "AX",
      str_detect(param, "^estAY_") ~ "AY"
    ),

    method = param %>%
      str_remove("^est(XY|YX|A|AY)_") %>%
      recode(
        CLPM      = "CLPM",
        CLPM_Adj  = "Lin adj",
        CLPM_LBCA = "CLPM linear BCA",
        CLPM_XGB  = "CLPM XGB BCA",
        DPM       = "DPM",
        RI_CLPM   = "RI-CLPM"
      )
  ) %>%

  # attach true parameter values
  left_join(true_params, by = "estimand") %>%

  # compute RMSE and Monte Carlo SE
  group_by(scenario, occasion, estimand, method) %>%
  reframe(
    nsim = n(),

    # RMSE
    rmse = sqrt(mean((estimate - true)^2)),

    # Monte Carlo SE of RMSE
    mcse_rmse =
      sqrt(
        sum(
          ((estimate - true)^2 - mean((estimate - true)^2))^2
        ) /
        (nsim * (nsim - 1))
      ) / (2 * rmse)
  ) %>%

  # clean factor ordering
  mutate(
    scenario = factor(scenario, levels = c("constant", "stepwise")),
    method   = factor(
      method,
      levels = c(
        "CLPM",
        "Lin adj",
        "CLPM linear BCA",
        "CLPM XGB BCA",
        "DPM",
        "RI-CLPM"
      )
    )
  )

#####################################################
# Colour palette for methods (shared across plots)
#####################################################

pal_method <- viridis(
  n = nlevels(rmse_df$method),
  option = "viridis",
  begin  = 0.1,
  end    = 0.9
)
names(pal_method) <- levels(rmse_df$method)

#####################################################
# Plot 1: cross-lagged effect of X on Y (XY)
#####################################################

plot_rmse_XY <- rmse_df %>%
  filter(estimand == "XY") %>%

  ggplot(aes(
    x     = occasion,
    y     = rmse,
    color = method,
    group = method
  )) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  geom_errorbar(
    aes(
      ymin = rmse - mcse_rmse,
      ymax = rmse + mcse_rmse
    ),
    width = 0.15,
    linewidth = 0.4
  ) +
  ggh4x::facet_wrap2(~ scenario, nrow = 1, axes = "y") +
  scale_color_manual(values = pal_method) +
  labs(x = "Occasion", y = "RMSE", color = NULL) +
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12)
  )

#####################################################
# Plot 2: cross-lagged effect of Y on X (YX)
#####################################################

plot_rmse_YX <- rmse_df %>%
  filter(estimand == "YX") %>%

  ggplot(aes(
    x     = occasion,
    y     = rmse,
    color = method,
    group = method
  )) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  geom_errorbar(
    aes(
      ymin = rmse - mcse_rmse,
      ymax = rmse + mcse_rmse
    ),
    width = 0.15,
    linewidth = 0.4
  ) +
  ggh4x::facet_wrap2(~ scenario, nrow = 1, axes = "y") +
  scale_color_manual(values = pal_method) +
  labs(x = "Occasion", y = "RMSE", color = NULL) +
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12)
  )

#####################################################
# Plot 3: autoregressive effect of X on X (AX)
#####################################################

plot_rmse_AX <- rmse_df %>%
  filter(estimand == "AX") %>%

  ggplot(aes(
    x     = occasion,
    y     = rmse,
    color = method,
    group = method
  )) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  geom_errorbar(
    aes(
      ymin = rmse - mcse_rmse,
      ymax = rmse + mcse_rmse
    ),
    width = 0.15,
    linewidth = 0.4
  ) +
  ggh4x::facet_wrap2(~ scenario, nrow = 1, axes = "y") +
  scale_color_manual(values = pal_method) +
  labs(x = "Occasion", y = "RMSE", color = NULL) +
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12)
  )

#####################################################
# Plot 4: autoregressive effect of Y on Y (AY)
#####################################################

plot_rmse_AY <- rmse_df %>%
  filter(estimand == "AY") %>%

  ggplot(aes(
    x     = occasion,
    y     = rmse,
    color = method,
    group = method
  )) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  geom_errorbar(
    aes(
      ymin = rmse - mcse_rmse,
      ymax = rmse + mcse_rmse
    ),
    width = 0.15,
    linewidth = 0.4
  ) +
  ggh4x::facet_wrap2(~ scenario, nrow = 1, axes = "y") +
  scale_color_manual(values = pal_method) +
  labs(x = "Occasion", y = "RMSE", color = NULL) +
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12)
  )

#####################################################
# Display all plots
#####################################################

plot_rmse_XY
plot_rmse_YX
plot_rmse_AX
plot_rmse_AY

```

### Plot 4.1: Coverage Simulation Study 2


```{r}

#####################################################
# Build coverage data for all lagged parameters
#####################################################

# true parameters from the data-generating model
true_params <- tibble(
  estimand = c("XY", "YX", "AX", "AY"),
  true     = c(
    A[2, 1],  # XY
    A[1, 2],  # YX
    A[1, 1],  # AX
    A[2, 2]   # AY
  )
)

#####################################################
# --- estimates ---
#####################################################

est_df <- results_sim2 %>%

  # only occasions with lagged parameters
  filter(occasion >= 2) %>%

  pivot_longer(
    cols = matches("^est(XY|YX|A|AY)_"),
    names_to  = "param",
    values_to = "estimate"
  ) %>%

  filter(!is.na(estimate)) %>%

  mutate(
    estimand = case_when(
      str_detect(param, "^estXY_") ~ "XY",
      str_detect(param, "^estYX_") ~ "YX",
      str_detect(param, "^estA_")  ~ "AX",
      str_detect(param, "^estAY_") ~ "AY"
    ),
    method = param %>%
      str_remove("^est(XY|YX|A|AY)_") %>%
      recode(
        CLPM      = "CLPM",
        CLPM_Adj  = "Lin adj",
        CLPM_LBCA = "CLPM linear BCA",
        CLPM_XGB  = "CLPM XGB BCA",
        DPM       = "DPM",
        RI_CLPM   = "RI-CLPM"
      )
  ) %>%
  select(run, occasion, scenario, estimand, method, estimate)

#####################################################
# --- CI lower bounds ---
#####################################################

ciL_df <- results_sim2 %>%
  filter(occasion >= 2) %>%

  pivot_longer(
    cols = matches("^ciL_(XY|YX|A|AY)_"),
    names_to  = "param",
    values_to = "ciL"
  ) %>%

  mutate(
    estimand = case_when(
      str_detect(param, "^ciL_XY_") ~ "XY",
      str_detect(param, "^ciL_YX_") ~ "YX",
      str_detect(param, "^ciL_A_")  ~ "AX",
      str_detect(param, "^ciL_AY_") ~ "AY"
    ),
    method = param %>%
      str_remove("^ciL_(XY|YX|A|AY)_") %>%
      recode(
        CLPM      = "CLPM",
        CLPM_Adj  = "Lin adj",
        CLPM_LBCA = "CLPM linear BCA",
        CLPM_XGB  = "CLPM XGB BCA",
        DPM       = "DPM",
        RI_CLPM   = "RI-CLPM"
      )
  ) %>%
  select(run, occasion, scenario, estimand, method, ciL)

#####################################################
# --- CI upper bounds ---
#####################################################

ciU_df <- results_sim2 %>%
  filter(occasion >= 2) %>%

  pivot_longer(
    cols = matches("^ciU_(XY|YX|A|AY)_"),
    names_to  = "param",
    values_to = "ciU"
  ) %>%

  mutate(
    estimand = case_when(
      str_detect(param, "^ciU_XY_") ~ "XY",
      str_detect(param, "^ciU_YX_") ~ "YX",
      str_detect(param, "^ciU_A_")  ~ "AX",
      str_detect(param, "^ciU_AY_") ~ "AY"
    ),
    method = param %>%
      str_remove("^ciU_(XY|YX|A|AY)_") %>%
      recode(
        CLPM      = "CLPM",
        CLPM_Adj  = "Lin adj",
        CLPM_LBCA = "CLPM linear BCA",
        CLPM_XGB  = "CLPM XGB BCA",
        DPM       = "DPM",
        RI_CLPM   = "RI-CLPM"
      )
  ) %>%
  select(run, occasion, scenario, estimand, method, ciU)

#####################################################
# --- combine everything ---
#####################################################

coverage_df <- est_df %>%
  left_join(ciL_df, by = c("run","occasion","scenario","estimand","method")) %>%
  left_join(ciU_df, by = c("run","occasion","scenario","estimand","method")) %>%
  left_join(true_params, by = "estimand") %>%

  filter(!is.na(ciL), !is.na(ciU)) %>%

  mutate(
    covered = (ciL <= true & true <= ciU)
  ) %>%

  group_by(scenario, occasion, estimand, method) %>%
  reframe(
    nsim = n(),
    coverage = mean(covered),
    mcse_coverage = sqrt(coverage * (1 - coverage) / nsim)
  ) %>%

  mutate(
    scenario = factor(scenario, levels = c("constant", "stepwise")),
    method   = factor(
      method,
      levels = c(
        "CLPM",
        "Lin adj",
        "CLPM linear BCA",
        "CLPM XGB BCA",
        "DPM",
        "RI-CLPM"
      )
    )
  )

#####################################################
# Colour palette for methods
#####################################################

pal_method <- qualitative_hcl(
  n = nlevels(coverage_df$method),
  palette = "Set 2"
)
names(pal_method) <- levels(coverage_df$method)

#####################################################
# Plot helper
#####################################################

plot_coverage <- function(df, estimand_label) {
  df %>%
    filter(estimand == estimand_label) %>%
    ggplot(aes(
      x = occasion,
      y = coverage,
      color = method,
      group = method
    )) +
    geom_line(linewidth = 0.9) +
    geom_point(size = 2) +
    geom_errorbar(
      aes(
        ymin = coverage - mcse_coverage,
        ymax = coverage + mcse_coverage
      ),
      width = 0.15,
      linewidth = 0.4
    ) +
    geom_hline(yintercept = 0.95, linetype = "dashed", linewidth = 0.4) +
    ggh4x::facet_wrap2(~ scenario, nrow = 1, axes = "y") +
    scale_color_manual(values = pal_method) +
    labs(
      x = "Occasion",
      y = "Coverage",
      color = NULL
    ) +
    theme_classic(base_size = 13) +
    theme(
      panel.spacing.x = unit(1.2, "cm"),
      strip.text      = element_text(size = 14),
      axis.title      = element_text(size = 13),
      axis.text       = element_text(size = 12),
      legend.position = "bottom",
      legend.text     = element_text(size = 12)
    )
}

#####################################################
# Build and display all four plots
#####################################################

plot_cov_XY <- plot_coverage(coverage_df, "XY")
plot_cov_YX <- plot_coverage(coverage_df, "YX")
plot_cov_AX <- plot_coverage(coverage_df, "AX")
plot_cov_AY <- plot_coverage(coverage_df, "AY")

plot_cov_XY
plot_cov_YX
plot_cov_AX
plot_cov_AY

```

### Plot 4.2: Type I Error Simulation Study 2

```{r}


#####################################################
# Build type I error data for all lagged parameters
#####################################################

# true parameters from the data-generating model
true_params <- tibble(
  estimand = c("XY", "YX", "AX", "AY"),
  true     = c(
    A[2, 1],  # XY
    A[1, 2],  # YX (typically 0)
    A[1, 1],  # AX
    A[2, 2]   # AY
  )
)

#####################################################
# --- p-values ---
#####################################################

pval_df <- results_sim2 %>%

  # only occasions with lagged parameters
  filter(occasion >= 2) %>%

  # reshape p-values to long format
  pivot_longer(
    cols = matches("^p(XY|YX|A|AY)_"),
    names_to  = "param",
    values_to = "pval"
  ) %>%

  # drop failed fits
  filter(!is.na(pval)) %>%

  # identify estimand type and method
  mutate(
    estimand = case_when(
      str_detect(param, "^pXY_") ~ "XY",
      str_detect(param, "^pYX_") ~ "YX",
      str_detect(param, "^pA_")  ~ "AX",
      str_detect(param, "^pAY_") ~ "AY"
    ),

    method = param %>%
      str_remove("^p(XY|YX|A|AY)_") %>%
      recode(
        CLPM      = "CLPM",
        CLPM_Adj  = "Lin adj",
        CLPM_LBCA = "CLPM linear BCA",
        CLPM_XGB  = "CLPM XGB BCA",
        DPM       = "DPM",
        RI_CLPM   = "RI-CLPM"
      )
  ) %>%
  select(run, occasion, scenario, estimand, method, pval)

#####################################################
# Build type I error data frame
#####################################################

type1_df <- pval_df %>%

  # attach true parameter values
  left_join(true_params, by = "estimand") %>%

  # only estimands where the null is true (theta = 0)
  filter(true == 0) %>%

  # indicator: reject at alpha
  mutate(
    reject = (pval <= 0.05)
  ) %>%

  # type I error rate and Monte Carlo SE
  group_by(scenario, occasion, estimand, method) %>%
  reframe(
    nsim = n(),

    # type I error rate
    type1 = mean(reject),

    # Monte Carlo SE of type I error rate
    mcse_type1 =
      sqrt(type1 * (1 - type1) / nsim)
  ) %>%

  # clean factor ordering
  mutate(
    scenario = factor(scenario, levels = c("constant", "stepwise")),
    method   = factor(
      method,
      levels = c(
        "CLPM",
        "Lin adj",
        "CLPM linear BCA",
        "CLPM XGB BCA",
        "DPM",
        "RI-CLPM"
      )
    )
  )

#####################################################
# Colour palette for methods (shared across plots)
#####################################################

pal_method <- qualitative_hcl(
  n = nlevels(type1_df$method),
  palette = "Set 2"
)
names(pal_method) <- levels(type1_df$method)

#####################################################
# Plot: type I error (only for null-true estimands)
#####################################################

plot_type1 <- type1_df %>%

  ggplot(aes(
    x     = occasion,
    y     = type1,
    color = method,
    group = method
  )) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  geom_errorbar(
    aes(
      ymin = type1 - mcse_type1,
      ymax = type1 + mcse_type1
    ),
    width = 0.15,
    linewidth = 0.4
  ) +

  # nominal alpha reference line
  geom_hline(yintercept = 0.05, linetype = "dashed", linewidth = 0.4) +

  # facet by scenario
  ggh4x::facet_wrap2(
    ~ scenario,
    nrow = 1,
    axes = "y"
  ) +

  # colours
  scale_color_manual(values = pal_method) +

  # labels
  labs(
    x = "Occasion",
    y = "Type I error",
    color = NULL
  ) +

  # scientific theme
  theme_classic(base_size = 13) +
  theme(
    panel.spacing.x = unit(1.2, "cm"),
    strip.text      = element_text(size = 14),
    axis.title      = element_text(size = 13),
    axis.text       = element_text(size = 12),
    legend.position = "bottom",
    legend.text     = element_text(size = 12)
  )

#####################################################
# Display plot
#####################################################

plot_type1

```

### 4.3: Type II Error Simulation Study 2


```{r}

#####################################################
# Build power data for all lagged parameters
#####################################################

# true parameters from the data-generating model
true_params <- tibble(
  estimand = c("XY", "YX", "AX", "AY"),
  true     = c(
    A[2, 1],  # XY
    A[1, 2],  # YX (usually 0)
    A[1, 1],  # AX
    A[2, 2]   # AY
  )
)

#####################################################
# --- p-values ---
#####################################################

pval_df <- results_sim2 %>%

  # only occasions with lagged parameters
  filter(occasion >= 2) %>%

  # reshape p-values to long format
  pivot_longer(
    cols = matches("^p(XY|YX|A|AY)_"),
    names_to  = "param",
    values_to = "pval"
  ) %>%

  # drop failed fits
  filter(!is.na(pval)) %>%

  # identify estimand type and method
  mutate(
    estimand = case_when(
      str_detect(param, "^pXY_") ~ "XY",
      str_detect(param, "^pYX_") ~ "YX",
      str_detect(param, "^pA_")  ~ "AX",
      str_detect(param, "^pAY_") ~ "AY"
    ),

    method = param %>%
      str_remove("^p(XY|YX|A|AY)_") %>%
      recode(
        CLPM      = "CLPM",
        CLPM_Adj  = "Lin adj",
        CLPM_LBCA = "CLPM linear BCA",
        CLPM_XGB  = "CLPM XGB BCA",
        DPM       = "DPM",
        RI_CLPM   = "RI-CLPM"
      )
  ) %>%
  select(run, occasion, scenario, estimand, method, pval)

#####################################################
# Build power data frame
#####################################################

power_df <- pval_df %>%

  # attach true parameter values
  left_join(true_params, by = "estimand") %>%

  # only estimands where the null is false
  filter(true != 0) %>%

  # indicator: reject at alpha
  mutate(
    reject = (pval <= 0.05)
  ) %>%

  # power and Monte Carlo SE
  group_by(scenario, occasion, estimand, method) %>%
  reframe(
    nsim = n(),

    # power
    power = mean(reject),

    # Monte Carlo SE of power
    mcse_power =
      sqrt(power * (1 - power) / nsim)
  ) %>%

  # clean factor ordering
  mutate(
    scenario = factor(scenario, levels = c("constant", "stepwise")),
    method   = factor(
      method,
      levels = c(
        "CLPM",
        "Lin adj",
        "CLPM linear BCA",
        "CLPM XGB BCA",
        "DPM",
        "RI-CLPM"
      )
    )
  )

#####################################################
# Colour palette for methods (shared across plots)
#####################################################

pal_method <- qualitative_hcl(
  n = nlevels(power_df$method),
  palette = "Set 2"
)
names(pal_method) <- levels(power_df$method)

#####################################################
# Plot helper (used for all three estimands)
#####################################################

plot_power <- function(df, estimand_label) {

  df %>%
    filter(estimand == estimand_label) %>%

    ggplot(aes(
      x     = occasion,
      y     = power,
      color = method,
      group = method
    )) +
    geom_line(linewidth = 0.9) +
    geom_point(size = 2) +
    geom_errorbar(
      aes(
        ymin = power - mcse_power,
        ymax = power + mcse_power
      ),
      width = 0.15,
      linewidth = 0.4
    ) +

    # reference line for high power
    geom_hline(yintercept = 0.80, linetype = "dashed", linewidth = 0.4) +

    # facet by scenario
    ggh4x::facet_wrap2(
      ~ scenario,
      nrow = 1,
      axes = "y"
    ) +

    # colours
    scale_color_manual(values = pal_method) +

    # labels
    labs(
      x = "Occasion",
      y = "Power",
      color = NULL
    ) +

    # scientific theme
    theme_classic(base_size = 13) +
    theme(
      panel.spacing.x = unit(1.2, "cm"),
      strip.text      = element_text(size = 14),
      axis.title      = element_text(size = 13),
      axis.text       = element_text(size = 12),
      legend.position = "bottom",
      legend.text     = element_text(size = 12)
    )
}

#####################################################
# Build and display all three power plots
#####################################################

plot_power_XY <- plot_power(power_df, "XY")
plot_power_AX <- plot_power(power_df, "AX")
plot_power_AY <- plot_power(power_df, "AY")

plot_power_XY
plot_power_AX
plot_power_AY

```

Now we need to concatenate the plots:

```{r}

library(patchwork)

# create legendless plots
plot_relbias_XY_noleg <- plot_relbias_XY +
  theme(legend.position = "none")

# combine the two
combined_XY <- plot_relbias_XY_noleg / plot_rmse_XY

ggsave(
  filename = "combined_XY_relbias_RMSE.png",
  plot     = combined_XY,
  path     = "C:/Users/Admin/Desktop",
  width    = 10,
  height   = 8,
  dpi      = 300
)

```